title: "GTCI Figures"
author: "Emerson Johnston"
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE
)

# Resolve this Rmd's directory, then locate setup_publication.R from there.
rmd_dir <- tryCatch(
  dirname(normalizePath(knitr::current_input(dir = TRUE), winslash = "/", mustWork = TRUE)),
  error = function(e) normalizePath(getwd(), winslash = "/", mustWork = TRUE)
)

setup_candidates <- c(
  file.path(rmd_dir, "inputs", "scripts", "setup_publication.R"),
  file.path(rmd_dir, "..", "inputs", "scripts", "setup_publication.R"),
  file.path(getwd(), "inputs", "scripts", "setup_publication.R")
)

setup_path <- setup_candidates[file.exists(setup_candidates)][1]

if (is.na(setup_path)) {
  stop(
    "Could not locate inputs/scripts/setup_publication.R from the Rmd location or working directory.",
    call. = FALSE
  )
}

source(setup_path)
```

# Introduction

### Figure 0.00 - Total inbound internationally mobile students
```{r}
meta <- get_meta("FIG000")

df_000 <- read_fig("FIG000") %>%
  rename(Country = 1, value = 2) %>%
  mutate(value = as.numeric(value)) %>%
  arrange(desc(value)) %>%
  mutate(Country = fct_reorder(Country, value))

labels_000 <- first_last_peak(df_000, value, value)

FIG000 <- ggplot(df_000, aes(value, Country, fill = Country)) +
  geom_col(width = 0.75, alpha = ALPHA_COL) +
  geom_text(aes(label = label_comma()(value), hjust = -0.2),
            family = FONT_FAMILY,
            fontface = "bold",
            size = SMALL_LABEL_SIZE / .pt,
            color = "grey25") +
  scale_x_continuous(labels = label_comma(),
                     expand = EXPAND_LABEL_RIGHT) +
  scale_fill_manual(values = highlight_colors(
    all_values  = levels(df_000$Country),
    highlighted = "United States of America",
    palette     = tpa_colors[1],
    default     = tpa_colors[2]
  )) +
  guides(fill = "none") +
  labs_meta(meta, x = "Inbound internationally mobile students (2023)") +
  coord_cartesian(clip = "off") +
  theme_horizontal_bar() +
  theme(
    axis.text.y = element_text(size = SMALL_LABEL_SIZE),
    plot.margin = publication_margin(right = PUB$spacing$edge_wide_pt)
  )

save_chart(FIG000, "fig_0_00", size = "medium")
```

### Figure 0.01 - Definition of 'International Students' Funnel
```{r fig_intro_funnel}
# Rounded-rectangle path builder (separate x/y radius so corners stay round
# when the panel isn't square).
rounded_rect <- function(xmin, xmax, ymin, ymax, rx = 0.02, ry = 0.02, n = 10) {
  rx <- min(rx, (xmax - xmin) / 2); ry <- min(ry, (ymax - ymin) / 2)
  arc <- function(cx, cy, start) {
    a <- seq(start, start + pi/2, length.out = n)
    cbind(cx + rx * cos(a), cy + ry * sin(a))
  }
  pts <- rbind(
    arc(xmax - rx, ymin + ry, -pi/2),
    arc(xmax - rx, ymax - ry,  0),
    arc(xmin + rx, ymax - ry,  pi/2),
    arc(xmin + rx, ymin + ry,  pi)
  )
  data.frame(x = pts[, 1], y = pts[, 2])
}

GREY_FILL <- tpa_colors[9]   # "#B9B4AC"
RED_FILL  <- tpa_colors[1]   # "#8C1515"

blocks <- tibble::tribble(
  ~label,                    ~count,   ~xmin, ~xmax, ~ymin, ~ymax, ~fill,     ~tcol,
  "Foreign-born",            "47.8M",   0.00,  2.00,  0.75,  1.00,  RED_FILL,  "white",
  "Noncitizens",             "~22.8M",  0.00,  1.20,  0.50,  0.75,  RED_FILL,  "white",
  "Naturalized\ncitizens",   "~25M",    1.20,  2.00,  0.00,  0.75,  GREY_FILL, "grey20",
  "Temporary\nvisa holders", "~3.5M",   0.00,  0.60,  0.25,  0.50,  RED_FILL,  "white",
  "Green card\nholders",     "12.8M",   0.60,  0.94,  0.00,  0.50,  GREY_FILL, "grey20",
  "Unauthorized\n& other",   "~11M",    0.94,  1.20,  0.00,  0.50,  GREY_FILL, "grey20",
  "Int.\nstudents",  "~1.1M",   0.00,  0.20,  0.00,  0.25,  RED_FILL,  "white",
  "Int.\nworkers",   "~1.4M",   0.20,  0.40,  0.00,  0.25,  RED_FILL,  "white",
  "Exchange\n& other",       "~1M",     0.40,  0.60,  0.00,  0.25,  GREY_FILL, "grey20"
) |>
  dplyr::mutate(
    lx     = (xmin + xmax) / 2,
    ly     = (ymin + ymax) / 2,
    n_line = vapply(strsplit(label, "\n"), length, integer(1)),
    size   = c(TITLE_SIZE, SUBTITLE_SIZE, AXIS_SIZE + 0.5, AXIS_SIZE,
               COMPACT_LABEL_SIZE, SMALL_LABEL_SIZE + 0.5, SWATCH_SIZE,
               SWATCH_SIZE, SWATCH_SIZE) / .pt,
    c_size = c(DATA_LABEL_SIZE - 0.5, DATA_LABEL_SIZE - 0.5, COMPACT_LABEL_SIZE,
               SMALL_LABEL_SIZE + 0.5, SMALL_LABEL_SIZE, SWATCH_SIZE,
               MILESTONE_LABEL_SIZE, MILESTONE_LABEL_SIZE, MILESTONE_LABEL_SIZE) / .pt,
    lab_y  = ly + dplyr::if_else(n_line == 2, 0.030, 0.014),
    cnt_y  = ly - dplyr::if_else(n_line == 2, 0.055, 0.034)
  )

poly <- do.call(rbind, lapply(seq_len(nrow(blocks)), function(i) {
  b  <- blocks[i, ]
  rr <- rounded_rect(b$xmin, b$xmax, b$ymin, b$ymax, rx = 0.018, ry = 0.03)
  rr$label <- b$label
  rr$fill  <- b$fill
  rr
}))

fig_funnel <- ggplot() +
  geom_polygon(data = poly, aes(x = x, y = y, group = label, fill = fill),
               colour = "white", linewidth = 1) +
  geom_text(data = blocks,
            aes(x = lx, y = lab_y, label = label, colour = tcol, size = size),
            family = FONT_FAMILY, fontface = "bold", lineheight = 0.9) +
  geom_text(data = blocks,
            aes(x = lx, y = cnt_y, label = count, colour = tcol, size = c_size),
            family = FONT_FAMILY) +
  scale_fill_identity() +
  scale_colour_identity() +
  scale_size_identity() +
  scale_x_continuous(limits = c(0, 2), expand = EXPAND_NONE) +
  scale_y_continuous(limits = c(0, 1), expand = EXPAND_NONE) +
  labs(caption = "Not to scale. Census (ACS) and DHS/OHSS estimates, 2023-24; sources differ and do not sum.") +
  theme_arrows() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )


save_chart(fig_funnel, "fig_0_01")
```

# Chapter 1

### Figure 1.01 - How Many Research Doctorates Are Awarded by U.S. Universities?
```{r figure 003}
meta <- get_meta("FIG101")

df_003 <- read_fig("FIG101") %>%
  select(Year, value = 2) %>%
  mutate(across(everything(), as.numeric))

labels_003 <- first_last_peak(df_003, Year, value) %>%
  mutate(lab = ifelse(value < 1000,
                      format(round(value), big.mark = ","),
                      fmt_num_k(value)))

FIG101 <- ggplot(df_003, aes(Year, value)) +
  geom_line(linewidth = LINE_WIDTH, color = tpa_colors[1]) +
  geom_label_repel_tpa(
    data         = labels_003,
    mapping      = aes(label = lab),
    color        = tpa_colors[1],
    nudge_y      = 3000,
    with_segment = TRUE
  ) +
  scale_x_years(df_003$Year) +
  scale_y_continuous(labels = fmt_num_k,
                     expand = EXPAND_LINE_TALL) +
  labs_meta(meta, y = "Doctorate recipients") +
  theme_tpa()
save_chart_aligned(FIG101, "fig_1_01")
FIG101
```



### Figure 1.02 - How Has American Degree Production Changed Over Time
```{r figure 98}
meta <- get_meta("FIG102")

m1 <- function(x) sprintf("%.1fM", x / 1e6)

fmt_start <- function(x) {
  out <- dplyr::case_when(
    x >= 1e6 ~ sprintf("%.1fM", x / 1e6),
    TRUE     ~ sprintf("%.1fK", x / 1e3)
  )
  sub("^0\\.", ".", out)
}

df_098 <- read_levels("FIG102", value_col = "Count",
                      levels_order = c("Bachelors", "Masters", "Doctorate")) %>%
  mutate(Level = fct_recode(Level,
                            "Bachelor's" = "Bachelors",
                            "Master's"   = "Masters"))

last_098 <- last_by_group(df_098, Year, Level) %>%
  mutate(
    .label = paste0(Level, ": ", fmt_start(Count))
  ) %>%
  prepare_right_labels(df_098$Year)

FIG102 <- ggplot(df_098, aes(Year, Count, color = Level, group = Level)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_label_repel_tpa(
    data    = last_098,
    mapping = aes(x = .label_x, label = .label),
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_098$Year, right_labels = last_098) +
  scale_y_continuous(labels = m1, expand = EXPAND_LINE) +
  scale_color_levels() +
  guides(color = "none") +
  labs_meta(meta, y = "Degrees awarded") +
  coord_cartesian(clip = "off") +
  theme_tpa() +
  theme_right_labels(last_098)

save_chart_aligned(FIG102, "fig_1_02")
```

### Figure 1.03 - How Many STEM Degrees Are Awarded in the United States
```{r figure 0001}
meta <- get_meta("FIG103")
df_001 <- read_levels("FIG103", value_col = "Count",
                      levels_order = c("Doctorate", "Masters", "Bachelors")) %>%
  mutate(Level = fct_recode(Level,
                            "Bachelor's" = "Bachelors",
                            "Master's"   = "Masters"))
FIG103 <- ggplot(df_001, aes(Year, Count, fill = Level)) +
  geom_area(alpha = ALPHA_AREA) +
  scale_x_years(df_001$Year) +
  scale_y_continuous(breaks = c(200000, 400000, 600000, 800000),
                     labels = fmt_num_k,
                     expand = EXPAND_BAR) +
  scale_fill_levels(reverse_legend = FALSE) +
  labs_meta(meta, y = "Degrees awarded") +
  coord_cartesian(ylim = c(200000, 800000)) +
  theme_tpa()
save_chart_aligned(FIG103, "fig_1_03")
```


### Figure 1.04 - International Students in U.S. Higher Education
```{r figure 009}
meta <- get_meta("FIG104")

df_009 <- read_fig("FIG104") %>% 
  rename(Year = 1, Total = 2, Share = 3) %>%
  mutate(across(everything(), as.numeric),
         Share = fill_gaps(Year, Share))

scale_factor <- dual_axis_scale(
  primary_max   = max(df_009$Total, na.rm = TRUE),
  secondary_max = 0.07
)

FIG104 <- ggplot(df_009, aes(Year)) +
  geom_col(aes(y = Total, fill = "Total"),
           width = 0.7, alpha = ALPHA_COL_LIGHT) +
  geom_line(aes(y = Share * scale_factor, color = "Share"),
            linewidth = LINE_WIDTH, na.rm = TRUE) +
  scale_x_years(df_009$Year) +
  scale_y_continuous(
    name   = "Count of international students",
    labels = fmt_kbm,
    expand = EXPAND_BAR,
    sec.axis = sec_axis_scaled(scale_factor,
                               name   = "Share of total US enrollment",
                               labels = label_percent(accuracy = 1))
  ) +
  scale_fill_manual(name = NULL, values = c("Total" = tpa_colors[2])) +
  scale_color_manual(name = NULL, values = c("Share" = tpa_colors[1])) +
  labs_meta(meta) +
  theme_dual_axis()

save_chart_aligned(FIG104, "fig_1_04")
```

### Figure 1.05 - What Share of All Degrees Awarded in the US are Earned by Int. Students
```{r figure 099}
meta <- get_meta("FIG105")
pct  <- label_percent(accuracy = 1)
df_099 <- read_levels(
  "FIG105",
  value_col    = "Share",
  levels_order = c("Bachelors", "Masters",
                  "Doctorates")
) %>%
  filter(!is.na(Share)) %>%
  mutate(Level = fct_recode(Level,
                            "Bachelor's" = "Bachelors",
                            "Master's"   = "Masters"))
last_099 <- last_by_group(df_099, Year, Level) %>%
  mutate(
    .label = paste0(pct(Share), " - ", Level)
  ) %>%
  prepare_right_labels(df_099$Year)
first_099 <- df_099 %>% group_by(Level) %>% slice_min(Year, n = 1) %>% ungroup()

FIG105 <- ggplot(df_099, aes(Year, Share, color = Level, group = Level)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_label_repel_tpa(
    data    = last_099,
    mapping = aes(x = .label_x, label = .label),
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  geom_label_repel_tpa(
    data         = first_099,
    mapping      = aes(label = pct(Share)),
    nudge_x      = -0.5,
    hjust        = 1,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_099$Year, right_labels = last_099) +
  scale_y_continuous(labels = pct, expand = EXPAND_LINE_WIDE) +
  scale_color_levels() +
  guides(color = "none") +
  labs_meta(meta, y = "Share of completions") +
  coord_cartesian(ylim = c(0, .35), clip = "off") +
  theme_tpa() +
  theme_right_labels(last_099)
save_chart_aligned(FIG105, "fig_1_05")
FIG105
```

### Figure 1.06 - What Share of STEM Degrees Are Earned by International Students?
```{r figure 005}
meta <- get_meta("FIG106")

pct  <- label_percent(accuracy = 1)
df_005 <- read_levels(
  "FIG106",
  value_col    = "Share",
  levels_order = c("Bachelors", "Masters",
                   "Doctorates")
) %>%
  filter(!is.na(Share)) %>%
  mutate(Level = fct_recode(Level,
                            "Bachelor's" = "Bachelors",
                            "Master's"   = "Masters"))

last_005 <- last_by_group(df_005, Year, Level) %>%
  mutate(
    .label = paste0(pct(Share), ": ", Level)
  ) %>%
  prepare_right_labels(df_005$Year)
first_005 <- df_005 %>% group_by(Level) %>% slice_min(Year, n = 1) %>% ungroup()

FIG106 <- ggplot(df_005, aes(Year, Share, color = Level, group = Level)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_label_repel_tpa(
    data    = last_005,
    mapping = aes(x = .label_x, label = .label),
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  geom_label_repel_tpa(
    data         = first_005,
    mapping      = aes(label = pct(Share)),
    nudge_x      = -0.5,
    hjust        = 1,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_005$Year, pad_left = 0.05, right_labels = last_005) +
  scale_y_continuous(labels = pct, expand = EXPAND_LINE_WIDE) +
  scale_color_levels() +
  guides(color = "none") +
  labs_meta(meta, y = "Share of completions") +
  coord_cartesian(clip = "off") +
  theme_tpa() +
  theme_right_labels(last_005)

save_chart_aligned(FIG106, "fig_1_06")
FIG106
```

### Figure 1.07 - Projected STEM PhD Production by Country### Figure 1.07 - Projected STEM PhD Production by Country
```{r figure 008}
meta <- get_meta("FIG107")
df_008 <- read_fig("FIG107") %>%
  pivot_observed_projected() %>%
  order_by_latest(Country, value)
common_end <- max(df_008$Year)
labels_008 <- last_by_group(df_008, Year, Country) %>%
  mutate(
    .label = paste0(Country, ": ", fmt_num_k(value)),   # K-formatted values
    Year   = common_end                                  # anchor to shared right edge
  ) %>%
  prepare_right_labels(df_008$Year)

# Reserve the same right gutter as 1.08 so both panels end at the same x.
SHARED_PROJ_RIGHT_LIMIT <- attr(
  prepare_right_labels(
    last_by_group(
      read_fig("FIG108") %>% pivot_observed_projected() %>% order_by_latest(Country, value),
      Year, Country
    ) %>% mutate(.label = paste0(Country, ": ", round(value))),
    (read_fig("FIG108") %>% pivot_observed_projected())$Year
  ),
  "tpa_right_limit"
)
labels_008 <- set_right_limit(labels_008, SHARED_PROJ_RIGHT_LIMIT)

proj_start <- proj_start_year(df_008)
y_top      <- 150000                       # fixed ceiling; breaks and band key off this
milestones <- tribble(
  ~year, ~label,
  2019, "China Tops US in STEM PhDs",
  2033, "India Tops US in STEM PhDs",
  2042, "India Tops China in STEM PhDs,\nBoth now produce 1.5x US",
  2049, "India Doubles US STEM Production"
)
FIG107 <- ggplot(df_008, aes(Year, value, color = Country, group = Country)) +
  annotate_projected(proj_start, max(df_008$Year),
                     label_y = 140000, ymin = 0, ymax = 150000) +   # bound band to plot range
  geom_op_lines(df_008) +
  annotate_milestones(
    milestones, y_top* 1.035,
    gap_low  = 0.30,
    gap_high = 0.58,
    bg_for = \(year) if_else(year >= proj_start, "grey94", "white")
  ) +
  geom_label_repel_tpa(
    data    = labels_008,
    mapping = aes(x = .label_x, label = .label),
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_008$Year, right_labels = labels_008) +
  scale_y_continuous(labels = fmt_num_k0,
                     breaks = seq(0, 150000, by = 30000),
                     limits = c(0, 150000),
                     expand = expansion(mult = c(0, 0))) +
  scale_color_tpa() +
  guides(color = "none") +
  labs_meta(meta, y = "STEM PhDs per year") +
  coord_cartesian(clip = "off") +
  theme_op() +
  theme_right_labels(labels_008)
save_chart_aligned(FIG107, "fig_1_07")
FIG107
```

### Figure 1.08 - Projected STEM PhD Production, per Capita, by Country
```{r figure 007}
meta <- get_meta("FIG108")
df_007 <- read_fig("FIG108") %>%
  pivot_observed_projected() %>%
  order_by_latest(Country, value)
labels_007 <- last_by_group(df_007, Year, Country) %>%
  mutate(
    .label = paste0(Country, ": ", round(value))
  ) %>%
  prepare_right_labels(df_007$Year)
labels_007 <- set_right_limit(labels_007, SHARED_PROJ_RIGHT_LIMIT)
proj_start <- proj_start_year(df_007)
y_top      <- 500                          # fixed ceiling; breaks and band key off this
FIG108 <- ggplot(df_007, aes(Year, value, color = Country, group = Country)) +
  annotate_projected(proj_start, max(df_007$Year),
                     label_y = 466.75, ymin = 0, ymax = 500) +   # bound band to plot range
  geom_op_lines(df_007) +
  geom_label_repel_tpa(
    data    = labels_007,
    mapping = aes(x = .label_x, label = .label),
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_007$Year, right_labels = labels_007) +
  scale_y_continuous(labels = label_comma(),
                     breaks = seq(0, 500, by = 100),
                     limits = c(0, 500),
                     expand = expansion(mult = c(0, 0))) +
  scale_color_tpa() +
  guides(color = "none") +
  labs_meta(meta, y = "STEM PhDs per year") +
  coord_cartesian(clip = "off") +
  theme_op() +
  theme_right_labels(labels_007)
save_chart_aligned(FIG108, "fig_1_08")
FIG108
```

# Chapter 2

### Figure 2.01 - Which Countries Acccount for the Largest Share of International Undergraduates? 
```{r FIG203}
meta <- get_meta("FIG201")

df_109 <- read_fig("FIG201") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Country", values_to = "Count") %>%
  filter(!is.na(Count))

top_countries <- df_109 %>%
  filter(Year == max(Year)) %>%
  slice_max(Count, n = 8) %>%
  pull(Country)

df_109 <- df_109 %>%
  filter(Country %in% top_countries) %>%
  order_by_latest(Country, Count)

labels_109 <- last_by_group(df_109, Year, Country) %>%
  mutate(
    .label = paste0(Country, ": ", label_percent(accuracy = 0.1)(Count))
  ) %>%
  prepare_right_labels(df_109$Year)

FIG203 <- ggplot(df_109, aes(Year, Count, color = Country, group = Country)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_label_repel_tpa(
    data    = labels_109,
    mapping = aes(x = .label_x, label = .label),
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_109$Year, right_labels = labels_109) +
  scale_y_continuous(labels = label_percent(accuracy = 1),
                expand = EXPAND_LINE) +
  scale_color_tpa() +
  guides(color = "none") +
  labs_meta(meta, y = "Share of Enrolled International Students") +
  coord_cartesian(ylim = c(0, 0.4), clip = "off") +
  theme_tpa() +
  theme_right_labels(labels_109)

save_chart(FIG203, "fig_2_01")
```


### Figure 2.02 - Which Countries Account for the Largest Share of International Graduates? 
```{r 110}
meta <- get_meta("FIG202")

df_111 <- read_fig("FIG202") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Country", values_to = "Count") %>%
  filter(!is.na(Count))

top_countries <- df_111 %>%
  filter(Year == max(Year)) %>%
  slice_max(Count, n = 8) %>%
  pull(Country)

df_111 <- df_111 %>%
  filter(Country %in% top_countries) %>%
  order_by_latest(Country, Count)

labels_111 <- last_by_group(df_111, Year, Country) %>%
  mutate(
    .label = paste0(Country, ": ", label_percent(accuracy = 0.1)(Count))
  ) %>%
  prepare_right_labels(df_111$Year)

FIG204 <- ggplot(df_111, aes(Year, Count, color = Country, group = Country)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_label_repel_tpa(
    data    = labels_111,
    mapping = aes(x = .label_x, label = .label),
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_111$Year, right_labels = labels_111) +
  scale_y_continuous(labels = label_percent(accuracy = 1),
                expand = EXPAND_LINE) +
  scale_color_tpa() +
  guides(color = "none") +
  labs_meta(meta, y = "Share of Enrolled International Students") +
  coord_cartesian(ylim = c(0, 0.4), clip = "off") +
  theme_tpa() +
  theme_right_labels(labels_111)

save_chart(FIG204, "fig_2_02")
```

### Figure 2.03 - International Students Enrolled in STEM Fields by Discipline
```{r figure 014}
meta <- get_meta("FIG203")

df_014 <- read_fig("FIG203") %>%
  select(-`Non STEM`,-`Social Sciences`) %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Field", values_to = "Count") %>%
  filter(!is.na(Count)) %>%
  order_by_latest(Field, Count)

labels_014 <- last_by_group(df_014, Year, Field) %>%
  mutate(
    .label = case_when(
      Field == "Math and Computer Science" ~
        paste0(
          "Math and Computer\nScience: ",
          label_comma()(Count)
        ),

      Field == "Physical and Life Sciences" ~
        paste0(
          "Physical and Life\nSciences: ",
          label_comma()(Count)
        ),

      TRUE ~ paste0(
        Field,
        ": ",
        label_comma()(Count)
      )
    )
  ) %>%
  prepare_right_labels(
    df_014$Year,
    allow_multiline = TRUE
  )

FIG205 <- ggplot(df_014, aes(Year, Count, color = Field, group = Field)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_label_repel_tpa(
    data    = labels_014,
    mapping = aes(x = .label_x, label = .label),
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(
  df_014$Year,
  right_labels = labels_014
  ) +
  scale_y_continuous(labels = fmt_num_k,
                   expand = EXPAND_LINE) +
  scale_color_tpa() +
  guides(color = "none") +
  labs_meta(meta, y = "Enrolled students") +
  coord_cartesian(clip = "off") +
  theme_tpa() +
  theme_right_labels(labels_014)

save_chart(FIG205, "fig_2_03")
```


### Figure 2.04 - International Students Enrolled in STEM Fields by Country
```{r figure 013}
meta <- get_meta("FIG204")

df_013 <- read_fig("FIG204") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Country", values_to = "Count") %>%
  filter(!is.na(Count))

top_countries <- df_013 %>%
  group_by(Country) %>%
  summarise(Count = sum(Count), .groups = "drop") %>%
  slice_max(Count, n = 8) %>%
  pull(Country)

df_013 <- df_013 %>%
  filter(Country %in% top_countries) %>%
  order_by_latest(Country, Count)

labels_013 <- last_by_group(df_013, Year, Country) %>%
  mutate(
    .label = paste0(Country, ": ", label_comma()(Count))
  ) %>%
  prepare_right_labels(df_013$Year)

FIG206 <- ggplot(df_013, aes(Year, Count, color = Country, group = Country)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_label_repel_tpa(
    data    = labels_013,
    mapping = aes(x = .label_x, label = .label),
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_013$Year, right_labels = labels_013) +
  scale_y_log10(labels = fmt_num_k,
                breaks = c(0, 1000, 5000, 10000, 50000, 100000, 250000),
                expand = EXPAND_LINE) +
  scale_color_tpa() +
  guides(color = "none") +
  labs_meta(meta, y = "Enrolled students (log scale)") +
  coord_cartesian(clip = "off") +
  theme_tpa() +
  theme_right_labels(labels_013)

save_chart(FIG206, "fig_2_04")
```


### Figure 2.05 - Economic Contributuions of International Students 
```{r figure 010}
meta <- get_meta("FIG205")

df_010 <- read_fig("FIG205") %>%
  rename(value = 2) %>%
  mutate(Year  = as.numeric(Year),
         value = as.numeric(value) / 1e9)

labels <- first_last_peak(df_010, Year, value)

FIG209 <- ggplot(df_010, aes(Year, value)) +
  geom_col(fill = tpa_colors[1], width = 0.7, alpha = ALPHA_COL) +
  geom_text_tpa(data    = labels,
                mapping = aes(label = sprintf("%.2f", value)),
                color   = tpa_colors[1]) +
  coord_cartesian(clip = "off") +
  scale_x_years(df_010$Year) +
  scale_y_continuous(labels = label_dollar(suffix = "B"),
                     expand = EXPAND_BAR_TALL) +
  labs_meta(meta, y = "USD (billions)") +
  theme_tpa()

save_chart(FIG209, "fig_2_05")
```


### Figure 2.06 - How Much Did International Enrollment Change in 2025
```{r FIG210}
meta <- get_meta("FIG206")

df_104 <- read_fig("FIG206")

FIG210 <- ggplot() +
  block_arrow_layers(df_104, Category, Change,
                     fills = arrow_fill,
                     divider_after = 1) +
  labs_meta(meta) +
  theme_arrows()

save_chart(FIG210, "fig_2_06", size = "standard")
```

### Figure 2.07 - Where are American Institutions Focusing Their Recruitment Efforts Abroad?
```{r FIG211}
meta <- get_meta("FIG207")

panel_colors <- c(
  "Undergraduate" = tpa_colors[1],   
  "Graduate"      = tpa_colors[2]    
)

track_max <- 1.0   # grey track length; nudge up toward ~1.1 if fills look too long

df_107 <- read_fig("FIG207") %>%
  transmute(
    Rank                  = as.integer(str_remove(Rank, "Rank ")),
    Undergraduate_Country = `Undergraduate Country`,
    Undergraduate_Value   = as.numeric(`Undergraduate Percent`),
    Graduate_Country      = `Graduate Country`,
    Graduate_Value        = as.numeric(`Graduate Percent`)
  ) %>%
  pivot_longer(-Rank, names_to = c("Panel", ".value"), names_sep = "_") %>%
  mutate(
    Panel   = factor(Panel, levels = c("Undergraduate", "Graduate")),
    Country = str_wrap(Country, width = 18),
    Country = reorder_within(Country, -Rank, Panel)   # rank 1 on top, per panel
  )

FIG211 <- ggplot(df_107, aes(y = Country)) +
  geom_col(aes(x = track_max), fill = "grey90", width = 0.3) +
  geom_col(aes(x = Value, fill = Panel), width = 0.3) +
  geom_text(aes(x = track_max, label = label_percent(accuracy = 1)(Value)),
            hjust = -0.25, fontface = "bold", color = "grey20",
            size = DATA_LABEL_SIZE / .pt, family = FONT_FAMILY) +
  facet_wrap(~ Panel, scales = "free_y") +
  scale_x_continuous(expand = expansion(mult = c(0, 0.14))) +
  scale_y_reordered(expand = expansion(add = 1)) +
  scale_fill_manual(values = panel_colors) +
  coord_cartesian(clip = "off") +
  labs_meta(meta) +
  theme_bullet()

save_chart(FIG211, "fig_2_07")
```

### Figure 2.08 - What Do Institutions Cite as the Reason for Decreasing International Enrollments?
```{r FIG212}
meta <- get_meta("FIG208")

df_106 <- read_fig("FIG208") %>%
  rename(Reason = 1) %>%
  mutate(across(-Reason, as.numeric)) %>%
  pivot_longer(-Reason, names_to = "Cohort", values_to = "Share") %>%
  filter(!is.na(Share)) %>%
  mutate(
    Cohort = factor(Cohort, levels = c("Fall 2018", "Fall 2021", "Fall 2025")),
    Reason = str_wrap(Reason, width = 40),                 # <- lower width = narrower text column
    Reason = factor(Reason, levels = rev(unique(Reason)))
  ) 

cohort_colors <- c(
  "Fall 2018" = tpa_colors[1],
  "Fall 2021" = tpa_colors[2],
  "Fall 2025" = tpa_colors[3]
)

# Spread crowded value labels horizontally; gap is in percentage points.
gap <- 4
labels_106 <- df_106 %>%
  group_by(Reason) %>%
  arrange(Share, .by_group = TRUE) %>%
  mutate(
    spread  = accumulate(Share, ~ max(.y, .x + gap)),
    label_x = spread - (mean(spread) - mean(Share))
  ) %>%
  ungroup()

FIG212 <- ggplot(df_106, aes(Share, Reason, color = Cohort)) +
  geom_point(size = 2.5) +
  geom_text(
    data = labels_106,
    aes(x = label_x, label = paste0(Share, "%")),
    nudge_y = -0.34,
    size = 8 / .pt,
    fontface = "bold",
    family = FONT_FAMILY,
    show.legend = FALSE
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.07, 0.07))) +
  scale_y_discrete(expand = expansion(add = c(0.8, 0.6))) +
  scale_color_manual(values = cohort_colors) +
  guides(color = guide_legend(override.aes = list(size = 3.5))) +
  coord_cartesian(clip = "off") +
  labs_meta(meta) +
  theme_dotplot()

save_chart(FIG212, "fig_2_08", size="standard")
```


# Chapter 3

### Figure 3.01 - Where is international talent in the education-to-work pipeline?
```{r FIG201}
meta <- get_meta("FIG301")

df_102 <- read_fig("FIG301") %>%
  rename(
    `Graduate (Other)` = `Graduate - Unspecified`,
    `Non-degree`             = `Non-Degree`,
    OPT                      = `Optional Practical Training (OPT)`
  ) %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Level", values_to = "Count") %>%
  order_by_latest(Level, Count)

labels_102 <- last_by_group(df_102, Year, Level) %>%
  mutate(
    .label = paste0(Level, ": ", label_comma()(Count))
  ) %>%
  prepare_right_labels(df_102$Year)

FIG201 <- ggplot(df_102, aes(Year, Count, color = Level, group = Level)) +
  geom_line(linewidth = 1) +
  geom_label_repel_tpa(
    data    = labels_102,
    mapping = aes(x = .label_x, label = .label),
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_102$Year, right_labels = labels_102) +
  scale_y_continuous(labels = fmt_num_k,
                     expand = expansion(mult = c(0.02, 0.05))) +
  scale_color_manual(values = acad_level_colors) +
  guides(color = "none") +
  labs_meta(meta, y = "International students enrolled") +
  coord_cartesian(clip = "off") +
  theme_tpa() +
  theme_right_labels(labels_102)

save_chart(FIG201, "fig_3_01")
```

### Figure 3.02 - International Postdoctoral Researchers by Field
```{r figure 020}
meta <- get_meta("FIG302")

df_020 <- read_fig("FIG302") %>%
  rename(Year = 1,
         Science     = 2,
         Engineering = 3,
         Health      = 4) %>%
  mutate(across(everything(), as.numeric)) %>%
  pivot_longer(-Year, names_to = "Field", values_to = "Share") %>%
  filter(!is.na(Share)) %>%
  order_by_latest(Field, Share)

labels_020 <- last_by_group(df_020, Year, Field) %>%
  mutate(
    .label = paste0(Field, ": ", sprintf("%.1f%%", Share))
  ) %>%
  prepare_right_labels(df_020$Year)

FIG207 <- ggplot(df_020, aes(Year, Share, color = Field, group = Field)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_label_repel_tpa(
    data    = labels_020,
    mapping = aes(x = .label_x, label = .label),
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_020$Year, right_labels = labels_020) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(0, 80),
                     breaks = seq(0, 80, 10),
                     expand = EXPAND_BAR_TIGHT) +
  scale_color_tpa() +
  guides(color = "none") +
  labs_meta(meta, y = "International share of postdocs") +
  coord_cartesian(clip = "off") +
  theme_tpa() +
  theme_right_labels(labels_020)

save_chart(FIG207, "fig_3_02")
```

### Figure 3.03 - Who Holds Postdoctoral Positions at Federally Funded Research Centers
```{r figure 021}
meta <- get_meta("FIG303")

df_021 <- read_fig("FIG303") %>%
  rename(Year = 1,
         `U.S. citizens and permanent residents` = 2,
         `Temporary visa holders`                = 3) %>%
  mutate(across(everything(), as.numeric)) %>%
  pivot_longer(-Year, names_to = "Status", values_to = "Count") %>%
  filter(!is.na(Year), !is.na(Count)) %>%
  order_by_latest(Status, Count)

labels_021 <- df_021 %>%
  nudge_by_series(Status,
                  above  = "Temporary visa holders",
                  amount = 120)

FIG208 <- ggplot(df_021, aes(Year, Count, color = Status, group = Status)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_point(size = 2) +
  geom_label_repel_tpa(
    data         = labels_021,
    mapping      = aes(label = label_comma()(Count)),
    nudge_y      = labels_021$nudge_y,
    direction    = "y",
    with_segment = FALSE
  ) +
  scale_x_years(df_021$Year, pad_right = PAD_RIGHT_TIGHT) +
  scale_y_continuous(labels = label_comma(),
                     limits = c(0, 2500),
                     breaks = seq(0, 2500, 500),
                     expand = EXPAND_BAR_TIGHT) +
  scale_color_manual(values = citizenship_colors_alt) +
  labs_meta(meta, y = "Postdoctoral researchers") +
  theme_tpa()

save_chart(FIG208, "fig_3_03")
```

### Figure 3.04 - Who Actually Works in STEM After Earning a STEM Degree?
```{r}
# shared prep -- combine US-born (FIG301a) and foreign-born (FIG301b)
# into one long frame with a Population column
prep_conversion <- function() {
  read_one <- function(name, pop) {
    read_fig(name) %>%
      mutate(across(-Year, as.numeric)) %>%
      pivot_longer(-Year, names_to = "Degree", values_to = "Rate") %>%
      mutate(Rate = Rate * 100, Population = pop)
  }

  bind_rows(
    read_one("FIG304a", "U.S.-born"),
    read_one("FIG304b", "Foreign-born")
  ) %>%
    mutate(
      Degree = recode(Degree,
        "bachelors"    = "Bachelors",
        "masters"      = "Masters",
        "doctorates"   = "Doctorates",
        "doctors"      = "Doctorates",
        "professional" = "Professional"),
      Degree = factor(Degree, levels = c("Doctorates", "Professional", "Masters", "Bachelors"))
    )
}

meta <- get_meta("FIG304")

first_yr <- 2003
last_yr  <- 2023

# long form: one row per Degree × Population × Year, with Year as a factor
# so it can drive the open/filled shape legend
df_db <- prep_conversion() %>%
  filter(Year %in% c(first_yr, last_yr)) %>%
  mutate(Year = factor(Year, levels = c(first_yr, last_yr)))

# wide, just for the connecting segments
df_seg <- df_db %>%
  select(Degree, Population, Year, Rate) %>%
  pivot_wider(names_from = Year, values_from = Rate, names_prefix = "Y")

fig_db <- ggplot(df_db, aes(y = Degree)) +
  geom_segment(data = df_seg,
               aes(x = Y2003, xend = Y2023, yend = Degree, color = Population),
               linewidth = 1.1, alpha = 0.55) +
  # both years drawn as shape 21; Year drives fill via the shape legend
  geom_point(aes(x = Rate, color = Population, shape = Year),
             fill = "white", stroke = 1.2, size = 3.8) +
  geom_point(data = filter(df_db, Year == last_yr),
             aes(x = Rate, color = Population, fill = Population),
             shape = 21, stroke = 0, size = 4) +
  scale_x_continuous(labels = label_number(suffix = "%"),
                     limits = c(20, 85), breaks = seq(20, 80, 10),
                     expand = expansion(mult = c(0.05, 0.10))) +
  scale_color_manual(values = c("U.S.-born"    = tpa_colors[2],
                                "Foreign-born" = tpa_colors[1])) +
  scale_fill_manual(values = c("U.S.-born"    = tpa_colors[2],
                               "Foreign-born" = tpa_colors[1]),
                    guide = "none") +
  scale_shape_manual(values = c("2003" = 21, "2023" = 19),
                     labels = c("2003" = "2003 (open)", "2023" = "2023 (filled)")) +
  guides(
    color = guide_legend(order = 1, override.aes = list(shape = 19, size = 4)),
    shape = guide_legend(order = 2,
                         override.aes = list(color = "grey40", fill = "white"))
  ) +
  labs_meta(meta, y = NULL) +
  theme_horizontal_bar()

save_chart(fig_db, "fig_3_04")
```


### Figure 3.05 - Top Employers Recieving H-1B Approvals
```{r figure 030}
# ---- shared FIG302 prep ----
# Full firm-level H-1B approvals (NAICS 54). Computes shares against
# the annual Grand Total row, excludes the partial 2026 column.
fig305_shares <- function() {
  raw <- read_fig("FIG305")
  names(raw) <- trimws(names(raw))

  total_row <- raw %>% filter(trimws(Company) == "Grand Total")
  totals <- total_row %>%
    select(-Company, -`Grand Total`) %>%
    pivot_longer(everything(), names_to = "Year", values_to = "Total") %>%
    mutate(Year = as.integer(Year), Total = as.numeric(Total))

  raw %>%
    filter(!trimws(Company) %in% c("Grand Total")) %>%
    mutate(Company = trimws(Company)) %>%
    select(-`Grand Total`) %>%
    pivot_longer(-Company, names_to = "Year", values_to = "Count") %>%
    mutate(Year = as.integer(Year),
           Count = as.numeric(replace_na(Count, 0))) %>%
    filter(Year != 2026) %>%                     # drop partial year
    left_join(totals, by = "Year") %>%
    mutate(Share = Count / Total * 100)
}

meta <- get_meta("FIG305")

short_names <- c(
  "COGNIZANT TECHNOLOGY SOLUTIONS US CORP"      = "Cognizant",
  "TATA CONSULTANCY SERVICES LIMITED"           = "Tata",
  "INFOSYS LIMITED"                             = "Infosys",
  "GOOGLE INC"                                  = "Google",
  "Deloitte and Subsidiaries"                   = "Deloitte",
  "WIPRO LIMITED"                               = "WIPRO",
  "ERNST & YOUNG LLP"                           = "Ernst & Young",
  "TECH MAHINDRA AMERICAS INC"                  = "Tech Mahindra",
  "HCL Technologies Limited and Subsidiaries"   = "HCL",
  "CAPGEMINI AMERICA INC"                       = "Capgemini",
  "LTIMINDTREE LIMITED"                         = "LTIMindtree",
  "IBM and subsidiaries"                        = "IBM",
  "LARSEN & TOUBRO INFOTECH LIMITED"            = "Larsen & Toubro"
)

shares <- fig305_shares()

# rank union firms by their max share across the two years, so the
# biggest bars (Cognizant's 2017, Google's 2025) sit on the left
df_bar <- shares %>%
  filter(Year %in% c(2017, 2025), Company %in% names(short_names)) %>%
  mutate(Company = recode(Company, !!!short_names),
         Year = factor(Year))

bar_order <- df_bar %>%
  group_by(Company) %>% summarise(m = max(Share)) %>%
  arrange(desc(m)) %>% pull(Company)

df_bar <- df_bar %>% mutate(Company = factor(Company, levels = bar_order))

FIG302a <- ggplot(df_bar, aes(Company, Share, fill = Year)) +
  geom_col(position = position_dodge(width = BAR_WIDTH),
           width = BAR_WIDTH, alpha = ALPHA_COL) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(0, 25), breaks = seq(0, 25, 5),
                     expand = EXPAND_BAR) +
  scale_fill_manual(values = c("2017" = tpa_colors[2], "2025" = tpa_colors[1])) +
  labs_meta(meta, y = "Share of H-1B approvals (NAICS 54)") +
  theme_tpa() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

save_chart(FIG302a, "fig_3_05")
```

### Figure 3.06 - Top Employers Receiving H-1B Approvals, Share Over Time
```{r fig_116}
meta <- get_meta("FIG306")

# true top-10-by-year share: rank all firms each year, sum the top 10
df_116 <- fig305_shares() %>%
  group_by(Year) %>%
  slice_max(Count, n = 10) %>%
  summarise(value = sum(Share), .groups = "drop") %>%
  mutate(Year = as.numeric(Year))

labels <- first_last_peak(df_116, Year, value) %>%
  mutate(hj = dplyr::case_when(
    Year == min(df_116$Year) ~ -0.4,   # first point: label to the RIGHT (inward)
    Year == max(df_116$Year) ~  1.3,   # last point:  label to the LEFT (inward)
    TRUE                       ~  0.5    # peak: stays centered
  ))

fig_116 <- ggplot(df_116, aes(Year, value)) +
  geom_line(color = tpa_colors[1], linewidth = LINE_WIDTH) +
  geom_point(data = labels, color = tpa_colors[1], size = POINT_SIZE) +
  # peak label: house wrapper, centered above the point
  geom_text_tpa(data    = filter(labels, hj == .5),
                mapping = aes(label = sprintf("%.1f%%", value)),
                color   = tpa_colors[1],
                vjust   = 1.5) +
  # endpoint labels: nudged inward so they don't clip the panel edges
  geom_text(data    = filter(labels, hj != 0.5),
            mapping = aes(label = sprintf("%.1f%%", value), hjust = hj),
            color   = tpa_colors[1], fontface = "bold",
            family  = FONT_FAMILY, size = DATA_LABEL_SIZE / .pt) +
  scale_x_years(df_116$Year) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(50, 75),
                     breaks = seq(50, 75, 5),
                     expand = EXPAND_BAR_TALL) +
  labs_meta(meta, y = "Top 10 employers' share of approvals") +
  theme_tpa()

save_chart(fig_116, "fig_3_06")
```


### Figure 3.07 - Who Founded Leading U.S. AI Companies
```{r figure 031}
meta <- get_meta("FIG307")

df_031 <- read_fig("FIG307") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Composition", values_to = "Share") %>%
  filter(!is.na(Share)) %>%
  mutate(
    Share       = Share * 100,
    Composition = case_when(
      Composition == "all us"       ~ "All U.S. Born",
      Composition == "all foreign"  ~ "All Foreign Born",
      Composition == "some foreign" ~ "Some Foreign Born"
    )
  ) %>%
  order_by_latest(Composition, Share)

palette_031 <- c(
  "All U.S. Born"     = tpa_colors[1],
  "All Foreign Born"  = tpa_colors[2],
  "Some Foreign Born" = tpa_colors[4]
)

labels_031 <- last_by_group(df_031, Year, Composition) %>%
  mutate(
    .label = paste0(Composition, ": ", fmt_pct_1(Share))
  ) %>%
  prepare_right_labels(df_031$Year)

FIG304 <- ggplot(df_031, aes(Year, Share, color = Composition, group = Composition)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_point(size = POINT_SIZE) +
  geom_label_repel_tpa(
    data    = labels_031,
    mapping = aes(x = .label_x, label = .label),
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_031$Year, right_labels = labels_031) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(20, 45),
                     breaks = seq(0, 45, 5),
                     expand = EXPAND_BAR_TIGHT) +
  scale_color_manual(values = palette_031) +
  guides(color = "none") +
  labs_meta(meta, y = "Share of Forbes AI 50") +
  coord_cartesian(clip = "off") +
  theme_tpa() +
  theme_right_labels(labels_031)

save_chart(FIG304, "fig_3_07")
```

### Figure 3.08 - Where Immigrant AI Founders Come From
```{r figure 032}
meta <- get_meta("FIG308")

df_032 <- read_fig("FIG308") %>%
  rename(Country = 1, Count = 2) %>%
  mutate(
    Count = as.numeric(Count),
    Country = case_when(
      Country == "the Netherlands"     ~ "Netherlands",
      Country == "United Arab Emirates" ~ "United Arab Emirates",
      Country == "Scotland"             ~ "United Kingdom",
      Country == "Hong Kong"            ~ "Hong Kong S.A.R.",
      TRUE                              ~ Country
    )
  ) %>%
  group_by(Country) %>%
  summarise(Count = sum(Count, na.rm = TRUE), .groups = "drop")

world_032 <- join_country_data(df_032)

FIG305 <- ggplot(world_032) +
  geom_sf(aes(fill = Count), color = "white", linewidth = 0.2) +
  geom_country_labels(world_032, value_col = "Count", min_value = 20, color = "black") +
  scale_fill_gradient(
    low = TPA_RED_LIGHT,
    high = tpa_colors[1],
    na.value = "grey92",
    breaks = c(1, 20, 40, 60, 76),
    labels = label_comma(),
    name = "Founders"
  ) +
  coord_sf(crs = "+proj=robin", expand = FALSE, datum = NA,
           xlim = c(-16000000, 16000000),
           ylim = c(-6000000, 8500000)) +
  labs_meta(meta) +
  theme_map() 
save_chart(FIG305, "fig_3_08")
```

### Figure 3.09 - Which Leading AI Companies Were Founded by Immigrants?
```{r figure 033}
### Figure 33 - Founder Composition and Valuation of Forbes AI50 Startups
fig033_data <- function() {
  raw <- read_fig("FIG309")
  names(raw) <- make.unique(names(raw))   # de-dupe the repeated headers
  foreign <- raw %>%
    transmute(Company,
              Valuation = as.numeric(`Pitchbook Valuation`),
              Origin = "Foreign-founded") %>%
    filter(!is.na(Valuation))
  domestic <- raw %>%
    transmute(Company,
              Valuation = as.numeric(`Pitchbook Valuation.1`),
              Origin = "Domestic-founded") %>%
    filter(!is.na(Valuation))
  bind_rows(foreign, domestic) %>%
    mutate(ValB = Valuation / 1e9)
}

meta <- get_meta("FIG309")

df_top <- fig033_data() %>%
  slice_max(ValB, n = 25) %>%
  mutate(Company = fct_reorder(Company, ValB))

FIG306 <- ggplot(df_top, aes(ValB, Company, color = Origin)) +
  geom_point(size = 3.2) +
  scale_x_log10(labels = label_dollar(suffix = "B", accuracy = 1),
                breaks = c(1, 3, 10, 30, 100, 300),
                expand = expansion(mult = c(0.04, 0.08))) +
  scale_color_manual(values = c("Foreign-founded"  = "#C46C5D",
                                "Domestic-founded" = "#A6C5DE")) +
  labs_meta(meta, y = NULL) +
  labs(x = "Valuation (log scale)") +
  theme_dotplot() +
  theme(
    panel.grid.major.y = element_line(color = "grey92", linewidth = 0.3),
    # theme_dotplot blanks the x-axis (value is normally on each dot);
    # this chart has no dot labels, so restore the axis and title it
    axis.text.x  = element_text(family = FONT_FAMILY, color = "grey25",
                                size = AXIS_SIZE, angle = 0, hjust = 0.5,
                                margin = margin(t = PUB$spacing$axis_text_gap_pt)),
    axis.ticks.x = element_line(color = "grey50", linewidth = 0.35),
    axis.line.x  = element_line(color = "grey50", linewidth = 0.4),
    axis.title.x = element_text(family = FONT_FAMILY, color = "grey25",
                                size = AXIS_SIZE, margin = margin(t = PUB$spacing$axis_title_gap_pt))
  )

save_chart(FIG306, "fig_3_09")
```




### Figure 3.10 - Where Researchers at Leading AI Companies Are Located?
```{r figure 034}
meta <- get_meta("FIG310")

company_order <- c("Alphabet","Anthropic", "OpenAI", "Alibaba", "DeepSeek", "Moonshot")
category_order <- c("Rest of World", "China", "United States")

df_034 <- read_fig("FIG310") %>%
  rename(Company = 1) %>%
  pivot_longer(-Company, names_to = "Affiliation", values_to = "Share") %>%
  mutate(
    Share = as.numeric(Share) * 100,
    Affiliation = str_remove(Affiliation, "^Current Affiliation - ?"),
    Affiliation = factor(Affiliation, levels = category_order),
    Company = factor(Company, levels = company_order)
  )

palette_034 <- c(
  "United States"      = tpa_colors[2],
  "China"              = tpa_colors[1],
  "Rest of World"      = tpa_colors[3]
)

FIG308 <- ggplot(df_034, aes(Share, Company, fill = Affiliation)) +
  geom_col(width = BAR_WIDTH_NARROW, alpha = ALPHA_COL) +
  geom_text(aes(label = fmt_pct_0(Share)),
            position = position_stack(vjust = 0.5),
            color = "white",
            fontface = "bold",
            family = FONT_FAMILY,
            size = DATA_LABEL_SIZE / .pt) +
  scale_x_continuous(labels = label_number(suffix = "%"),
                     breaks = seq(0, 100, 10),
                     limits = c(0, 101),
                     expand = expansion(mult = c(0, 0))) +
  scale_fill_manual(values = palette_034,
                    breaks = c("United States", "China", "Rest of World")) +
  labs_meta(meta, x = "Share of researchers") +
  theme_horizontal_bar()

save_chart(FIG308, "fig_3_10")
```

# Chapter 4

### Figure 4.01 - How Do American Students Score on International Test?
```{r figure 035}
meta <- get_meta("FIG401")

subject_order <- c("Reading", "Science", "Math")   # Math last -> drawn on top

df_035 <- read_fig("FIG401") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Subject", values_to = "Score") %>%
  filter(!is.na(Score)) %>%
  mutate(Subject = factor(Subject, levels = subject_order))

palette_035 <- c(
  "Math"    = tpa_colors[1],   # red accent -- the chapter's argument
  "Science" = "grey62",
  "Reading" = "grey78"
)

df_math  <- filter(df_035, Subject == "Math")
df_muted <- filter(df_035, Subject != "Subject" & Subject != "Math")

# point-level labels: every Math point, muted series only at endpoints
math_labels  <- df_math
muted_labels <- df_muted %>% filter(Year %in% range(Year))

# right-edge series labels replacing the legend
end_labels <- last_by_group(df_035, Year, Subject) %>%
  mutate(.label = as.character(Subject)) %>%
  prepare_right_labels(df_035$Year)

FIG401 <- ggplot(df_035, aes(Year, Score, color = Subject, group = Subject)) +
  geom_hline(yintercept = 0, color = "grey50", linewidth = LINE_WIDTH_REF) +
  geom_line(data = df_muted, linewidth = LINE_WIDTH) +
  geom_point(data = df_muted, size = 1.8) +
  geom_line(data = df_math, linewidth = LINE_WIDTH + 0.4) +
  geom_point(data = df_math, size = 2.4) +
  geom_text(data = math_labels, aes(label = sprintf("%+d", Score)),
            vjust = 1.9, size = DATA_LABEL_SIZE / .pt,
            fontface = "bold", family = FONT_FAMILY, show.legend = FALSE) +
  geom_text(data = muted_labels, aes(label = sprintf("%+d", Score)),
            vjust = -1.1, size = DATA_LABEL_SIZE / .pt,
            fontface = "bold", family = FONT_FAMILY, show.legend = FALSE) +
  geom_label_repel_tpa(
    data    = end_labels,
    mapping = aes(x = .label_x, label = .label),
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  annotate("text", x = min(df_035$Year), y = 1.6, label = "OECD average",
           hjust = 0, vjust = 0, color = "grey50",
           size = SMALL_LABEL_SIZE / .pt, family = FONT_FAMILY) +
  scale_x_years(df_035$Year, pad_left = 0.03, right_labels = end_labels) +
  scale_y_continuous(limits = c(-20, 32), breaks = seq(-20, 30, 5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  scale_color_manual(values = palette_035) +
  guides(color = "none") +
  labs_meta(meta, y = "Points above or below OECD average") +
  coord_cartesian(clip = "off") +
  theme_tpa() +
  theme_right_labels(end_labels) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5))

save_chart(FIG401, "fig_4_01")
```

### Figure 4.02 - How Do OECD Countries Rank in Math?
```{r FIG402}
meta <- get_meta("FIG402")
num  <- function(x) parse_number(as.character(x))

OECD_MATH_MEAN <- 472  # OECD 2022 mathematics average

oecd_members <- c(
  "Australia","Austria","Belgium","Canada","Chile","Colombia","Costa Rica",
  "Czech Republic","Denmark","Estonia","Finland","France","Germany","Greece",
  "Hungary","Iceland","Ireland","Israel","Italy","Japan","Korea","Lat\via",
  "Lithuania","Mexico","Netherlands","New Zealand","Norway","Poland","Portugal",
  "Slovak Republic","Slovenia","Spain","Sweden","Switzerland","Türkiye",
  "United Kingdom","United States"
)

df_120 <- read_fig("FIG402") %>%
  filter(Country %in% oecd_members) %>%
  mutate(Gap = num(PISA_Math_Mean) - OECD_MATH_MEAN,
         Country = fct_reorder(Country, Gap))

palette_120 <- highlight_colors(
  all_values  = levels(df_120$Country),
  highlighted = "United States",
  palette     = tpa_colors[1],   # red
  default     = tpa_colors[2]    # navy
)

FIG402 <- ggplot(df_120, aes(Gap, Country, fill = Country)) +
  geom_vline(xintercept = 0, color = "grey50", linewidth = LINE_WIDTH_REF) +
  geom_col(width = 0.75, alpha = ALPHA_COL) +
  geom_text(aes(label = if_else(Gap == 0, "0", sprintf("%+.0f", Gap)),
                hjust = if_else(Gap >= 0, -0.2, 1.2)),
            family = FONT_FAMILY,
            fontface = "bold",
            size = SMALL_LABEL_SIZE / .pt,
            color = "grey25") +
  scale_x_continuous(limits = c(-100, 75),
                     breaks = seq(-100, 75, 25),
                     expand = EXPAND_NONE) +
  scale_fill_manual(values = palette_120) +
  guides(fill = "none") +
  labs_meta(meta, x = "Points above or below the OECD average") +
  theme_horizontal_bar() +
  theme(axis.text.y  = element_text(size = SMALL_LABEL_SIZE),
        axis.title.x = element_text(family = FONT_FAMILY,
                                    size = SMALL_LABEL_SIZE,
                                    color = "grey25",
                                    margin = margin(t = PUB$spacing$axis_title_compact_pt)))
save_chart(FIG402, "fig_4_02", size = "medium")
```

### Figure 4.03 - How Do OECD Countries Rank in Math Proficiency?
```{r}
meta <- get_meta("FIG403")
num  <- function(x) parse_number(as.character(x))

keep_rows <- c(
  "Australia","Austria","Belgium","Canada","Chile","Colombia","Costa Rica",
  "Czech Republic","Denmark","Estonia","Finland","France","Germany","Greece",
  "Hungary","Iceland","Ireland","Israel","Italy","Japan","Korea","Latvia",
  "Lithuania","Mexico","Netherlands","New Zealand","Norway","Poland","Portugal",
  "Slovak Republic","Slovenia","Spain","Sweden","Switzerland","Türkiye",
  "United Kingdom","United States","OECD average"
)

df_122 <- read_fig("FIG403") %>%
  rename(Country = 1) %>%
  mutate(Country = str_remove(Country, "\\*+$"),      # drop sampling asterisks
         `Level 5` = num(`Level 5`),
         `Level 6` = num(`Level 6`),
         Top = `Level 5` + `Level 6`) %>%
  filter(Country %in% keep_rows) %>%
  mutate(Country = fct_reorder(Country, Top))

palette_122 <- highlight_colors(
  all_values  = levels(df_122$Country),
  highlighted = c("United States", "OECD average"),
  palette     = c(tpa_colors[1], tpa_colors[3]),   # red, gold
  default     = tpa_colors[2]                       # navy
)

FIG403 <- ggplot(df_122, aes(Top, Country, fill = Country)) +
  geom_col(width = 0.75, alpha = ALPHA_COL) +
  geom_text(aes(label = sprintf("%.1f", Top), hjust = -0.2),
            family = FONT_FAMILY,
            fontface = "bold",
            size = SMALL_LABEL_SIZE / .pt,
            color = "grey25") +
  scale_x_continuous(limits = c(0, 25),
                     breaks = seq(0, 25, 5),
                     expand = EXPAND_NONE) +
  scale_fill_manual(values = palette_122) +
  guides(fill = "none") +
  labs_meta(meta, x = "Share of students scoring at Level 5 or 6 (%)") +
  theme_horizontal_bar() +
  theme(axis.text.y  = element_text(size = SMALL_LABEL_SIZE),
        axis.title.x = element_text(family = FONT_FAMILY,
                                    size = SMALL_LABEL_SIZE,
                                    color = "grey25",
                                    margin = margin(t = PUB$spacing$axis_title_compact_pt)))
save_chart(FIG403, "fig_4_03", size = "medium")
```


### Figure 4.04 - Does Spending More on Schools Buy Better Results?
```{r fig_119}
meta <- get_meta("FIG404")

num <- function(x) parse_number(as.character(x))

df_119 <- read_fig("FIG404") %>%
  mutate(
    K12  = num(`K-12 (in USD)`),
    PISA = as.numeric(`OECD Math Average`)
  ) %>%
  filter(!is.na(K12), !is.na(PISA),
         !Country %in% c("EU25 average", "G20 average", "China")) %>%
  mutate(pt_group = case_when(
    Group == "United States" ~ "United States",
    Group == "OECD average"  ~ "OECD average",
    TRUE                     ~ "Other"
  ))

oecd_x <- df_119$K12[df_119$Country == "OECD average"]
oecd_y <- df_119$PISA[df_119$Country == "OECD average"]

# manually placed anchors (US, OECD average) vs. repelled grey labels
key_countries  <- c("United States", "OECD average")
label_grey <- df_119 %>%
  filter(Country %in% c("Korea", "Japan", "Peru", "Luxembourg",
                        "Türkiye", "Estonia", "Poland", "Latvia"))
label_key  <- df_119 %>%
  filter(Country %in% key_countries) %>%
  mutate(
    nx = if_else(Country == "United States",  1200, -1600),
    ny = if_else(Country == "United States",    -6,     9)
  )

pal_c <- c("United States" = tpa_colors[1],
           "OECD average"  = tpa_colors[3],
           "Other"         = "grey80")

fig_119 <- ggplot(df_119, aes(K12, PISA)) +
  geom_hline(yintercept = oecd_y, color = "grey70",
             linetype = "dashed", linewidth = LINE_WIDTH_REF) +
  geom_vline(xintercept = oecd_x, color = "grey70",
             linetype = "dashed", linewidth = LINE_WIDTH_REF) +
  geom_point(aes(color = pt_group, size = pt_group), alpha = ALPHA_BUBBLE) +
  # grey country labels: let repel place them
  geom_label_repel_tpa(
    data          = label_grey,
    mapping       = aes(label = Country, color = pt_group),
    seed          = 42,
    force         = 3,
    with_segment  = TRUE,
    direction     = "both"
  ) +
  # US and OECD average: placed by hand
  geom_label_repel_tpa(
    data         = label_key,
    mapping      = aes(label = Country, color = pt_group),
    nudge_x      = label_key$nx,
    nudge_y      = label_key$ny,
    seed         = 42,
    force        = 0,
    with_segment = TRUE,
    direction    = "both"
  ) +
  annotate("text", x = max(df_119$K12) * 0.98, y = oecd_y - 6,
           label = "OECD average score", hjust = 1, vjust = 1,
           color = "grey55", size = MICRO_LABEL_SIZE / .pt,
           family = FONT_FAMILY) +
  annotate("text", x = oecd_x - 350, y = min(df_119$PISA) + 4,
           label = "OECD average spending", angle = 90,
           hjust = 0, vjust = 1, color = "grey55",
           size = MICRO_LABEL_SIZE / .pt, family = FONT_FAMILY) +
  scale_x_continuous(labels = label_dollar(scale = 1e-3, suffix = "K"),
                     expand = expansion(mult = c(0.04, 0.08))) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.08))) +
  scale_color_manual(values = pal_c) +
  scale_size_manual(values = c("United States" = 4, "OECD average" = 3.5, "Other" = 2.4)) +
  guides(color = "none", size = "none") +
  coord_cartesian(clip = "off") +
  labs_meta(meta,
            x = "K-12 spending per student (USD, PPP)",
            y = "PISA math score, 2022") +
  theme_tpa() +
  theme(
    axis.text.y = element_text(size = SMALL_LABEL_SIZE),
    axis.title.x = element_text(family = FONT_FAMILY, color = "grey25",
                             size = AXIS_SIZE, margin = margin(t = PUB$spacing$axis_title_gap_pt))
  )

save_chart(fig_119, "fig_4_04", size = "standard")
```

### Figure 4.05 - How Lopsided Is U.S. Education Spending?
```{r fig_118}
meta <- get_meta("FIG405")

num <- function(x) parse_number(as.character(x))

df_118 <- read_fig("FIG405") %>%
  mutate(
    K12   = num(`K-12 (in USD)`),
    Tert  = num(`Tertiary (in USD)`),
    Ratio = Tert / K12
  ) %>%
  filter(!Country %in% c("EU25 average", "G20 average", "China")) %>%
  mutate(Country = fct_reorder(Country, Ratio))

pal_b <- highlight_colors(
  levels(df_118$Country),
  highlighted = c("United States", "OECD average"),
  palette     = c(tpa_colors[1], tpa_colors[3]),
  default     = tpa_colors[2]
)

fig_118 <- ggplot(df_118, aes(Ratio, Country, fill = Country)) +
  geom_vline(xintercept = 1, color = "grey50", linewidth = LINE_WIDTH_REF) +
  geom_col(width = 0.72, alpha = ALPHA_COL) +
  geom_text(aes(label = sprintf("%.1f\u00d7", Ratio),
                hjust = if_else(Ratio >= 1, -0.2, 1.2)),
            family = FONT_FAMILY, fontface = "bold",
            size = SMALL_LABEL_SIZE / .pt, color = "grey25") +
  annotate("text", x = 1.02, y = 25, label = "spends equally on both levels",
           angle = 90, hjust = 0, vjust = 0, color = "white",
           size = MICRO_LABEL_SIZE / .pt, family = FONT_FAMILY) +
  scale_x_continuous(breaks = seq(0, 3, 0.5),
                     labels = function(x) paste0(x, "\u00d7"),
                     expand = expansion(mult = c(0, 0.08))) +
  scale_fill_manual(values = pal_b) +
  guides(fill = "none") +
  coord_cartesian(clip = "off") +
  labs_meta(meta, x = "Tertiary-to-K-12 spending ratio (per student)") +
  theme_horizontal_bar() +
  theme(
    axis.text.y = element_text(size = SMALL_LABEL_SIZE),
    axis.title.x = element_text(family = FONT_FAMILY, color = "grey25",
                             size = AXIS_SIZE, margin = margin(t = PUB$spacing$axis_title_gap_pt))
  )

save_chart(fig_118, "fig_4_05", size = "medium")
```


### Figure 4.06 - STEM's Losses are Lateral: More Switch Fields Than Drop Out
```{r figure 038b}
meta <- get_meta("FIG406")

short_fields <- c(
  "Computer & information sciences"   = "Computer science",
  "Engineering"                       = "Engineering",
  "Bio/physical sciences, math & ag"  = "Sciences & math",
  "Non-STEM field"                    = "Non-STEM"
)

field_order <- c("Computer science", "Engineering", "Sciences & math")
col_order   <- c(field_order, "Non-STEM")

df_038b <- read_fig("FIG406") %>%
  mutate(pct = as.numeric(pct),
         is_diag      = as.character(entry_field) == as.character(degree_field),
         entry_field  = recode(entry_field, !!!short_fields),
         degree_field = recode(degree_field, !!!short_fields),
         entry_field  = factor(entry_field, levels = rev(field_order)),
         degree_field = factor(degree_field, levels = col_order),
         cohort       = factor(cohort, levels = c("2003/04", "2011/12")),
         lab = case_when(
           is.na(pct)                        ~ "n/a",
           reliability %in% c("unstable_30_50", "unstable_50plus") ~ paste0(fmt_pct_1(pct), "*"),
           TRUE                              ~ fmt_pct_1(pct)))

FIG406 <- ggplot(df_038b, aes(degree_field, entry_field, fill = pct)) +
  geom_tile(color = "white", linewidth = 1.5) +
  geom_tile(data = ~filter(.x, is_diag),
            color = "grey15", linewidth = 1, fill = NA) +
  geom_text(aes(label = lab, color = pct > 40),
            family = FONT_FAMILY, size = COMPACT_LABEL_SIZE / .pt) +
  scale_fill_gradient(low = "#FCEBEB", high = tpa_colors[1],  # or "#791F1F"
                      na.value = "grey92", guide = "none") +
  scale_color_manual(values = c(`TRUE` = "#FCEBEB", `FALSE` = "#501313"),
                     na.value = "grey45", guide = "none") +
  facet_wrap(~cohort) +
  labs_meta(meta, x = "Graduated in", y = "Entered") +
  theme_tpa() +
  theme(axis.title.x = element_text(family = FONT_FAMILY, color = "grey25",
                                    size = AXIS_SIZE, margin = margin(t = PUB$spacing$axis_title_gap_pt)),
        axis.title.y = element_text(family = FONT_FAMILY, color = "grey25",
                                    size = AXIS_SIZE, margin = margin(r = PUB$spacing$axis_label_right_gap_pt)),
        axis.text.x  = element_text(angle = 30, hjust = 1),
        panel.grid   = element_blank())

save_chart(FIG406, "fig_4_06")
```

### Figure 4.07 - Computing Holds its Majors While Broad Science Loses Half of Its Graduates
```{r figure 038c}
meta <- get_meta("FIG407")

short_fields <- c(
  "Computer & information sciences"   = "Computer science",
  "Engineering"                       = "Engineering",
  "Bio/physical sciences, math & ag"  = "Sciences & math",
  "Non-STEM field"                    = "Non-STEM"
)

field_order <- c("Computer science", "Engineering", "Sciences & math")

df_038c <- read_fig("FIG407") %>%
  mutate(across(-field, as.numeric)) %>%
  pivot_longer(-field, names_to = "cohort", values_to = "pct") %>%
  mutate(cohort = recode(cohort,
                         entry_2003 = "2003/04 cohort",
                         entry_2011 = "2011/12 cohort"),
         field  = recode(field, !!!short_fields),
         field  = factor(field, levels = field_order))

FIG407 <- ggplot(df_038c, aes(x = field, y = pct, fill = cohort)) +
  geom_col(width = 0.6, position = position_dodge(width = 0.65)) +
  geom_text(aes(label = fmt_pct_1(pct)),
            position = position_dodge(width = 0.65), vjust = -0.4,
            family = FONT_FAMILY, size = COMPACT_LABEL_SIZE / .pt) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(0, 80), breaks = seq(0, 80, 20),
                     expand = EXPAND_BAR_TIGHT) +
  scale_fill_tpa() +
  labs_meta(meta, y = "Graduated in entry field") +
  theme_tpa() +
  theme(axis.title.y = element_text(family = FONT_FAMILY, color = "grey25",
                                    size = AXIS_SIZE, margin = margin(r = PUB$spacing$axis_label_right_gap_pt)))

save_chart(FIG407, "fig_4_07")
```

### Figure 4.08 - When Do Non-Completers Leave College
```{r figure 083}
meta <- get_meta("FIG408")

year_levels <- c(
  "No degree, left without return year 1" = "Year 1",
  "No degree, left without return year 2" = "Year 2",
  "No degree, left without return year 3" = "Year 3",
  "No degree, left without return year 4" = "Year 4",
  "No degree, left without return year 5" = "Year 5",
  "No degree, left without return year 6" = "Year 6"
)

df_083 <- read_fig("FIG408") %>%
  rename(Cohort = 1) %>%
  select(Cohort, all_of(names(year_levels))) %>%   # timing only; still-enrolled/attained live in fig_084 + text
  mutate(across(-Cohort, as.numeric)) %>%
  pivot_longer(-Cohort, names_to = "ExitYear", values_to = "Share") %>%
  mutate(ExitYear = recode(ExitYear, !!!year_levels),
         Cohort   = factor(Cohort, levels = c("1996/2001", "2004/2009", "2012/2017")))

FIG408 <- ggplot(df_083, aes(ExitYear, Share, fill = Cohort)) +
  geom_col(width = 0.7, position = position_dodge(width = 0.78)) +
  geom_text(aes(label = fmt_pct_1(Share)),
            position = position_dodge(width = 0.78), vjust = -0.4,
            family = FONT_FAMILY, size = SMALL_LABEL_SIZE / .pt) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(0, 9), breaks = seq(0, 8, 2),
                     expand = EXPAND_BAR_TIGHT) +
  scale_fill_tpa() +
  labs_meta(meta, y = "Share of STEM entrants") +
  theme_tpa() +
  theme(axis.title.y = element_text(family = FONT_FAMILY, color = "grey25",
                                    size = AXIS_SIZE, margin = margin(r = PUB$spacing$axis_label_right_gap_pt)))

save_chart(FIG408, "fig_4_08")
```


### Figure 4.09 - How Many Students Study Abroad From Each Country### Figure 4.09 - How Many Students Study Abroad From Each Country
```{r FIG410}
meta <- get_meta("FIG409")
num  <- function(x) parse_number(as.character(x))   # parse_number chokes on numeric cols

# Wide CSV: one column per year (1997–2023); first column is labelled
# "Year" but holds country names. Pivot long, coerce, keep latest year.
mobility <- read_fig("FIG409") %>%
  rename(country = 1) %>%
  select(-any_of("Grand Total")) %>%
  pivot_longer(-country, names_to = "year", values_to = "ratio") %>%
  mutate(year = as.integer(year), ratio = num(ratio)) %>%
  filter(!is.na(ratio))

# OECD members present in the data (Korea = "South Korea") plus the two
# rivals. Luxembourg excluded: ratio > 100% (no full domestic system).
oecd_plus <- c(
  "Australia", "Austria", "Belgium", "Canada", "Chile", "Colombia",
  "Costa Rica", "Czechia", "Denmark", "Estonia", "Finland", "France",
  "Germany", "Greece", "Hungary", "Iceland", "Ireland", "Israel", "Italy",
  "Japan", "South Korea", "Latvia", "Lithuania", "Mexico", "Netherlands",
  "New Zealand", "Norway", "Poland", "Portugal", "Slovakia", "Slovenia",
  "Spain", "Sweden", "Switzerland", "Turkey", "United Kingdom",
  "United States", "China", "India"
)

df_124 <- mobility %>%
  filter(country %in% oecd_plus) %>%
  group_by(country) %>%
  slice_max(year, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(
    Country = recode(country, "South Korea" = "Korea"),
    grp = case_when(
      country == "United States"      ~ "United States",
      country %in% c("China", "India") ~ "Rival",
      TRUE                             ~ "OECD"
    )
  ) %>%
  arrange(ratio) %>%
  mutate(Country = fct_inorder(Country))   # US (lowest) at the floor

# House highlight pattern, mirroring FIG402/122: US in red, the rivals
# picked out in gold, all other OECD in navy.
palette_124 <- highlight_colors(
  all_values  = df_124$grp,
  highlighted = c("United States", "Rival"),
  palette     = c(tpa_colors[1], tpa_colors[3]),
  default     = tpa_colors[2]
)

FIG410 <- ggplot(df_124, aes(ratio, Country, fill = grp)) +
  geom_col(width = 0.75, alpha = ALPHA_COL) +
  geom_text(aes(label = sprintf("%.1f", ratio), hjust = -0.2),
            family = FONT_FAMILY, fontface = "bold",
            size = SMALL_LABEL_SIZE / .pt, color = "grey25") +
  scale_x_continuous(labels = label_number(suffix = "%"),
                     expand = EXPAND_LABEL_RIGHT) +
  scale_fill_manual(values = palette_124) +
  guides(fill = "none") +
  coord_cartesian(clip = "off") +
  labs_meta(meta, x = "Share of tertiary students studying abroad") +
  theme_horizontal_bar() +
  theme(axis.text.y  = element_text(size = SMALL_LABEL_SIZE),
        axis.title.x = element_text(family = FONT_FAMILY, color = "grey25",
                                    size = SMALL_LABEL_SIZE, margin = margin(t = PUB$spacing$axis_title_compact_pt)))

save_chart(FIG410, "fig_4_09", size = "medium")
```

### Figure 4.10 - Who Imports Students
```{r figure 042}
meta <- get_meta("FIG410")

df_042 <- read_fig("FIG410") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Country", values_to = "Flow") %>%
  mutate(Flow = Flow / 1e6) %>%              # keep NA rows so lines break at gaps
  order_by_latest(Country, Flow)

highlighted <- c("United States of America", "China")

df_042 <- df_042 %>% mutate(is_hi = Country %in% highlighted)

labels_042 <- df_042 %>%
  filter(!is.na(Flow), Country %in% highlighted) %>%
  group_by(Country) %>% slice_max(Year, n = 1) %>% ungroup() %>%
  mutate(
    .label = paste0(
      if_else(Country == "United States of America", "United States", Country),
      ": ",
      sprintf("%.2fM", Flow)
    )
  ) %>%
  prepare_right_labels(df_042$Year)

FIG411 <- ggplot() +
  geom_hline(yintercept = 0, color = "grey50", linewidth = LINE_WIDTH_REF) +
  geom_line(data = filter(df_042, !is_hi),
            aes(Year, Flow, group = Country),
            color = "grey80", linewidth = 0.35, na.rm = FALSE) +
  geom_line(data = filter(df_042, is_hi),
            aes(Year, Flow, color = Country, group = Country),
            linewidth = LINE_WIDTH_DENSE, na.rm = FALSE) +
  geom_label_repel_tpa(
    data = labels_042,
    aes(x = .label_x, y = Flow, label = .label, color = Country),
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_042$Year, right_labels = labels_042) +
  scale_y_continuous(labels = function(x) paste0(sprintf("%.1f", x), "M"),
                     limits = c(-1, 1),
                     breaks = seq(-1, 1, 0.5),
                     expand = expansion(mult = c(0.05, 0.05))) +
  scale_color_manual(values = c("United States of America" = tpa_colors[2],
                                "China" = tpa_colors[1])) +
  guides(color = "none") +
  labs_meta(meta, y = "Net flow (millions of students)") +
  coord_cartesian(clip = "off") +
  theme_tpa() +
  theme_right_labels(labels_042)

save_chart(FIG411, "fig_4_10")
```



# Chapter 5

### Figure 5.01 - Share of Authors of NeurIPS Oral Presentations by Primary Country of Affiliation (2013-2025)
```{r figure 047}
meta <- get_meta("FIG501")

df_047 <- read_fig("FIG501") %>%
  filter(!is.na(Year)) %>%
  mutate(across(everything(), as.numeric)) %>%
  pivot_longer(-Year, names_to = "Region", values_to = "Share") %>%
  filter(!is.na(Share)) %>%
  mutate(
    Share  = Share * 100,
    Region = factor(Region, levels = c("United States", "Rest of the World", "China"))
  )

palette_047 <- c(
  china_us_colors,
  "Rest of the World" = "grey55"
)

labels_047 <- last_by_group(df_047, Year, Region) %>%
  mutate(
    .label = paste0(Region, ": ", fmt_pct_1(Share))
  ) %>%
  prepare_right_labels(df_047$Year)

fig_047 <- ggplot(df_047, aes(Year, Share, color = Region, group = Region)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_point(size = 2) +
  geom_label_repel_tpa(
    data    = labels_047,
    mapping = aes(x = .label_x, label = .label),
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_047$Year, right_labels = labels_047) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(0, 90),
                     breaks = seq(0, 90, 10),
                     expand = EXPAND_BAR_TIGHT) +
  scale_color_manual(values = palette_047) +
  guides(color = "none") +
  labs_meta(meta, y = "Share of NeurIPS oral presentations") +
  coord_cartesian(clip = "off") +
  theme_tpa() +
  theme_right_labels(labels_047) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5))

save_chart(fig_047, "fig_5_01")
```

### Figure 5.02 - Who Produces the Most-Cited Research in Emerging Technologies?
```{r figure 044}
meta <- get_meta("FIG502")

df_044 <- read_fig("FIG502") %>%
  filter(!is.na(Year)) %>%
  mutate(across(everything(), as.numeric)) %>%
  pivot_longer(-Year, names_to = "Country", values_to = "Share") %>%
  mutate(Share   = Share * 100,
         Country = case_when(Country == "china" ~ "China",
                             Country == "us"    ~ "United States")) %>%
  group_by(Year) %>%
  mutate(label_y = Share + ifelse(Share == max(Share), 1.3, -1.3)) %>%
  ungroup()

FIG501 <- ggplot(df_044, aes(Year, Share, color = Country, group = Country)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_point(size = 2) +
  geom_text(aes(y = label_y, label = fmt_pct_0(Share)),
            size = DATA_LABEL_SIZE / .pt, fontface = "bold",
            family = FONT_FAMILY, show.legend = FALSE) +
  scale_x_years(df_044$Year) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(0, 25), breaks = seq(0, 25, 5),
                     expand = EXPAND_LINE) +
  scale_color_china_us() +
  labs_meta(meta, y = "Fractional share of top-cited publications") +
  coord_cartesian(xlim = c(2014.5, 2025.5), clip = "off") +
  theme_tpa() +
  theme_x_horizontal() +
  theme(plot.margin = publication_margin(
    right = PUB$spacing$edge_pt,
    left  = PUB$spacing$edge_pt
  ))

save_chart(FIG501, "fig_5_02")
```

### Figure 5.03 - Whose Authors Appear on the World's Top Papers?
```{r figure 045}
meta <- get_meta("FIG503")

df_045 <- read_fig("FIG503") %>%
  filter(!is.na(Year)) %>%
  mutate(across(everything(), as.numeric)) %>%
  pivot_longer(-Year, names_to = "Country", values_to = "Share") %>%
  mutate(Share   = Share * 100,
         Country = case_when(Country == "pct_papers_with_us" ~ "United States",
                             Country == "pct_papers_with_cn" ~ "China")) %>%
  group_by(Year) %>%
  mutate(label_y = Share + ifelse(Share == max(Share), 0.9, -0.9)) %>%
  ungroup()

FIG502 <- ggplot(df_045, aes(Year, Share, color = Country, group = Country)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_point(size = 2) +
  geom_text(aes(y = label_y, label = fmt_pct_1(Share)),
            size = DATA_LABEL_SIZE / .pt, fontface = "bold",
            family = FONT_FAMILY, show.legend = FALSE) +
  scale_x_years(df_045$Year) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(15, 35), breaks = seq(15, 35, 2),
                     expand = EXPAND_LINE) +
  scale_color_china_us() +
  labs_meta(meta, y = "Share of top papers with at least one author") +
  coord_cartesian(xlim = c(2014.5, 2025.5), clip = "off") +
  theme_tpa() +
  theme_x_horizontal() +
  theme(plot.margin = publication_margin(
    right = PUB$spacing$edge_pt,
    left  = PUB$spacing$edge_pt
  ))

save_chart(FIG502, "fig_5_03")
```

### Figure 5.04 - Where Do the World's Most-Cited Researchers Work
```{r figure 046}
meta <- get_meta("FIG504")

df_046 <- read_fig("FIG504") %>%
  filter(!is.na(Year)) %>%
  mutate(across(everything(), as.numeric)) %>%
  transmute(Year,
            China         = CN_Share * 100,
            `United States` = US_Share * 100) %>%
  pivot_longer(-Year, names_to = "Country", values_to = "Share")

labels_046 <- last_by_group(df_046, Year, Country) %>%
  mutate(
    .label = paste0(Country, ": ", fmt_pct_0(Share))
  ) %>%
  prepare_right_labels(df_046$Year)

FIG503 <- ggplot(df_046, aes(Year, Share, color = Country, group = Country)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_point(size = POINT_SIZE) +
  geom_label_repel_tpa(
    data    = labels_046,
    mapping = aes(x = .label_x, label = .label),
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_046$Year, right_labels = labels_046) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(0, 60),
                     breaks = seq(0, 60, 15),
                     expand = EXPAND_LINE) +
  scale_color_china_us() +
  guides(color = "none") +
  labs_meta(meta, y = "Share of global top 1% most-cited researchers") +
  coord_cartesian(clip = "off") +
  theme_tpa() +
  theme_right_labels(labels_046) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5))

save_chart(FIG503, "fig_5_04")
```


### Figure 5.05 - Who Spends the Most on R&D? 
```{r figure 063}
meta <- get_meta("FIG505")
df   <- read_country_lines("FIG505", group_col = "Region", value_col = "Value") %>%
  mutate(Value = Value / 1000)

fig <- plot_country_lines(df, meta, "GERD (billions of constant 2020 PPP dollars)",
                          group_col   = "Region",
                          value_col   = "Value",
                          label_fmt   = function(x) paste0("$", sprintf("%.0f", x), "B"),
                          highlight   = c("United States", "China"),
                          color_scale = scale_color_china_us(),
                          y_args      = list(
                            labels = function(x) paste0("$", x, "B"),
                            limits = c(0, 875),
                            breaks = seq(0, 875, 125),
                            expand = EXPAND_BAR_TIGHT
                          ))

save_chart(fig, "fig_5_05")
```

### Figure 5.06 - How Intensely Does Each Country Invest in R&D?
```{r figure 064}
meta <- get_meta("FIG506")
df   <- read_country_lines("FIG506", group_col = "Country", value_col = "Share") %>%
  filter(Country != "Russia") %>%
  mutate(Country = if_else(Country == "US", "United States", Country),
         Share   = Share * 100)

fig <- plot_country_lines(df, meta, "R&D intensity (% of GDP)",
                          value_col   = "Share",
                          label_fmt   = fmt_pct_1,
                          highlight   = c("United States", "China"),
                          color_scale = scale_color_china_us(),
                          y_args      = list(
                            labels = label_number(suffix = "%"),
                            limits = c(0, 6),
                            breaks = seq(0, 6, 1),
                            expand = EXPAND_BAR_TIGHT
                          ))

save_chart(fig, "fig_5_06")
```


### Figure 5.07 - Is R&D Spending Aimed at Today or Tomorrow?
```{r FIG509}
meta <- get_meta("FIG507")

rd_type_colors <- c(
  "Basic research"           = tpa_colors[1],  
  "Applied research"         = tpa_colors[2],  
  "Experimental development" = tpa_colors[3]  
)

df <- read_fig("FIG507") %>%
  rename(Year = 1, Country = 2, Type = 3, value = 4) %>%
  mutate(
    Year    = as.integer(round(as.numeric(as.character(Year)))),
    value   = as.numeric(value),
    Country = factor(Country, levels = c("United States", "China")),
    Type    = recode(as.character(Type),
                     "Basic Research"           = "Basic research",
                     "Applied Research"          = "Applied research",
                     "Experimental Development"  = "Experimental development"),
    Type    = factor(Type, levels = c("Basic research", "Applied research",
                                      "Experimental development"))
  )

fig <- ggplot(df, aes(Year, value, fill = Type)) +
  geom_area(position = position_stack(reverse = TRUE), alpha = ALPHA_AREA) +
  facet_wrap(~ Country) +
  scale_fill_manual(values = rd_type_colors, breaks = names(rd_type_colors)) +
  scale_x_years(df$Year, pad_left = 0.02, pad_right = 0.05) +
  scale_y_continuous(
    labels = function(x) paste0("$", x, "B"),
    limits = c(0, 875),
    breaks = seq(0, 875, 125),
    expand = EXPAND_BAR_TIGHT
  ) +
  labs_meta(meta) +
  theme_facet()
save_chart(fig, "fig_5_07")
```

### Figure 5.08 - Business Now Dominates R&D in Both Countries
```{r fig_126}
meta <- get_meta("FIG508")

sector_colors <- c(
  "Business enterprise" = tpa_colors[3],
  "Government"          = tpa_colors[2],
  "Higher education"    = tpa_colors[1],
  "Private non-profit"  = tpa_colors[4]
)

df <- read_fig("FIG508") %>%
  rename(Year = 1, Country = 2, Sector = 3, value = 4) %>%
  mutate(
    Year   = as.integer(round(as.numeric(as.character(Year)))),
    value  = as.numeric(value),
    Sector = recode(as.character(Sector),
                    "Business enterprise sector" = "Business enterprise",
                    "Government sector"          = "Government",
                    "Higher education sector"    = "Higher education",
                    "Private non-profit sector"  = "Private non-profit")
  ) %>%
  filter(Sector %in% names(sector_colors)) %>%
  complete(Country, Sector, Year, fill = list(value = 0)) %>%   # China private non-profit -> $0 all years
  mutate(
    Country = factor(Country, levels = c("United States", "China")),
    Sector  = factor(Sector, levels = names(sector_colors))
  )

fig_126 <- ggplot(df, aes(Year, value, fill = Sector)) +
  geom_area(position = position_stack(reverse = FALSE), alpha = ALPHA_AREA) +
  facet_wrap(~ Country) +
  scale_fill_manual(values = sector_colors, breaks = names(sector_colors)) +
  scale_x_years(df$Year, pad_left = 0.02, pad_right = 0.05) +
  scale_y_continuous(
    labels = function(x) paste0("$", x, "B"),
    limits = c(0, 875),
    breaks = seq(0, 875, 125),
    expand = EXPAND_BAR_TIGHT
  ) +
  labs_meta(meta) +
  theme_facet()

save_chart(fig_126, "fig_5_08")
```

# Chapter 6

### Figure 6.01 - Do International PhD Recipients Plan to Stay in the United States
```{r figure 065}
meta <- get_meta("FIG601")

df_065 <- read_fig("FIG601") %>%
  rename(Year = 1, Share = 2) %>%
  filter(!is.na(Year)) %>%
  mutate(across(everything(), as.numeric),
         Share = Share * 100)

FIG601 <- ggplot(df_065, aes(Year, Share)) +
  geom_col(fill = tpa_colors[2], width = 0.7, alpha = ALPHA_COL) +
  geom_text(aes(label = fmt_pct_0(Share)),
            vjust = -0.5,
            color = tpa_colors[2],
            size = DATA_LABEL_SIZE / .pt,
            fontface = "bold",
            family = FONT_FAMILY) +
  scale_x_years(df_065$Year) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(60, 80),
                     breaks = seq(60, 80, 2),
                     expand = EXPAND_TIGHT_TOP,
                     oob = scales::oob_keep) +
  coord_cartesian(ylim = c(60, 80)) +
  labs_meta(meta, y = "Intention to stay in the United States") +
  theme_tpa()

save_chart(FIG601, "fig_6_01")
```

### Figure 6.02 - Stay Rates of International STEM PhD Recipients
```{r figure 067}
meta <- get_meta("FIG602")
df_067 <- read_fig("FIG602") %>%
  filter(!is.na(Year)) %>%
  mutate(across(everything(), as.numeric)) %>%
  pivot_longer(-Year, names_to = "Rate", values_to = "Share") %>%
  filter(!is.na(Share)) %>%
  mutate(
    Share = Share * 100,
    Rate  = factor(Rate, levels = c("5-year stay rate", "10-year stay rate"))
  )

# first point, last point, and peak of each line; first-point labels
# left-align so they extend right off the marker instead of over the axis
labels_067 <- flp_by_group(df_067, Year, Share, Rate) %>%
  mutate(hjust = if_else(Year == min(Year), 1, 0.5))

palette_067 <- c(
  "5-year stay rate"  = tpa_colors[2],   # navy
  "10-year stay rate" = tpa_colors[1]    # red
)

FIG602 <- ggplot(df_067, aes(Year, Share, color = Rate, group = Rate)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_point(size = 2) +
  geom_text_tpa(
    data        = labels_067,
    mapping     = aes(label = fmt_pct_1(Share), hjust = hjust),
    show.legend = FALSE
  ) +
  scale_x_years(df_067$Year, pad_right = .09, pad_left = 0.09) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(50, 75),
                     breaks = seq(50, 75, 5),
                     expand = expansion(mult = c(0, 0.05))) +
  scale_color_manual(values = palette_067) +
  coord_cartesian(clip = "off") +
  labs_meta(meta, y = "Stay rate") +
  theme_tpa() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5))

save_chart(FIG602, "fig_6_02")
```

### Figure 6.03 - STEM OPT Applications and Approvals
```{r figure 071}
meta <- get_meta("FIG603")

opt_years <- c(2008, 2013, 2018, 2023, 2025)

df_071 <- read_fig("FIG603") %>%
  filter(!is.na(Year)) %>%
  mutate(across(everything(), as.numeric)) %>%
  filter(Year %in% opt_years) %>%
  select(Year, Received, Approved) %>%
  pivot_longer(-Year, names_to = "Status", values_to = "Count") %>%
  mutate(Status = factor(Status, levels = c("Received", "Approved")),
         Year   = factor(Year, levels = opt_years))

FIG603 <- ggplot(df_071, aes(Year, Count, fill = Status)) +
  geom_col(position = position_dodge(width = 0.8),
           width = 0.7, alpha = ALPHA_COL) +
  geom_text(aes(label = label_comma()(Count)),
            position = position_dodge(width = 0.8),
            vjust = -0.5,
            size = SMALL_LABEL_SIZE / .pt,
            fontface = "bold",
            family = FONT_FAMILY) +
  scale_x_discrete() +
  scale_y_continuous(labels = fmt_num_k,
                     expand = EXPAND_TIGHT_TOP) +
  scale_fill_manual(name = NULL,
                    values = c("Received" = tpa_colors[2],
                               "Approved" = tpa_colors[1])) +
  labs_meta(meta, y = "STEM OPT applications") +
  coord_cartesian(clip = "off") +
  theme_tpa()

save_chart(FIG603, "fig_6_03")
```

### Figure 6.04 - H-1B Approvals Over Time
```{r figure 087}
meta <- get_meta("FIG604")

total_approvals <- function(name) {
  read_fig(name) %>%
    filter(!is.na(Year)) %>%
    select(Year, `New Approved`, `Renewal Approved`) %>%
    mutate(across(everything(), as.numeric)) %>%
    transmute(Year, Total = `New Approved` + `Renewal Approved`)
}

df_072 <- total_approvals("FIG604a") %>% rename(All = Total) %>%
  left_join(total_approvals("FIG604b") %>% rename(Tech = Total), by = "Year")

FIG604a <- ggplot(df_072, aes(Year)) +
  geom_area(aes(y = All),  fill = tpa_colors[2], alpha = ALPHA_AREA) +
  geom_area(aes(y = Tech), fill = tpa_colors[1], alpha = ALPHA_AREA) +
  geom_line(aes(y = All),  color = tpa_colors[2], linewidth = LINE_WIDTH) +
  geom_line(aes(y = Tech), color = tpa_colors[1], linewidth = LINE_WIDTH) +
  annotate("text", x = 2017, y = 150000, label = "All industries",
           color = "white", fontface = "bold", hjust = 0.5,
           family = FONT_FAMILY, size = DATA_LABEL_SIZE / .pt) +
  annotate("text", x = 2017, y = 50000, label = "Tech occupations\n(NAICS 54)",
           color = "white", fontface = "bold", hjust = 0.5, lineheight = 0.9,
           family = FONT_FAMILY, size = COMPACT_LABEL_SIZE / .pt) +
  scale_x_years(df_072$Year) +
  scale_y_continuous(labels = fmt_num_k, expand = EXPAND_BAR) +
  labs_meta(meta, y = "H-1B approvals") +
  theme_tpa()

save_chart(FIG604a, "fig_6_04")
```


### Figure 6.05 - H-1B Denials Over Time
```{r figure 087}
meta <- get_meta("FIG605")

df_087 <- read_hb_denials("FIG605a") %>% filter(Type == "New Denials") %>%
  select(Year, All = Count) %>%
  left_join(
    read_hb_denials("FIG605b") %>% filter(Type == "New Denials") %>%
      select(Year, Tech = Count),
    by = "Year"
  )

FIG605a <- ggplot(df_087, aes(Year)) +
  geom_area(aes(y = All),  fill = tpa_colors[2], alpha = ALPHA_AREA) +
  geom_area(aes(y = Tech), fill = tpa_colors[1], alpha = ALPHA_AREA) +
  geom_line(aes(y = All),  color = tpa_colors[2], linewidth = LINE_WIDTH) +
  geom_line(aes(y = Tech), color = tpa_colors[1], linewidth = LINE_WIDTH) +
  annotate("text", x = 2017, y = 32000, label = "All industries",
           color = tpa_colors[2], fontface = "bold", hjust = 0.5,
           family = FONT_FAMILY, size = DATA_LABEL_SIZE / .pt) +
  annotate("text", x = 2018.75, y = 14000, label = "Tech occupations\n(NAICS 54)",
           color = "white", fontface = "bold", hjust = 0.5, lineheight = 0.9,
           family = FONT_FAMILY, size = COMPACT_LABEL_SIZE / .pt) +
  scale_x_years(df_087$Year) +
  scale_y_continuous(labels = fmt_num_k, expand = EXPAND_BAR) +
  labs_meta(meta, y = "New H-1B denials") +
  theme_tpa()

save_chart(FIG605a, "fig_6_05")
```
### Figure 6.06 - Outcomes of PERM Applications
```{r figure 076}
meta <- get_meta("FIG606")

df_076 <- read_fig("FIG606") %>%
  filter(!is.na(Year)) %>%
  select(Year,
         `Used`    = Certified,
         `Expired` = `Certified-expired`,
         `Denied`  = Denied) %>%
  mutate(across(everything(), as.numeric)) %>%
  pivot_longer(-Year, names_to = "Status", values_to = "Count") %>%
  filter(!is.na(Count)) %>%
  mutate(Status = factor(Status, levels = c("Denied", "Expired", "Used")))

palette_076 <- c(
  "Used"    = tpa_colors[3],   # gold — certifications actually used
  "Expired" = tpa_colors[1],   # red  — certified but lapsed unused
  "Denied"  = tpa_colors[6]
)

FIG606 <- ggplot(df_076, aes(Year, Count, fill = Status)) +
  geom_area(alpha = ALPHA_AREA) +
  scale_x_years(df_076$Year) +
  scale_y_continuous(labels = fmt_num_k, expand = EXPAND_BAR) +
  scale_fill_manual(values = palette_076,
                    breaks = c("Used", "Expired", "Denied")) +
  labs_meta(meta, y = "PERM certifications") +
  theme_tpa()

save_chart(FIG606, "fig_6_06")
```


### Figure 6.07 - Top Countries of Origin for J-1 Participants
```{r figure 078}
meta <- get_meta("FIG607")

df_078 <- read_fig("FIG607") %>%
  filter(!is.na(Year)) %>%
  mutate(across(everything(), as.numeric)) %>%
  pivot_longer(-Year, names_to = "Country", values_to = "Share") %>%
  filter(!is.na(Share)) %>%
  mutate(Share    = Share * 100,
         is_china = Country == "China")

china_078 <- filter(df_078, is_china)
other_078 <- filter(df_078, !is_china)
china_last <- china_078 %>%
  filter(Year == max(Year)) %>%
  mutate(.label = paste0("China: ", fmt_pct_1(Share))) %>%
  prepare_right_labels(df_078$Year)

FIG607 <- ggplot() +
  geom_line(data = other_078,
            aes(Year, Share, group = Country),
            color = "grey75", linewidth = LINE_WIDTH) +
  geom_line(data = china_078,
            aes(Year, Share),
            color = tpa_colors[1], linewidth = LINE_WIDTH_DENSE) +
  geom_point(data = china_last,
             aes(Year, Share), color = tpa_colors[1], size = 2.5) +
  geom_label_repel_tpa(
    data    = china_last,
    mapping = aes(x = .label_x, y = Share, label = .label),
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  annotate("text", x = 2015, y = 11.6, label = "China", hjust = 0,
           color = tpa_colors[1], fontface = "bold",
           family = FONT_FAMILY, size = DATA_LABEL_SIZE / .pt) +
  annotate("text", x = 2020.2, y = 7.0, label = "Other top-five countries",
           hjust = 0, color = "grey55",
           family = FONT_FAMILY, size = SMALL_LABEL_SIZE / .pt) +
  scale_x_years(df_078$Year, right_labels = china_last) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(0, 12), breaks = seq(0, 12, 2),
                     expand = EXPAND_BAR_TIGHT) +
  labs_meta(meta, y = "Share of J-1 participants") +
  coord_cartesian(clip = "off") +
  theme_tpa() +
  theme_right_labels(china_last)

save_chart(FIG607, "fig_6_07")
```

### Figure 6.08 - Workforce Participation Rates of Immigrants versus Native Born Workers
```{r figure 079}
meta <- get_meta("FIG608")

df_079 <- read_fig("FIG608") %>%
  rename(Gap = `Gap with the native-born`) %>%
  mutate(Gap = as.numeric(Gap),
         Country = fct_reorder(Country, Gap))

palette_079 <- highlight_colors(
  all_values  = levels(df_079$Country),
  highlighted = c("United States", "OECD average"),
  palette     = c(tpa_colors[1], tpa_colors[3]),   # red, gold
  default     = tpa_colors[2]                       # navy
)

FIG608 <- ggplot(df_079, aes(Gap, Country, fill = Country)) +
  geom_vline(xintercept = 0, color = "grey50", linewidth = LINE_WIDTH_REF) +
  geom_col(width = 0.75, alpha = ALPHA_COL) +
  geom_text(aes(label = sprintf("%.1f", Gap),
                hjust = if_else(Gap >= 0, -0.2, 1.2)),
            family = FONT_FAMILY,
            fontface = "bold",
            size = SMALL_LABEL_SIZE / .pt,
            color = "grey25") +
  scale_x_continuous(limits = c(-15, 20),
                     breaks = seq(-15, 20, 5),
                     expand = EXPAND_NONE) +
  scale_fill_manual(values = palette_079) +
  guides(fill = "none") +
  labs_meta(meta, x = "Percentage point gap with the native-born") +
  theme_horizontal_bar() +
  theme(axis.text.y = element_text(size = SMALL_LABEL_SIZE))

save_chart(FIG608, "fig_6_08", size = "medium")
```

### Figure 6.09 - How Much STEM Talent Does the U.S. Train and then Lose? 
```{r FIG609}
# FIG609 (EVJ213) — chapter closer, by field.
# One cohort (2012-14 doctorate recipients), one horizon (~10-year U.S. stay
# rate), split into four STEM fields. Each field's departures valued at that
# field's own time to degree x the NSF GRFP support rate. Benchmark, not a
# per-person accounting (see note).

cohort <- read_fig("FIG609") %>%
  mutate(across(c(recipients, stay_rate_10yr, ttd_years), as.numeric))

COST_PER_YEAR <- 53000

fld <- cohort %>%
  mutate(
    stayed     = recipients * stay_rate_10yr,
    left       = recipients - stayed,
    left_share = left / recipients,
    cost_phd   = COST_PER_YEAR * ttd_years,
    value_left = left * cost_phd
  ) %>%
  arrange(left_share) %>%                       # lowest leak -> bottom, highest -> top
  mutate(field = factor(field, levels = field))

value_total <- sum(fld$value_left)              # ~3.94e9

# long form for stacking; Left leftmost so its right edge = departure count
seg <- fld %>%
  select(field, Stayed = stayed, Left = left) %>%
  pivot_longer(c(Stayed, Left), names_to = "outcome", values_to = "count") %>%
  mutate(outcome = factor(outcome, levels = c("Stayed", "Left")))

# leak label anchored just past the red segment's right edge (x = left)
leak_lab <- fld %>%
  transmute(field, left,
            llab = paste0(label_comma(accuracy = 100)(left),
                          "  (", round(left_share * 100), "%)"))

outcome_fill <- c("Stayed" = tpa_colors[9],   # muted
                  "Left"   = tpa_colors[1])   # emphasis red

meta <- get_meta("FIG609")
meta$caption <- paste0(meta$caption)

x_max <- max(fld$recipients)

p_127 <- ggplot(seg, aes(x = count, y = field, fill = outcome)) +
  geom_col(width = 0.68, alpha = ALPHA_COL) +
  geom_text(
    data = leak_lab, inherit.aes = FALSE,
    aes(x = left, y = field, label = llab),
    hjust = -0.08, family = FONT_FAMILY, fontface = "bold",
    size = DATA_LABEL_SIZE / .pt, color = tpa_colors[1]
  ) +
  annotate(
    "text", x = 0, y = 0.32, hjust = 0,
    label = sprintf("\u2248 $%.1f billion in STEM training investment left with them",
                    value_total / 1e9),
    family = FONT_FAMILY, fontface = "bold",
    size = TITLE_SIZE / .pt, color = tpa_colors[1]
  ) +
  scale_x_continuous(labels = label_comma(), limits = c(0, x_max),
                     expand = expansion(mult = c(0, 0.05))) +
  scale_y_discrete(expand = expansion(add = c(0.95, 0.6))) +
  scale_fill_manual(values = outcome_fill, breaks = c("Stayed", "Left")) +
  labs_meta(meta, x = NULL, y = NULL) +
  theme_horizontal_bar()

save_chart(p_127, "fig_6_09", size = "standard")
```


