title: "GTCI Tables"
author: "Emerson Johnston"
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE
)

# Resolve this Rmd's directory, then locate setup_publication.R from there.
rmd_dir <- tryCatch(
  dirname(normalizePath(knitr::current_input(dir = TRUE), winslash = "/", mustWork = TRUE)),
  error = function(e) normalizePath(getwd(), winslash = "/", mustWork = TRUE)
)

setup_candidates <- c(
  file.path(rmd_dir, "inputs", "scripts", "setup_publication.R"),
  file.path(rmd_dir, "..", "inputs", "scripts", "setup_publication.R"),
  file.path(getwd(), "inputs", "scripts", "setup_publication.R")
)

setup_path <- setup_candidates[file.exists(setup_candidates)][1]

if (is.na(setup_path)) {
  stop(
    "Could not locate inputs/scripts/setup_publication.R from the Rmd location or working directory.",
    call. = FALSE
  )
}

source(setup_path)

# Start a fresh table-export audit for this knit. save_table() appends one row
# per exported table and the final audit chunk verifies width + typography.
if (exists("reset_table_export_audit")) {
  reset_table_export_audit()
}
```

# Chapter 1 

### Table 1.01 - How Much of U.S. STEM Training Goes to International Students?
```{r table 006}
meta_d <- get_meta("TAB101")

change_type <- "ppt"  # choose "relative" or "ppt"

TAB101 <- build_change_table_wide(
  "TAB101",
  meta_d,
  change_type = change_type
)

save_table(
  TAB101,
  "tab_1_01",
  profile = "ultrawide",
  stub_width_px = 350
)
```


# Chapter 2 

### Table 2.01 - Where International Students in the United States Come From
```{r table 007}
meta_007 <- get_meta("TAB201")
TAB201 <- build_region_table(
  name = "TAB201",
  meta = meta_007
)
save_table(TAB201, "tab_2_01", profile = "ultrawide", stub_width_px = 260)
```


### Table 2.02 Which Fields Are Each Countries' Students Leaving? 
```{r}
meta <- get_meta("TAB202")
change_type <- "ppt"   # this table only makes sense as percentage points

# The nine field files use two units. TAB202a/b/c are proportions (0.388);
# TAB202d–i are already percent numbers (11.2). Scale only a/b/c by 100 so
# every field lands in the same percent unit and a country's fields sum to ~100.
prop_files <- c("TAB202a", "TAB202b", "TAB202c")
field_files <- c(
  "Engineering"              = "TAB202a",
  "Math & Computer Science"  = "TAB202b",
  "Physical & Life Sciences" = "TAB202c",
  "Business and Management"  = "TAB202d",
  "Education"                = "TAB202e",
  "Fine/Applied Arts"        = "TAB202f",
  "Health Professions"       = "TAB202g",
  "Social Sciences"          = "TAB202h"
)

# Short column headers. The data columns are named by the long field labels
# above; these map each to a compact header whose longest single word fits one
# column at this width, so no header word overflows its cell. Field ordering,
# formatting, and color all key off the long names via field_order and are
# unaffected.
field_header_short <- c(
  "Engineering"              = "Engineering",
  "Math & Computer Science"  = "Math/Computer",
  "Physical & Life Sciences" = "Phys./Life Sci.",
  "Business and Management"  = "Business/Mgmt",
  "Education"                = "Education",
  "Fine/Applied Arts"        = "Fine/Applied Arts",
  "Health Professions"       = "Health Prof.",
  "Social Sciences"          = "Social Sci."
)

# Rows shown, in order. Sorted by total enrollment, largest first.
country_order <- c("China", "India", "South Korea", "Iran",
                   "Nigeria", "Nepal", "Pakistan", "Bangladesh","Saudi Arabia","Turkey/Türkiye","Taiwan","Spain","Mexico","Colombia","France","Indonesia","Ghana","Italy","Brazil","Vietnam","Hong Kong","Canada","Germany","Japan","United Kingdom","Kuwait","Russia","Thailand","Venezuela")

long_all <- imap_dfr(field_files, function(file, field) {
  read_fig(file) %>%
    mutate(across(-Year, ~ suppressWarnings(as.numeric(na_if(as.character(.x), "-"))))) %>%
    pivot_longer(-Year, names_to = "Country", values_to = "Share") %>%
    filter(!is.na(Share)) %>%
    mutate(
      Share = if (file %in% prop_files) Share * 100 else Share,
      Field = field
    )
})

# First and last available year per country/field, then the point change.
change_long <- long_all %>%
  filter(Country %in% country_order) %>%
  group_by(Country, Field) %>%
  filter(n() >= 2) %>%
  summarise(
    Change = table_change(Share[which.min(Year)], Share[which.max(Year)], change_type),
    .groups = "drop"
  )

# Order fields left-to-right by mean movement across countries, so fields
# gaining share sit at the left and fields being left behind at the right.
field_order <- change_long %>%
  group_by(Field) %>%
  summarise(.avg = mean(Change, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(.avg)) %>%
  pull(Field)

# Wide: countries in rows, fields in columns.
mat <- change_long %>%
  pivot_wider(names_from = Field, values_from = Change) %>%
  mutate(Country = factor(Country, levels = country_order)) %>%
  arrange(Country) %>%
  mutate(Country = as.character(Country)) %>%
  select(Country, all_of(field_order))

# One symmetric domain across the whole matrix so color is comparable cell-to-cell.
lim <- max(abs(as.matrix(mat[field_order])), na.rm = TRUE)

tab_202 <- mat %>%
  gt(rowname_col = "Country", id = "TAB202") %>%
  fmt_number(columns = all_of(field_order), decimals = 1,
             force_sign = TRUE, pattern = "{x}%") %>%
  sub_missing(missing_text = "\u2014") %>%
  data_color(
    columns = all_of(field_order),
    fn = scales::col_numeric(
      palette  = TABLE_CHANGE_PALETTE,   # blue (gain) → white (0) → red (loss)
      domain   = c(-lim, lim),
      na.color = "white"
    )
  ) %>%
  cols_align(align = "right", columns = all_of(field_order)) %>%
  cols_label(.list = {
    lbl <- field_header_short[field_order]
    lbl[is.na(lbl)] <- field_order[is.na(lbl)]   # fall back to full name if unmapped
    setNames(as.list(lbl), field_order)
  }) %>%
  style_stub_left() %>%
  tab_spanner(
    label   = "Change in share of the country's U.S.-bound students (pct. points)",
    columns = all_of(field_order)
  ) %>%
  tab_footnote(
    footnote = paste(
      "Each cell is the percentage-point change in the share of that country's",
      "U.S.-bound international students enrolled in the field, from the first",
      "to the most recent year available (2010\u20132025 for most countries;",
      "2012 for Iran, 2018 for Bangladesh). Red marks fields gaining share,",
      "blue marks fields losing it. Columns are ordered from the field gaining",
      "most across countries to the field losing most. Countries are ordered by",
      "total international enrollment, largest first."
    ),
    locations = cells_column_spanners(
      spanners = "Change in share of the country's U.S.-bound students (pct. points)"
    )
  ) %>%
  theme_gt_tpa(meta = meta) %>%
  opt_css(css = sprintf(
    "#%s .gt_column_spanner { border-bottom-color: %s !important; }",
    "TAB202", tpa_colors[1]
  )) %>%
  with_table_profile("ultrawide") %>%
  with_size("standard")

save_table(tab_202, "tab_2_02", profile = "ultrawide", stub_width_px = 150)
```



# Chapter 3 

### Table 3.01 - Which STEM Graduates Work in STEM Occupations
```{r tab_14}
meta <- get_meta("TAB301")   # or reuse TAB301's meta
change_type <- "ppt"  # choose "relative" or "ppt"

df_029t <- read_fig("TAB301") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Level", values_to = "Share") %>%
  filter(!is.na(Share), Year %in% c(1993, 2003, 2013, 2023)) %>%
  mutate(
    Share = Share * 100,
    Level = case_when(
      str_detect(Level, regex("bach", ignore_case = TRUE)) ~ "Bachelors",
      str_detect(Level, regex("mast", ignore_case = TRUE)) ~ "Masters",
      str_detect(Level, regex("doct", ignore_case = TRUE)) ~ "Doctorate"
    )
  ) %>%
  pivot_wider(names_from = Year, values_from = Share, names_prefix = "Y") %>%
  mutate(
    Chg2003 = table_change(Y1993, Y2003, change_type),
    Chg2013 = table_change(Y2003, Y2013, change_type),
    Chg2023 = table_change(Y2013, Y2023, change_type),
    Change  = table_change(Y1993, Y2023, change_type),
    Occupation = factor(Level, levels = c("Bachelors", "Masters", "Doctorate"))
  ) %>%
  arrange(Occupation) %>%
  select(Occupation, Y1993, Chg2003, Chg2013, Chg2023, Y2023, Change)

tab_014 <- build_occupation_share_table(
  df_029t, "tab_014", meta,
  level_cols    = c("Y1993", "Y2023"),
  chg_cols      = c("Chg2003", "Chg2013", "Chg2023"),
  spanner_label = "Share employed in STEM",
  change_type   = change_type,
  footnote = paste(
    "1993 is the share of graduates at each degree level employed in a STEM",
    "occupation. 2003, 2013, and 2023 show the",
    if (change_type == "relative") "relative change" else "percentage-point change",
    "from the prior period shown rather than the share itself. Net Change uses",
    "the same change type from 1993 to 2023."
  )
)

save_table(tab_014, "tab_3_01", profile = "wide", stub_width_px = 260)
```

### Table 3.02 - How Concentrated Are Temporary Visa Holders in STEM Occupations 
```{r}
meta <- get_meta("TAB302")
change_type <- "ppt"  # choose "relative" or "ppt"

short_names <- c(
  "Biological, agricultural, and other life scientists" = "Biological scientists",
  "Computer and mathematical scientists"                = "Computer/math scientists",
  "Physical and related scientists"                     = "Physical scientists",
  "Social and related scientists"                        = "Social scientists",
  "Engineers"                                            = "Engineers",
  "S&E-related occupations"                               = "S&E-related occupations",
  "Non-S&E occupations"                                   = "Non-S&E occupations"
)

row_order <- c("Computer/math scientists", "Engineers", "Biological scientists",
               "Physical scientists", "Social scientists",
               "S&E-related occupations", "Non-S&E occupations")

df_011 <- read_fig("TAB302") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Occupation", values_to = "Rate") %>%
  filter(!is.na(Rate), Year %in% c(2017, 2019, 2021, 2023)) %>%
  mutate(Rate = Rate * 100,
         Occupation = recode(Occupation, !!!short_names)) %>%
  pivot_wider(names_from = Year, values_from = Rate, names_prefix = "Y") %>%
  mutate(
    Chg2019 = table_change(Y2017, Y2019, change_type),
    Chg2021 = table_change(Y2019, Y2021, change_type),
    Chg2023 = table_change(Y2021, Y2023, change_type),
    Change  = table_change(Y2017, Y2023, change_type),
    Occupation = factor(Occupation, levels = row_order)
  ) %>%
  arrange(Occupation) %>%
  select(Occupation, Y2017, Chg2019, Chg2021, Chg2023, Y2023, Change)

TAB302 <- build_occupation_share_table(
  df_011, "TAB302", meta,
  level_cols  = c("Y2017", "Y2023"),
  chg_cols    = c("Chg2019", "Chg2021", "Chg2023"),
  change_type = change_type
)

save_table(TAB302, "tab_3_02", profile = "wide", stub_width_px = 300)
```


### Table 3.03 - Top Employers Receiving H-1B Approvals
```{r tab_012}
meta <- get_meta("TAB303")
change_type <- "ppt"  # choose "relative" or "ppt"

fig305_shares <- function() {
  raw <- read_fig("FIG305")
  names(raw) <- trimws(names(raw))

  total_row <- raw %>% filter(trimws(Company) == "Grand Total")
  totals <- total_row %>%
    select(-Company, -`Grand Total`) %>%
    pivot_longer(everything(), names_to = "Year", values_to = "Total") %>%
    mutate(Year = as.integer(Year), Total = as.numeric(Total))

  raw %>%
    filter(!trimws(Company) %in% c("Grand Total")) %>%
    mutate(Company = trimws(Company)) %>%
    select(-`Grand Total`) %>%
    pivot_longer(-Company, names_to = "Year", values_to = "Count") %>%
    mutate(Year = as.integer(Year),
           Count = as.numeric(replace_na(Count, 0))) %>%
    filter(Year != 2026) %>%                     # drop partial year
    left_join(totals, by = "Year") %>%
    mutate(Share = Count / Total * 100)
}

shares <- fig305_shares()

# proper-case helper (acronyms preserved, minor words lowercased)
proper_case <- function(x,
                        keep_upper = c("IBM", "HCL", "LLP", "LLC", "US", "USA", "EY"),
                        small = c("and", "of", "the", "for")) {
  vapply(x, function(s) {
    words <- str_split(s, "\\s+")[[1]]
    out <- vapply(seq_along(words), function(i) {
      w <- words[i]
      if (toupper(w) %in% keep_upper)             toupper(w)
      else if (w == "&")                          "&"
      else if (tolower(w) %in% small && i != 1)   tolower(w)
      else                                        str_to_title(w)
    }, character(1))
    paste(out, collapse = " ")
  }, character(1), USE.NAMES = FALSE)
}

# union of firms that were top-10 in EITHER year
union_firms <- shares %>%
  filter(Year %in% c(2017, 2025)) %>%
  group_by(Year) %>%
  slice_max(Count, n = 10) %>%
  ungroup() %>%
  distinct(Company) %>%
  pull(Company)

wide <- shares %>%
  filter(Year %in% c(2017, 2025), Company %in% union_firms) %>%
  select(Company, Year, Share) %>%
  pivot_wider(names_from = Year, values_from = Share, names_prefix = "Y")

df_012 <- wide %>%
  mutate(
    Company = proper_case(recode(Company, !!!short_names)),
    Y1 = Y2017, Y2 = Y2025,
    Change = if_else(Y2017 > 0 | change_type == "ppt",
                     table_change(Y2017, Y2025, change_type), NA_real_)
  ) %>%
  arrange(desc(Change)) %>%          # order by 2025 share (today's leaders first)
  select(Company, Y1, Y2, Change)

tab_012 <- build_h1b_top10_table(df_012, "tab_012", meta,
                                 first_year = 2017, last_year = 2025,
                                 change_type = change_type)

save_table(tab_012, "tab_3_03", profile = "standard", stub_width_px = 360)
```


# Chapter 4

### Table 4.01 - Four in Ten STEM Entrants Finish with a STEM Degree — and That Barely Moved
```{r tab_018}
meta <- get_meta("TAB401")   # or a dedicated tab_038a entry
change_type <- "ppt"  # choose "relative" or "ppt"

tab_038a <- build_bookend_table(
  name          = "TAB401",
  meta          = meta,
  first         = "2003/04",
  last          = "2011/12",
  id_col        = "cohort",
  spanner_label = "Share of STEM entrants",
  change_type   = change_type,
  recode_cats = c(
    stem_bachelors    = "Completed Bachelor's Degree in STEM Field",
    nonstem_bachelors = "Completed Bachelor's Degree in Another, Non-STEM Field",
    subbac_credential = "Completed A Subbac Credential",
    still_enrolled    = "Still Enrolled",
    left_no_degree    = "Left with No Degree"
  ),
  cat_order = c(
    "Completed Bachelor's Degree in STEM Field",
    "Completed Bachelor's Degree in Another, Non-STEM Field",
    "Completed A Subbac Credential",
    "Still Enrolled",
    "Left with No Degree"
  ),
  footnote = paste(
    "Values are the share of each entering STEM cohort ending in each outcome;",
    "each year column sums to 100%. Net change is the",
    if (change_type == "relative") "relative change" else "percentage-point change",
    "from the 2003/04 cohort to the 2011/12 cohort."
  )
)
save_table(tab_038a, "tab_4_01", profile = "standard", stub_width_px = 340)
```


### Table 4.02 - How Many Times Do STEM Majors Change Majors
```{r TAB402}
meta <- get_meta("TAB402")
change_type <- "ppt"  # choose "relative" or "ppt"
TAB402 <- build_bookend_table(
  "TAB402", meta,
  first = "2004/2009 Survey", last = "2012/2017 Survey",
  id_col = "Survey",
  first_label = "2004/2009", last_label = "2012/2017",
  spanner_label = "Share of STEM entrants",
  change_type = change_type,
  recode_cats = c("Never"          = "Never switched",
                  "Once"           = "Switched once",
                  "More than Once" = "Switched 2+ times"),
  cat_order = c("Never switched", "Switched once", "Switched 2+ times"),
  footnote = paste(
    "Share of STEM entrants by number of major switches. Net change is the",
    if (change_type == "relative") "relative change" else "percentage-point change",
    "from the 2004/2009 to the 2012/2017 cohort."
  )
)
save_table(TAB402, "tab_4_02", profile = "standard")
```

### Table 4.03 - What Do Americans Study Abroad?
```{r tab_016}
meta <- get_meta("TAB403")
change_type <- "ppt"  # choose "relative" or "ppt"
tab_016 <- build_bookend_table(
  "TAB403", meta,
  first = 2005, last = 2024,
  spanner_label = "Share of students",
  change_type = change_type,
  sort_by = "change",
  recode_cats = c("STEM Fields"           = "STEM",
                  "Business & Management" = "Business & management",
                  "Social Sciences"       = "Social sciences",
                  "Fine and Applied Arts" = "Fine & applied arts"),
  flagged_rows = c("Social sciences", "Humanities", "Other"),
  flag_note = paste(
    "Open Doors revised its field taxonomy between 2013 and 2014; shares for",
    "these categories are not directly comparable across that break, so the",
    "reported change reflects reclassification as well as any real shift."
  ),
  footnote = paste(
    "Share of U.S. study-abroad students by field of study.",
    "Net change is the",
    if (change_type == "relative") "relative change" else "percentage-point change",
    "from 2005 to 2024."
  )
)

save_table(tab_016, "tab_4_03", profile = "standard", stub_width_px = 260)
```

### Table 4.04 - Where Do American's Study Abroad
```{r tab_017}
meta <- get_meta("TAB404")
change_type <- "ppt"  # choose "relative" or "ppt"
tab_017 <- build_bookend_table(
  "TAB404", meta,
  first = 2005, last = 2024,
  spanner_label = "Share of students",
  change_type = change_type,
  sort_by = "change",
  drop_cols = "Antarctica",
  recode_cats = c("Africa, Sub-Saharan"   = "Sub-Saharan Africa",
                  "Multiple Destinations" = "Multiple destinations"),
  footnote = paste(
    "Share of U.S. study-abroad students by host region. Net change is the",
    if (change_type == "relative") "relative change" else "percentage-point change",
    "from 2005 to 2024.",
    if (change_type == "relative")
      "Relative change is volatile for small-share regions (e.g., the Middle East & North Africa moved 1.0% to 1.6%)."
    else
      ""
  )
)

save_table(tab_017, "tab_4_04", profile = "standard", stub_width_px = 280)
```


# Chapter 5 

### Table 5.01 - When Did China Catch At Each AI Conference
```{r table 019}
change_type <- "ppt"  # choose "relative" or "ppt"

conf_names <- c(
  acml         = "ACML (ML)",
  nips         = "NeurIPS (ML / AI)",
  icml         = "ICML (ML)",
  cvpr         = "CVPR (Vision)",
  iccv         = "ICCV (Vision)",
  iclr         = "ICLR (ML)",
  emnlp        = "EMNLP (NLP)",
  ijcai        = "IJCAI (AI)",
  corl         = "CoRL (Robotics)",
  rss          = "RSS (Robotics)",
  iros         = "IROS (Robotics)",
  coling       = "COLING (NLP)",
  icra         = "ICRA (Robotics)",
  uai          = "UAI (Prob. ML)",
  naacl        = "NAACL (NLP)",
  acl          = "ACL (NLP)",
  aistats      = "AISTATS (Stats)",
  colt         = "COLT (Theory)",
  siggraphasia = "SIGGRAPH Asia",
  wacv         = "WACV (Vision)"
)

tab_501a <- build_conference_summary_table_split(
  name       = "tab_019a",
  meta       = get_meta("TAB501"),
  conf_names = conf_names,
  data_name  = "TAB501",
  row_slice  = 1:20,
  change_type = change_type
)

save_table(tab_501a, "tab_5_01", profile = "ultrawide", stub_width_px = 155)
```

### Table 5.02 - How Many U.S. Universities Rank in the Global Top 100? 
```{r TAB502}
meta <- get_meta("TAB502")
change_type <- "ppt"  # choose "relative" or "ppt"
num  <- function(x) parse_number(as.character(x))

field_order <- c("Physical Sciences", "Life Sciences", "Engineering", "Computer Science")
years_shown <- c(2014, 2018, 2022, 2026)

raw <- read_fig("TAB502") %>%
  rename(Year = 1) %>%
  mutate(Year = as.integer(num(Year)))

cs_2017 <- raw %>% filter(Year == 2017) %>% pull(`Computer Science`) %>% num()

df_001 <- raw %>%
  select(Year, all_of(field_order)) %>%                 # drop precomputed Average
  filter(Year %in% years_shown) %>%
  pivot_longer(-Year, names_to = "Field", values_to = "value") %>%
  mutate(value = num(value)) %>%
  pivot_wider(names_from = Year, values_from = value, names_prefix = "Y") %>%
  mutate(
    # computer science was first ranked in 2017; drop its 2017 count into the
    # 2014 slot so its change and net columns chain consistently
    Y2014   = if_else(Field == "Computer Science" & is.na(Y2014), cs_2017, Y2014),
    Chg2018 = table_change(Y2014, Y2018, change_type),
    Chg2022 = table_change(Y2018, Y2022, change_type),
    Chg2026 = table_change(Y2022, Y2026, change_type),
    Change  = table_change(Y2014, Y2026, change_type),
    Field   = factor(Field, levels = field_order)
  ) %>%
  arrange(Field) %>%
  select(Field, Y2014, Chg2018, Chg2022, Chg2026, Y2026, Change)

TAB502 <- build_ranking_delta_table(
  df_001, "TAB502", meta,
  level_cols  = c("Y2014", "Y2026"),
  chg_cols    = c("Chg2018", "Chg2022", "Chg2026"),
  change_type = change_type,
  footnote = paste(
    "Counts are U.S. universities in the global top 100 in each field, in 2014 and",
    "2026. Interior columns show the",
    if (change_type == "relative") "relative change" else "absolute change",
    "in that count from the prior shown year; Net Change uses the same change type",
    "from 2014 to 2026."
  ),
  cell_footnote       = paste(
    "Computer science was first ranked in 2017; its 2017 count is shown here in the",
    "2014 column so its change and net columns can be computed."
  ),
  cell_footnote_field = "Computer Science",
  cell_footnote_col   = "Y2014"
)
save_table(TAB502, "tab_5_02", profile = "standard")
```

### Table 5.03 - Is R&D Spending Aimed at Today or Tomorrow? 
```{r table_020}
meta <- get_meta("TAB503")

type_labels <- c(
  "Basic research"                              = "Basic research (for tomorrow)",
  "Applied research"                            = "Applied research (for today)",
  "Experimental development"                    = "Experimental development (for today)",
  "Capital expenditure not elsewhere recorded"  = "Capital expenditure (unallocated)"
)
type_order <- c("Basic research (for tomorrow)",
                "Applied research (for today)",
                "Experimental development (for today)",
                "Capital expenditure (unallocated)",
                "Total")

df_125 <- read_fig("TAB503") %>%
  rename(Year = 1, Country = 2, Type = 3, value = 4) %>%
  mutate(Year  = as.numeric(as.character(Year)),
         value = as.numeric(value)) %>%
  filter(Year %in% c(1994, 2004, 2014, 2024)) %>%
  mutate(Type = recode(Type, !!!type_labels)) %>%
  pivot_wider(names_from = Year, values_from = value, names_prefix = "Y")

totals_125 <- df_125 %>%
  group_by(Country) %>%
  summarise(across(c(Y1994, Y2004, Y2014, Y2024), ~ sum(.x, na.rm = TRUE)),
            .groups = "drop") %>%
  mutate(Type = "Total")

df_125 <- bind_rows(df_125, totals_125) %>%
  mutate(
    Change  = table_change(Y1994, Y2024, "relative"),
    Country = factor(Country, levels = c("United States", "China")),
    Type    = factor(Type, levels = type_order)
  ) %>%
  arrange(Country, Type) %>%
  select(Country, Type, Y1994, Y2004, Y2014, Y2024, Change)

tab_021 <- build_rd_type_level_table(
  df_125, "tab_5_03", meta,
  level_cols = c("Y1994", "Y2004", "Y2014", "Y2024")
)

save_table(
  tab_021,
  "tab_5_03",
  profile = "ultrawide",
  stub_width_px = 420
)
```

### Table 5.04 - Business Now Dominates R&D in Both Countries 
```{r tab_022}
meta <- get_meta("TAB504")

sector_labels <- c(
  "Business enterprise sector" = "Business enterprise",
  "Government sector"          = "Government",
  "Higher education sector"    = "Higher education",
  "Private non-profit sector"  = "Private non-profit"
)
sector_order <- c("Business enterprise", "Government",
                  "Higher education", "Private non-profit", "Total")

df_022 <- read_fig("TAB504") %>%
  rename(Year = 1, Country = 2, Type = 3, value = 4) %>%
  mutate(Year  = as.numeric(as.character(Year)),
         value = as.numeric(value)) %>%
  filter(Type %in% names(sector_labels),
         Year %in% c(1994, 2004, 2014, 2024)) %>%
  mutate(Type = recode(Type, !!!sector_labels)) %>%
  complete(Country, Type, Year, fill = list(value = 0)) %>%   # China private non-profit -> $0
  pivot_wider(names_from = Year, values_from = value, names_prefix = "Y")

# per-country total row, summed across the four performing sectors
totals_022 <- df_022 %>%
  group_by(Country) %>%
  summarise(across(c(Y1994, Y2004, Y2014, Y2024), ~ sum(.x, na.rm = TRUE)),
            .groups = "drop") %>%
  mutate(Type = "Total")

df_022 <- bind_rows(df_022, totals_022) %>%
  mutate(
    Change  = table_change(Y1994, Y2024, "relative"),
    Change  = ifelse(is.finite(Change), Change, 0),   # 0/0 -> 0%
    Country = factor(Country, levels = c("United States", "China")),
    Type    = factor(Type, levels = sector_order)
  ) %>%
  arrange(Country, Type) %>%
  select(Country, Type, Y1994, Y2004, Y2014, Y2024, Change)

tab_022 <- build_rd_type_level_table(
  df_022, "tab_5_04", meta,
  level_cols = c("Y1994", "Y2004", "Y2014", "Y2024"),
  footnote = paste(
    "Spending levels in billions of US dollars, PPP converted, at current prices",
    "(not adjusted for inflation). Net Change is the relative change from the first",
    "to the last shown year. Sectors denote where R&D is performed. China reports no",
    "private non-profit R&D; it is shown as $0, and its net-change cell is undefined",
    "from a zero base and appears as 0%. Totals may not sum exactly because of rounding."
  )
)

save_table(
  tab_022,
  "tab_5_04",
  profile = "ultrawide",
  stub_width_px = 420
)
```

### Table 5.05 - Who Patents in Critical Technologies
```{r table 020}
change_type <- "relative"  # choose "relative" or "ppt"

build_patents_table(
  name    = "tab_020",     # output id; adjust to next free table number
  meta    = get_meta("TAB505"),
  sources = c(
    "Biotechnology"       = "TAB505a",
    "Semiconductors"      = "TAB505b",
    "Computer technology" = "TAB505c"
  ),
  change_type = change_type
) %>%
  save_table("tab_5_05", profile = "ultrawide")
```

### Table 5.06 - Patent Leadership Shifted From U.S. and Japanese Incumbents to Asian Challengers
```{r TAB506}
meta <- get_meta("TAB506")
change_type <- "ppt"
num  <- function(x) parse_number(as.character(x))

flag_slug <- c(
  "United States" = "united-states",
  "Japan"         = "japan",
  "Taiwan"        = "taiwan",
  "South Korea"   = "south-korea",
  "Germany"       = "germany",
  "China"         = "china"
)

df_023 <- read_fig("TAB506") %>%
  rename(Company = 1, Country = 2, Y2005 = 3, Y2015 = 4, Y2025 = 5) %>%
  mutate(across(c(Y2005, Y2015, Y2025), num)) %>%
  mutate(
    flag     = paste0("inputs/flags/", recode(Country, !!!flag_slug), ".png"),
    Change   = table_change(Y2005, Y2025, change_type),
    Rank2005 = dplyr::min_rank(dplyr::desc(Y2005)),   # 1 = most 2005 grants
    Rank2025 = dplyr::min_rank(dplyr::desc(Y2025))    # 1 = most 2025 grants
  )

# mark the five firms that were top 5 in 2005 with a colored rank arrow.
# arrow points down (fell in rank) or up (rose) between 2005 and 2025.
arrow_span <- function(rank2005, fell) {
  color <- if (fell) "#8C1515" else "#3B7A57"   # TPA red for decline, green for rise
  glyph <- if (fell) "\u25BE" else "\u25B4"      # ▾ / ▴
  paste0(" <span style='color:", color, ";font-weight:700;'>",
         glyph, " #", rank2005, "</span>")
}

df_023 <- df_023 %>%
  mutate(
    Company = ifelse(
      Rank2005 <= 5,
      paste0(Company, mapply(arrow_span, Rank2005, Rank2025 > Rank2005)),
      Company
    )
  ) %>%
  arrange(Rank2025) %>%                            # sort by 2025 rank, not net change
  select(Company, flag, Y2005, Y2025, Change)

TAB506 <- build_patent_company_delta_table(
  df_023, "TAB506", meta,
  level_cols  = c("Y2005", "Y2025"),
  chg_cols    = character(0),
  change_type = change_type,
  footnote = paste(
    "The first and last columns are patent grants in 2005 and 2025.",
    "Net Change is the absolute change across the window.",
    "Arrows mark the five firms that ranked in the top five by 2005 grants,",
    "with their 2005 rank; red points to a fall in rank by 2025, green to a rise.",
    "Rows are ordered by 2025 grants."
  )
)
save_table(TAB506, "tab_5_06", profile = "ultrawide", stub_width_px = 360)
```





# Chapter 6

### Table 6.01 - Do Stay Intentions Differ by Region of Origin? 
```{r table 028}
meta <- get_meta("TAB601")
change_type <- "ppt"  # choose "relative" or "ppt"

tab_028 <- build_bookend_table(
  "TAB601", meta,
  first = 2002, last = 2024,
  spanner_label = "Intend to stay",
  scale = 100,
  drop_cols = "Region and country or economy unknown",
  change_type = change_type
)

save_table(tab_028, "tab_6_01", profile = "standard")
```


### Table 6.02 - What New STEM Phds Plan to Do After Graduations
```{r table 025}
meta <- get_meta("TAB602")
change_type <- "ppt"  # choose "relative" or "ppt"

tab_025 <- build_bookend_table(
  "TAB602", meta,
  first = 2021, last = 2024,
  spanner_label = "Share with U.S. plans",
  sort_by = "change",
  scale = 100,
  change_type = change_type
)

save_table(tab_025, "tab_6_02", profile = "standard")
```

### Table 6.03 - Where U.S. Awarded PhDs Work After Graduation
```{r table 027}
meta <- get_meta("TAB603")
change_type <- "ppt"  # choose "relative" or "ppt"

TAB603 <- build_sector_bookend_table(meta, change_type = change_type)

save_table(TAB603, "tab_6_03", profile = "standard")
TAB603
```

### Table 6.04 - How Many PERM Certifications Go Unusued
```{r table 029}
meta <- get_meta("TAB604")
change_type <- "ppt"  # choose "relative" or "ppt"

TAB604 <- build_perm_expiry_table(meta, change_type = change_type)

save_table(TAB604, "tab_6_04", profile = "standard")
```

### Table 6.05 - Employment-Based Green Card Wait Times by Nationality
```{r table 026}
meta <- get_meta("TAB605")

TAB605 <- build_eb_combined_table(meta)

save_table(TAB605, "tab_6_05", profile = "standard")
TAB605
```


# Table publication audit
```{r table-publication-audit, include=FALSE}
EXPECTED_TABLE_EXPORTS <- c(
  "tab_1_01",
  "tab_2_01", "tab_2_02",
  "tab_3_01", "tab_3_02", "tab_3_03",
  "tab_4_01", "tab_4_02", "tab_4_03", "tab_4_04",
  "tab_5_01", "tab_5_02", "tab_5_03", "tab_5_04", "tab_5_05", "tab_5_06",
  "tab_6_01", "tab_6_02", "tab_6_03", "tab_6_04", "tab_6_05"
)

table_layout_audit <- write_table_export_audit(
  path = file.path(PATHS$final_tables, "table_layout_audit_ch1_ch6.csv"),
  expected_names = EXPECTED_TABLE_EXPORTS
)

# A second, compact console check makes the publication contract visible during
# knitting without adding anything to the rendered document.
message(
  "TABLE AUDIT PASS: ",
  nrow(table_layout_audit),
  " tables; common width = ",
  unique(table_layout_audit$expected_width_px),
  " px; body/column/spanner/title/source = ",
  paste(
    unique(table_layout_audit$body_font_pt),
    unique(table_layout_audit$column_label_font_pt),
    unique(table_layout_audit$spanner_font_pt),
    unique(table_layout_audit$title_font_pt),
    unique(table_layout_audit$source_font_pt),
    sep = "/"
  ),
  " pt; row padding = ",
  unique(table_layout_audit$row_padding_pt),
  " pt."
)
```
