---
title: "GTCI Tables"
author: "Emerson Johnston"
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE
)

# Resolve this Rmd's directory, then locate setup_publication.R from there.
rmd_dir <- tryCatch(
  dirname(normalizePath(knitr::current_input(dir = TRUE), winslash = "/", mustWork = TRUE)),
  error = function(e) normalizePath(getwd(), winslash = "/", mustWork = TRUE)
)

setup_candidates <- c(
  file.path(rmd_dir, "inputs", "scripts", "setup_publication.R"),
  file.path(rmd_dir, "..", "inputs", "scripts", "setup_publication.R"),
  file.path(getwd(), "inputs", "scripts", "setup_publication.R")
)

setup_path <- setup_candidates[file.exists(setup_candidates)][1]

if (is.na(setup_path)) {
  stop(
    "Could not locate inputs/scripts/setup_publication.R from the Rmd location or working directory.",
    call. = FALSE
  )
}

source(setup_path)

# Start a fresh table-export audit for this knit. save_table() appends one row
# per exported table and the final audit chunk verifies width + typography.
if (exists("reset_table_export_audit")) {
  reset_table_export_audit()
}
```

# Header rules are centralized in tables.R so all builders use the same red spanner rule.

# Chapter 1 

### Table 1.01 - How Much of U.S. STEM Training Goes to International Students?
```{r table 006}
meta_d <- get_meta("TAB101")
change_type <- "ppt"  # choose "relative" or "ppt"

TAB101 <- build_change_table_wide(
  "TAB101", 
  meta_d, 
  change_type = change_type) %>%
  opt_css(css = "#TAB101 .gt_col_heading { white-space: nowrap !important; }")

save_table(
  TAB101, 
  "tab_1_01", 
  profile = "ultrawide", 
  stub_width_px = 310
  )
```


# Chapter 2 

### Table 2.01 - Where International Students in the United States Come From
```{r table 007}
meta_007 <- get_meta("TAB201")
TAB201 <- build_region_table(
  name = "TAB201",
  meta = meta_007
)
save_table(TAB201, "tab_2_01", profile = "ultrawide", stub_width_px = 300)
```


### Table 2.02 Which Fields Are Each Countries' Students Leaving? 
```{r}
meta <- get_meta("TAB202")
change_type <- "ppt"   # this table only makes sense as percentage points

# The nine field files use two units. TAB202a/b/c are proportions (0.388);
# TAB202d–i are already percent numbers (11.2). Scale only a/b/c by 100 so
# every field lands in the same percent unit and a country's fields sum to ~100.
prop_files <- c("TAB202a", "TAB202b", "TAB202c")
field_files <- c(
  "Engineering"              = "TAB202a",
  "Math & Computer Science"  = "TAB202b",
  "Physical & Life Sciences" = "TAB202c",
  "Business and Management"  = "TAB202d",
  "Education"                = "TAB202e",
  "Fine/Applied Arts"        = "TAB202f",
  "Health Professions"       = "TAB202g",
  "Social Sciences"          = "TAB202h"
)

# Short column headers. The data columns are named by the long field labels
# above; these map each to a compact header whose longest single word fits one
# column at this width, so no header word overflows its cell. Field ordering,
# formatting, and color all key off the long names via field_order and are
# unaffected.
field_header_short <- c(
  "Engineering"              = "Engg",
  "Math & Computer Science"  = "Computing",
  "Physical & Life Sciences" = "Sciences",
  "Business and Management"  = "Business",
  "Education"                = "Edu",
  "Fine/Applied Arts"        = "Arts",
  "Health Professions"       = "Health",
  "Social Sciences"          = "Social Sci."
)

# Rows shown, in order. Sorted by total enrollment, largest first.
country_order <- c("China", "India", "South Korea", "Iran",
                   "Nigeria", "Nepal", "Pakistan", "Bangladesh","Saudi Arabia","Turkey/Türkiye","Taiwan","Spain","Mexico","Colombia","France","Indonesia","Ghana","Italy","Brazil","Vietnam","Hong Kong","Canada","Germany","Japan","United Kingdom","Kuwait","Russia","Thailand","Venezuela")

long_all <- imap_dfr(field_files, function(file, field) {
  read_fig(file) %>%
    mutate(across(-Year, ~ suppressWarnings(as.numeric(na_if(as.character(.x), "-"))))) %>%
    pivot_longer(-Year, names_to = "Country", values_to = "Share") %>%
    filter(!is.na(Share)) %>%
    mutate(
      Share = if (file %in% prop_files) Share * 100 else Share,
      Field = field
    )
})

# First and last available year per country/field, then the point change.
change_long <- long_all %>%
  filter(Country %in% country_order) %>%
  group_by(Country, Field) %>%
  filter(n() >= 2) %>%
  summarise(
    Change = table_change(Share[which.min(Year)], Share[which.max(Year)], change_type),
    .groups = "drop"
  )

# Order fields left-to-right by mean movement across countries, so fields
# gaining share sit at the left and fields being left behind at the right.
field_order <- change_long %>%
  group_by(Field) %>%
  summarise(.avg = mean(Change, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(.avg)) %>%
  pull(Field)

# Wide: countries in rows, fields in columns.
mat <- change_long %>%
  pivot_wider(names_from = Field, values_from = Change) %>%
  mutate(Country = factor(Country, levels = country_order)) %>%
  arrange(Country) %>%
  mutate(Country = as.character(Country)) %>%
  select(Country, all_of(field_order))

# One symmetric domain across the whole matrix so color is comparable cell-to-cell.
lim <- max(abs(as.matrix(mat[field_order])), na.rm = TRUE)

tab_202 <- mat %>%
  gt(rowname_col = "Country", id = "TAB202") %>%
  # These are percentage-point moves, not percent changes, so the old "{x}%"
  # suffix was wrong on all 232 cells. Every cell in this matrix shares one
  # unit, which the spanner and the footnote both state, so the cells carry a
  # bare signed number: repeating " pp" 232 times would cost roughly a column
  # of width in an already dense table. Swap in change_pattern(change_type) if
  # per-cell units are wanted instead.
  fmt_number(
  columns = all_of(field_order),
  n_sigfig = TABLE_SIGFIGS,
  force_sign = TRUE
  ) %>%
  sub_missing(missing_text = "\u2014") %>%
  data_color(
    columns = all_of(field_order),
    fn = scales::col_numeric(
      palette  = TABLE_CHANGE_PALETTE,   # blue (loss) → white (0) → red (gain)
      domain   = c(-lim, lim),
      na.color = "white"
    )
  ) %>%
  cols_align(align = "right", columns = all_of(field_order)) %>%
  cols_label(.list = {
    lbl <- field_header_short[field_order]
    lbl[is.na(lbl)] <- field_order[is.na(lbl)]   # fall back to full name if unmapped
    setNames(as.list(lbl), field_order)
  }) %>%
  style_stub_left() %>%
  tab_spanner(
    label   = "Change in share of the country's U.S.-bound students (pct. points)",
    columns = all_of(field_order)
  ) %>%
  tab_footnote(
    footnote = end_sentence(
      "Each cell is the", change_phrase(change_type),
      "in percentage points, in the share of that",
      "country's U.S.-bound international students enrolled in the field, from",
      "the first to the most recent year available (2010\u20132025 for most",
      "countries; 2012 for Iran, 2018 for Bangladesh). Column headings are",
      "abbreviated: Math/Comp. is mathematics and computer science, Phys./Life",
      "is physical and life sciences, Business is business and management, Fine",
      "Arts is fine and applied arts, Health is health professions, and Social",
      "Sci. is social sciences. Red marks fields gaining share, blue marks",
      "fields losing it. Columns are ordered from the field gaining most across",
      "countries to the field losing most. Countries are ordered by total",
      "international enrollment, largest first"
    ),
    locations = cells_column_spanners(
      spanners = "Change in share of the country's U.S.-bound students (pct. points)"
    )
  ) %>%
  theme_gt_tpa(meta = meta) %>%
  with_table_profile("ultrawide") %>%
  with_size("standard")

save_table(tab_202, "tab_2_02", profile = "ultrawide", stub_width_px = 150)
```



# Chapter 3 

### Table 3.01 - Which STEM Graduates Work in STEM Occupations
```{r tab_14}
meta <- get_meta("TAB301")
change_type <- "ppt"  # choose "relative" or "ppt"
df_029t <- read_fig("TAB301") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Level", values_to = "Share") %>%
  filter(!is.na(Share), Year %in% c(1993, 2003, 2013, 2023)) %>%
  mutate(
    Share = Share * 100,
    Level = case_when(
      str_detect(Level, regex("bach", ignore_case = TRUE)) ~ "Bachelors",
      str_detect(Level, regex("mast", ignore_case = TRUE)) ~ "Masters",
      str_detect(Level, regex("doct", ignore_case = TRUE)) ~ "Doctorate"
    )
  ) %>%
  pivot_wider(names_from = Year, values_from = Share, names_prefix = "Y") %>%
  mutate(
    Change     = table_change(Y1993, Y2023, change_type),
    Occupation = factor(Level, levels = c("Bachelors", "Masters", "Doctorate"))
  ) %>%
  arrange(Occupation) %>%
  select(Occupation, Y1993, Y2003, Y2013, Y2023, Change)
tab_014 <- build_occupation_share_table(
  df_029t, "tab_014", meta,
  level_cols    = c("Y1993", "Y2003", "Y2013", "Y2023"),
  spanner_label = "Share employed in STEM",
  change_type   = change_type,
  footnote = end_sentence(
    "Each year column is the share of graduates at each degree level employed",
    "in a STEM occupation.", change_col_label(FALSE), "is the",
    change_phrase(change_type), "from 1993 to 2023"
  )
)
save_table(tab_014, "tab_3_01", profile = "wide", stub_width_px = 300)
```

### Table 3.02 - How Concentrated Are Temporary Visa Holders in STEM Occupations 
```{r}
meta <- get_meta("TAB302")
change_type <- "ppt"
short_names <- c(
  "Biological, agricultural, and other life scientists" = "Biological scientists",
  "Computer and mathematical scientists"                = "Computer/math scientists",
  "Physical and related scientists"                     = "Physical scientists",
  "Social and related scientists"                        = "Social scientists",
  "Engineers"                                            = "Engineers",
  "S&E-related occupations"                               = "S&E-related occupations",
  "Non-S&E occupations"                                   = "Non-S&E occupations"
)
row_order <- c("Computer/math scientists", "Engineers", "Biological scientists",
               "Physical scientists", "Social scientists",
               "S&E-related occupations", "Non-S&E occupations")
df_011 <- read_fig("TAB302") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Occupation", values_to = "Rate") %>%
  filter(!is.na(Rate), Year %in% c(2017, 2019, 2021, 2023)) %>%
  mutate(Rate = Rate * 100,
         Occupation = recode(Occupation, !!!short_names)) %>%
  pivot_wider(names_from = Year, values_from = Rate, names_prefix = "Y") %>%
  mutate(
    Change     = table_change(Y2017, Y2023, change_type),
    Occupation = factor(Occupation, levels = row_order)
  ) %>%
  arrange(Occupation) %>%
  select(Occupation, Y2017, Y2019, Y2021, Y2023, Change)
TAB302 <- build_occupation_share_table(
  df_011, "TAB302", meta,
  level_cols  = c("Y2017", "Y2019", "Y2021", "Y2023"),
  change_type = change_type
)
save_table(TAB302, "tab_3_02", profile = "wide", stub_width_px = 300)
```


### Table 3.03 - Top Employers Receiving H-1B Approvals
```{r tab_012}
meta <- get_meta("TAB303")
change_type <- "ppt"  # choose "relative" or "ppt"

fig305_shares <- function() {
  raw <- read_fig("FIG305")
  names(raw) <- trimws(names(raw))

  total_row <- raw %>% filter(trimws(Company) == "Grand Total")
  totals <- total_row %>%
    select(-Company, -`Grand Total`) %>%
    pivot_longer(everything(), names_to = "Year", values_to = "Total") %>%
    mutate(Year = as.integer(Year), Total = as.numeric(Total))

  raw %>%
    filter(!trimws(Company) %in% c("Grand Total")) %>%
    mutate(Company = trimws(Company)) %>%
    select(-`Grand Total`) %>%
    pivot_longer(-Company, names_to = "Year", values_to = "Count") %>%
    mutate(Year = as.integer(Year),
           Count = as.numeric(replace_na(Count, 0))) %>%
    filter(Year != 2026) %>%                     # drop partial year
    left_join(totals, by = "Year") %>%
    mutate(Share = Count / Total * 100)
}

shares <- fig305_shares()

# proper-case helper (acronyms preserved, minor words lowercased)
proper_case <- function(x,
                        keep_upper = c("IBM", "HCL", "LLP", "LLC", "US", "USA", "EY"),
                        small = c("and", "of", "the", "for")) {
  vapply(x, function(s) {
    words <- str_split(s, "\\s+")[[1]]
    out <- vapply(seq_along(words), function(i) {
      w <- words[i]
      if (toupper(w) %in% keep_upper)             toupper(w)
      else if (w == "&")                          "&"
      else if (tolower(w) %in% small && i != 1)   tolower(w)
      else                                        str_to_title(w)
    }, character(1))
    paste(out, collapse = " ")
  }, character(1), USE.NAMES = FALSE)
}

# union of firms that were top-10 in EITHER year
union_firms <- shares %>%
  filter(Year %in% c(2017, 2025)) %>%
  group_by(Year) %>%
  slice_max(Count, n = 10) %>%
  ungroup() %>%
  distinct(Company) %>%
  pull(Company)

wide <- shares %>%
  filter(Year %in% c(2017, 2025), Company %in% union_firms) %>%
  select(Company, Year, Share) %>%
  pivot_wider(names_from = Year, values_from = Share, names_prefix = "Y")

df_012 <- wide %>%
  mutate(
    Company = proper_case(Company),
    Y1 = Y2017, Y2 = Y2025,
    Change = if_else(Y2017 > 0 | change_type == "ppt",
                     table_change(Y2017, Y2025, change_type), NA_real_)
  ) %>%
  arrange(desc(Change)) %>%          # order by 2025 share (today's leaders first)
  select(Company, Y1, Y2, Change)

tab_012 <- build_h1b_top10_table(df_012, "tab_012", meta,
                                 first_year = 2017, last_year = 2025,
                                 change_type = change_type)

save_table(tab_012, "tab_3_03", profile = "standard", stub_width_px = 360)
```


# Chapter 4

### Table 4.01 - Four in Ten STEM Entrants Finish with a STEM Degree — and That Barely Moved
```{r tab_018}
meta <- get_meta("TAB401")   # or a dedicated tab_038a entry
change_type <- "ppt"  # choose "relative" or "ppt"

tab_038a <- build_bookend_table(
  name          = "TAB401",
  meta          = meta,
  first         = "2003/04",
  last          = "2011/12",
  id_col        = "cohort",
  spanner_label = "Share of STEM entrants",
  change_type   = change_type,
  recode_cats = c(
    stem_bachelors    = "Completed Bachelor's Degree in STEM Field",
    nonstem_bachelors = "Completed Bachelor's Degree in Non-STEM Field",
    subbac_credential = "Completed A Subbac Credential",
    still_enrolled    = "Still Enrolled",
    left_no_degree    = "Left with No Degree"
  ),
  cat_order = c(
    "Completed Bachelor's Degree in STEM Field",
    "Completed Bachelor's Degree in Non-STEM Field",
    "Completed A Subbac Credential",
    "Still Enrolled",
    "Left with No Degree"
  ),
  footnote = end_sentence(
    "Values are the share of each entering STEM cohort ending in each outcome;",
    "each year column sums to 100%.", change_col_label(FALSE), "is the",
    change_phrase(change_type),
    "from the 2003/04 cohort to the 2011/12 cohort"
  )
)
save_table(tab_038a, "tab_4_01", profile = "standard", stub_width_px = 250)
```


### Table 4.02 - How Many Times Do STEM Majors Change Majors
```{r TAB402}
meta <- get_meta("TAB402")
change_type <- "ppt"  # choose "relative" or "ppt"
TAB402 <- build_bookend_table(
  "TAB402", meta,
  first = "2004/2009 Survey", last = "2012/2017 Survey",
  id_col = "Survey",
  first_label = "2003/04", last_label = "2011/12",
  spanner_label = "Share of STEM entrants",
  change_type = change_type,
  recode_cats = c("Never"          = "Never switched",
                  "Once"           = "Switched once",
                  "More than Once" = "Switched 2+ times"),
  cat_order = c("Never switched", "Switched once", "Switched 2+ times"),
  footnote = end_sentence(
    "Share of STEM entrants by number of major switches.",
    change_col_label(FALSE), "is the", change_phrase(change_type),
    "from the 2004/2009 to the 2012/2017 cohort"
  )
)
save_table(TAB402, "tab_4_02", profile = "standard", stub_width_px = 250)
```

### Table 4.03 - What Do Americans Study Abroad?
```{r tab_016}
meta <- get_meta("TAB403")
change_type <- "ppt"  # choose "relative" or "ppt"
tab_016 <- build_bookend_table(
  "TAB403", meta,
  first = 2005, last = 2024,
  spanner_label = "Share of students",
  change_type = change_type,
  sort_by = "change",
  recode_cats = c("STEM Fields"           = "STEM",
                  "Business & Management" = "Business & management",
                  "Social Sciences"       = "Social sciences",
                  "Fine and Applied Arts" = "Fine & applied arts"),
  flagged_rows = c("Social sciences", "Humanities", "Other"),
  flag_note = paste(
    "Open Doors revised its field taxonomy between 2013 and 2014; shares for",
    "these categories are not directly comparable across that break, so the",
    "reported change reflects reclassification as well as any real shift."
  ),
  footnote = end_sentence(
    "Share of U.S. study-abroad students by field of study.",
    change_col_label(FALSE), "is the", change_phrase(change_type),
    "from 2005 to 2024"
  )
)

save_table(tab_016, "tab_4_03", profile = "standard", stub_width_px = 250)
```

### Table 4.04 - Where Do American's Study Abroad
```{r tab_017}
meta <- get_meta("TAB404")
change_type <- "ppt"  # choose "relative" or "ppt"
tab_017 <- build_bookend_table(
  "TAB404", meta,
  first = 2005, last = 2024,
  spanner_label = "Share of students",
  change_type = change_type,
  sort_by = "change",
  drop_cols = "Antarctica",
  recode_cats = c("Africa, Sub-Saharan"   = "Sub-Saharan Africa",
                  "Multiple Destinations" = "Multiple destinations"),
  footnote = end_sentence(
    "Share of U.S. study-abroad students by host region.",
    change_col_label(FALSE), "is the", change_phrase(change_type),
    "from 2005 to 2024.",
    if (change_type == "relative")
      "Relative change is volatile for small-share regions (e.g., the Middle East & North Africa moved 1.0% to 1.6%)."
    else
      ""
  )
)

save_table(tab_017, "tab_4_04", profile = "standard", stub_width_px = 250)
```


# Chapter 5 

### Table 5.01 - When Did China Catch At Each AI Conference
```{r table 019}

change_type <- "ppt"  # choose "relative" or "ppt"

conf_names <- c(
  acml         = "ACML (ML)",
  nips         = "NeurIPS (ML / AI)",
  icml         = "ICML (ML)",
  cvpr         = "CVPR (Vision)",
  iccv         = "ICCV (Vision)",
  iclr         = "ICLR (ML)",
  emnlp        = "EMNLP (NLP)",
  ijcai        = "IJCAI (AI)",
  corl         = "CoRL (Robotics)",
  rss          = "RSS (Robotics)",
  iros         = "IROS (Robotics)",
  coling       = "COLING (NLP)",
  icra         = "ICRA (Robotics)",
  uai          = "UAI (Prob. ML)",
  naacl        = "NAACL (NLP)",
  acl          = "ACL (NLP)",
  aistats      = "AISTATS (Stats)",
  colt         = "COLT (Theory)",
  siggraphasia = "SIGGRAPH Asia (Graphics)",
  wacv         = "WACV (Vision)"
)

tab_501a <- build_conference_summary_table_split(
  name        = "tab_019a",
  meta        = get_meta("TAB501"),
  conf_names  = conf_names,
  data_name   = "TAB501",
  row_slice   = 1:20,
  change_type = change_type
) %>%

  # Narrow ONLY these four columns.
  # Every pixel removed from them is transferred to the stub.
  with_compact_columns(
    columns = c(
      "US_First",
      "US_Last",
      "CN_First",
      "CN_Last"
    ),
    width_px = 80
  ) %>%

  opt_css(css = "
    #tab_019a .gt_col_heading {
      white-space: nowrap !important;
    }
  ")


save_table(
  tab_501a,
  "tab_5_01",
  profile = "ultrawide",
  stub_width_px = 160
)
```

### Table 5.02 - How Many U.S. Universities Rank in the Global Top 100? 
```{r TAB502}
meta <- get_meta("TAB502")
change_type <- "ppt"  # choose "relative" or "ppt"
num  <- function(x) parse_number(as.character(x))
field_order <- c("Average", "Physical Sciences", "Life Sciences", "Engineering", "Computer Science")
years_shown <- c(2014, 2018, 2022, 2026)
raw <- read_fig("TAB502") %>%
  rename(Year = 1) %>%
  mutate(Year = as.integer(num(Year)))
cs_2017 <- raw %>% filter(Year == 2017) %>% pull(`Computer Science`) %>% num()
df_001 <- raw %>%
  select(Year, all_of(field_order)) %>%                 # drop precomputed Average
  filter(Year %in% years_shown) %>%
  pivot_longer(-Year, names_to = "Field", values_to = "value") %>%
  mutate(value = num(value)) %>%
  pivot_wider(names_from = Year, values_from = value, names_prefix = "Y") %>%
  mutate(
    # computer science was first ranked in 2017; drop its 2017 count into the
    # 2014 slot so its first-year level and Change are populated
    Y2014  = if_else(Field == "Computer Science" & is.na(Y2014), cs_2017, Y2014),
    Change = table_change(Y2014, Y2026, change_type),
    # build_occupation_share_table expects the stub column named "Occupation"
    Occupation = factor(Field, levels = field_order)
  ) %>%
  arrange(Occupation) %>%
  select(Occupation, Y2014, Y2018, Y2022, Y2026, Change)
TAB502 <- build_occupation_share_table(
  df_001, "TAB502", meta,
  level_cols    = c("Y2014", "Y2018", "Y2022", "Y2026"),
  spanner_label = "U.S. universities in the global top 100",
  change_type   = change_type,
  value_type    = "count",
  footnote = end_sentence(
    "Counts are U.S. universities in the global top 100 in each field, shown",
    "for 2014, 2018, 2022, and 2026.", change_col_label(FALSE), "is the",
    change_phrase(change_type, "count"), "from 2014 to 2026.",
    "Computer science was first ranked in 2017; its 2017 count is shown in the",
    "2014 column so its change can be computed"
  )
) %>%
  gt::fmt_number(
  columns = c(Y2014, Y2018, Y2022, Y2026),
  rows = Occupation == "Average",
  n_sigfig = TABLE_SIGFIGS
)

save_table(TAB502, "tab_5_02", profile = "standard")
```

### Table 5.03 - Is R&D Spending Aimed at Today or Tomorrow? 
```{r table_020}
meta <- get_meta("TAB503")

type_labels <- c(
  "Basic research"                              = "Basic research (for tomorrow)",
  "Applied research"                            = "Applied research (for today)",
  "Experimental development"                    = "Experimental development (for today)",
  "Capital expenditure not elsewhere recorded"  = "Capital expenditure (unallocated)"
)
type_order <- c("Basic research (for tomorrow)",
                "Applied research (for today)",
                "Experimental development (for today)",
                "Capital expenditure (unallocated)",
                "Total")

df_125 <- read_fig("TAB503") %>%
  rename(Year = 1, Country = 2, Type = 3, value = 4) %>%
  mutate(Year  = as.numeric(as.character(Year)),
         value = as.numeric(value)) %>%
  filter(Year %in% c(1994, 2004, 2014, 2024)) %>%
  mutate(Type = recode(Type, !!!type_labels)) %>%
  pivot_wider(names_from = Year, values_from = value, names_prefix = "Y")

totals_125 <- df_125 %>%
  group_by(Country) %>%
  summarise(across(c(Y1994, Y2004, Y2014, Y2024), ~ sum(.x, na.rm = TRUE)),
            .groups = "drop") %>%
  mutate(Type = "Total")

df_125 <- bind_rows(df_125, totals_125) %>%
  mutate(
    Change  = table_change(Y1994, Y2024, "relative"),
    Country = factor(Country, levels = c("United States", "China")),
    Type    = factor(Type, levels = type_order)
  ) %>%
  arrange(Country, Type) %>%
  select(Country, Type, Y1994, Y2004, Y2014, Y2024, Change)

tab_021 <- build_rd_type_level_table(
  df_125, "tab_5_03", meta,
  level_cols = c("Y1994", "Y2004", "Y2014", "Y2024")
)

save_table(
  tab_021,
  "tab_5_03",
  profile = "ultrawide",
  stub_width_px = 350
)
```

### Table 5.04 - Business Now Dominates R&D in Both Countries 
```{r tab_022}
meta <- get_meta("TAB504")

sector_labels <- c(
  "Business enterprise sector" = "Business enterprise sector",
  "Government sector"          = "Government sector",
  "Higher education sector"    = "Higher education sector",
  "Private non-profit sector"  = "Private non-profit sector"
)
sector_order <- c("Business enterprise sector", "Government sector",
                  "Higher education sector", "Private non-profit sector", "Total")

df_022 <- read_fig("TAB504") %>%
  rename(Year = 1, Country = 2, Type = 3, value = 4) %>%
  mutate(Year  = as.numeric(as.character(Year)),
         value = as.numeric(value)) %>%
  filter(Type %in% names(sector_labels),
         Year %in% c(1994, 2004, 2014, 2024)) %>%
  mutate(Type = recode(Type, !!!sector_labels)) %>%
  complete(Country, Type, Year, fill = list(value = 0)) %>%   # China private non-profit -> $0
  pivot_wider(names_from = Year, values_from = value, names_prefix = "Y")

# per-country total row, summed across the four performing sectors
totals_022 <- df_022 %>%
  group_by(Country) %>%
  summarise(across(c(Y1994, Y2004, Y2014, Y2024), ~ sum(.x, na.rm = TRUE)),
            .groups = "drop") %>%
  mutate(Type = "Total")

df_022 <- bind_rows(df_022, totals_022) %>%
  mutate(
    Change  = table_change(Y1994, Y2024, "relative"),
    Change  = ifelse(is.finite(Change), Change, 0),   # 0/0 -> 0%
    Country = factor(Country, levels = c("United States", "China")),
    Type    = factor(Type, levels = sector_order)
  ) %>%
  arrange(Country, Type) %>%
  select(Country, Type, Y1994, Y2004, Y2014, Y2024, Change)

tab_022 <- build_rd_type_level_table(
  df_022, "tab_5_04", meta,
  level_cols = c("Y1994", "Y2004", "Y2014", "Y2024"),
  footnote = end_sentence(
    "Spending levels in billions of US dollars, PPP converted, at current",
    "prices (not adjusted for inflation).", change_col_label(FALSE), "is the",
    "relative change from the first to the last shown year. Sectors denote",
    "where R&D is performed. China reports no private non-profit R&D; it is",
    "shown as $0, and its change cell is undefined from a zero base and appears",
    "as 0%. Totals may not sum exactly because of rounding"
  )
)

save_table(
  tab_022,
  "tab_5_04",
  profile = "ultrawide",
  stub_width_px = 350
)
```

### Table 5.05 - Who Patents in Critical Technologies
```{r table 020}
change_type <- "relative"  # choose "relative" or "ppt"

build_patents_table(
  name    = "tab_020",     # output id; adjust to next free table number
  meta    = get_meta("TAB505"),
  sources = c(
    "Biotechnology"       = "TAB505a",
    "Semiconductors"      = "TAB505b",
    "Computer technology" = "TAB505c"
  ),
  change_type = change_type
) %>%
  save_table("tab_5_05", profile = "ultrawide")
```

### Table 5.06 - Patent Leadership Shifted From U.S. and Japanese Incumbents to Asian Challengers
```{r TAB506}
meta <- get_meta("TAB506")
# "ppt" here means "native units": these are counts of grants, so the change is
# an absolute count difference rather than a percentage-point move.
change_type <- "ppt"
num  <- function(x) parse_number(as.character(x))
flag_slug <- c(
  "United States" = "united-states",
  "Japan"         = "japan",
  "Taiwan"        = "taiwan",
  "South Korea"   = "south-korea",
  "Germany"       = "germany",
  "China"         = "china"
)
df_023 <- read_fig("TAB506") %>%
  rename(Company = 1, Country = 2, Y2005 = 3, Y2015 = 4, Y2025 = 5) %>%
  mutate(across(c(Y2005, Y2015, Y2025), num)) %>%
  mutate(
    flag     = paste0("inputs/flags/", recode(Country, !!!flag_slug), ".png"),
    Change   = table_change(Y2005, Y2025, change_type),
    Rank2005 = dplyr::min_rank(dplyr::desc(Y2005)),   # 1 = most 2005 grants
    Rank2025 = dplyr::min_rank(dplyr::desc(Y2025))    # 1 = most 2025 grants
  )
# mark the five firms that were top 5 in 2005 with a colored rank arrow.
# arrow points down (fell in rank) or up (rose) between 2005 and 2025.
arrow_span <- function(rank2005, fell) {
  color <- TABLE_TEXT_COLOR                         # neutral; color is reserved for data cells
  glyph <- if (fell) "\u25BE" else "\u25B4"      # ▾ / ▴
  paste0(" <span style='color:", color, ";font-weight:700;'>",
         glyph, " #", rank2005, "</span>")
}
df_023 <- df_023 %>%
  mutate(
    Company = ifelse(
      Rank2005 <= 5,
      paste0(Company, mapply(arrow_span, Rank2005, Rank2025 > Rank2005)),
      Company
    )
  ) %>%
  arrange(Rank2025) %>%                            # sort by 2025 rank, not net change
  select(Company, flag, Y2005, Y2015, Y2025, Change)
TAB506 <- build_patent_company_delta_table(
  df_023, "TAB506", meta,
  level_cols  = c("Y2005", "Y2015", "Y2025"),
  chg_cols    = character(0),
  change_type = change_type,
  footnote = end_sentence(
    "The first three columns are patent grants in 2005, 2015, and 2025.",
    change_col_label(FALSE), "is the", change_phrase(change_type, "count"),
    "from 2005 to 2025. Arrows mark the five firms that ranked in the top five",
    "by 2005 grants, with their 2005 rank; a down arrow marks a fall in rank",
    "by 2025 and an up arrow marks a rise. Rows are ordered by 2025 grants"
  )
)
save_table(TAB506, "tab_5_06", profile = "ultrawide", stub_width_px = 360)
```

# Chapter 6

### Table 6.01 - Do Stay Intentions Differ by Region of Origin? 
```{r table 028}
meta <- get_meta("TAB601")
change_type <- "ppt"  # choose "relative" or "ppt"

tab_028 <- build_bookend_table(
  "TAB601", meta,
  first = 2002, last = 2024,
  spanner_label = "Intend to stay",
  scale = 100,
  drop_cols = "Region and country or economy unknown",
  change_type = change_type
)

save_table(tab_028, "tab_6_01", profile = "standard")
```


### Table 6.02 - What New STEM Phds Plan to Do After Graduations
```{r table 025}
meta <- get_meta("TAB602")
change_type <- "ppt"  # choose "relative" or "ppt"

tab_025 <- build_bookend_table(
  "TAB602", meta,
  first = 2021, last = 2024,
  spanner_label = "Share with U.S. plans",
  sort_by = "change",
  scale = 100,
  change_type = change_type
)

save_table(tab_025, "tab_6_02", profile = "standard")
```

### Table 6.03 - Where U.S. Awarded PhDs Work After Graduation
```{r table 027}
meta <- get_meta("TAB603")
change_type <- "ppt"  # choose "relative" or "ppt"

TAB603 <- build_sector_bookend_table(meta, change_type = change_type)

save_table(TAB603, "tab_6_03", profile = "standard")
TAB603
```

### Table 6.04 - Employment-Based Green Card Wait Times by Nationality
```{r table 026}
meta <- get_meta("TAB604")

TAB604 <- build_eb_combined_table(meta)

save_table(TAB604, "tab_6_04", profile = "standard")
TAB604
```

```{r}
library(magick)
library(dplyr)
library(purrr)
library(tibble)

# ---------------------------------------------------------------------------
# Diagnose rendered header-rule thickness directly from final PNGs
# ---------------------------------------------------------------------------

png_dir <- file.path(PATHS$final_tables, "pngs")

# Use the same red as the tables.
rule_red <- if (exists("tpa_colors")) tpa_colors[1] else "#8C1515"

# Convert "#RRGGBB" to numeric RGB.
hex_rgb <- function(hex) {
  hex <- gsub("#", "", hex)
  c(
    R = strtoi(substr(hex, 1, 2), 16L),
    G = strtoi(substr(hex, 3, 4), 16L),
    B = strtoi(substr(hex, 5, 6), 16L)
  )
}

target <- hex_rgb(rule_red)

# Lengths of all TRUE runs in one pixel row.
true_runs <- function(x) {
  rr <- rle(x)
  rr$lengths[rr$values]
}

measure_header_rules <- function(path,
                                 tolerance = 35,
                                 min_bar_fraction = 0.08,
                                 max_rule_thickness = 12) {

  img <- image_read(path)
  dat <- image_data(img, channels = "rgb")

  # magick array is channel x width x height
  w <- dim(dat)[2]
  h <- dim(dat)[3]

  r <- t(matrix(as.integer(dat[1, , ]), nrow = w, ncol = h))
  g <- t(matrix(as.integer(dat[2, , ]), nrow = w, ncol = h))
  b <- t(matrix(as.integer(dat[3, , ]), nrow = w, ncol = h))

  # "Red enough" to catch the rule even if Chromium anti-aliases an edge.
  dist <- sqrt(
    (r - target["R"])^2 +
    (g - target["G"])^2 +
    (b - target["B"])^2
  )

  red_mask <- dist <= tolerance

  # We only care about the header area.
  # This excludes nearly all heatmap cells farther down the table.
  y_min <- max(1, round(h * 0.07))
  y_max <- min(h, round(h * 0.45))

  min_bar_px <- round(w * min_bar_fraction)

  # Longest continuous red run in every row.
  longest_run <- vapply(
    seq_len(h),
    function(y) {
      runs <- true_runs(red_mask[y, ])
      if (length(runs) == 0) 0L else max(runs)
    },
    integer(1)
  )

  candidate_y <- which(
    seq_len(h) >= y_min &
    seq_len(h) <= y_max &
    longest_run >= min_bar_px
  )

  if (length(candidate_y) == 0) {
    return(
      tibble(
        table = basename(path),
        image_width_px = w,
        image_height_px = h,
        y_start = NA_integer_,
        y_end = NA_integer_,
        thickness_px = NA_integer_,
        number_of_bars = NA_integer_,
        bar_lengths_px = NA_character_,
        longest_bar_px = NA_integer_
      )
    )
  }

  # Group consecutive rows. A 3-pixel rule becomes one 3-row group.
  group_id <- cumsum(c(TRUE, diff(candidate_y) > 1))

  groups <- split(candidate_y, group_id)

  map_dfr(groups, function(ys) {

    thickness <- length(ys)

    # Large blocks are almost certainly heatmap/body fills, not header rules.
    if (thickness > max_rule_thickness) return(NULL)

    # Inspect the center row of the detected rule.
    y_mid <- ys[ceiling(length(ys) / 2)]

    runs <- true_runs(red_mask[y_mid, ])
    runs <- sort(runs[runs >= min_bar_px], decreasing = TRUE)

    tibble(
      table = basename(path),
      image_width_px = w,
      image_height_px = h,
      y_start = min(ys),
      y_end = max(ys),

      # THIS is the number we care about for rule consistency.
      thickness_px = thickness,

      # Useful secondary checks:
      number_of_bars = length(runs),
      bar_lengths_px = paste(runs, collapse = ", "),
      longest_bar_px = ifelse(length(runs), max(runs), NA_integer_)
    )
  })
}


# ---------------------------------------------------------------------------
# Run across every exported table
# ---------------------------------------------------------------------------

png_files <- list.files(
  png_dir,
  pattern = "^tab_.*\\.png$",
  full.names = TRUE
)

bar_diagnostics <- map_dfr(
  png_files,
  measure_header_rules
) %>%
  arrange(table, y_start)

print(bar_diagnostics, n = Inf)
```