title: "GTCI Figures"
author: "Emerson Johnston"
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = FALSE,
  warning   = FALSE,
  message   = FALSE,
  fig.width  = 8.0,
  fig.height = 5.5,
  dpi       = 300,
  fig.align = "center",
  dev       = "ragg_png"
)

library(tidyverse)
library(scales)
library(showtext)
library(ggtext)
library(ggrepel)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(gt)
library(gtExtras)
library(webshot2)
library(magick)
library(pdftools)

source("R/theme.R")
source("R/helpers.R")
source("R/tables.R")
source("R/chart_layers.R")
source("R/maps.R")
```

# Chapter 1

### Chapter 1 - Figure 01 - Total STEM Degree Completition in the United States, by Level (1995 - 2024)
```{r figure 0001}
meta <- get_meta("fig_001")

df_001 <- read_levels("fig_001", value_col = "Count",
                      levels_order = c("Doctorate", "Masters", "Bachelors"))

fig_001 <- ggplot(df_001, aes(Year, Count, fill = Level)) +
  geom_area(alpha = ALPHA_AREA) +
  scale_x_years(df_001$Year) +
  scale_y_continuous(labels = label_comma(),
                     expand = EXPAND_BAR) +
  scale_y_continuous(breaks = c(250000, 500000, 750000),
                   labels = label_comma(),
                   expand = EXPAND_BAR) +
  scale_fill_levels(reverse_legend = FALSE) +
  labs_meta(meta, y = "Degrees awarded") +
  theme_tpa()

save_chart(fig_001, "chapter1_fig_001")
```

### Chapter 1 - Figure 02 - STEM Degrees Awarded as a Share of Total Degrees Awarded - All levels
```{r figure 002}
meta <- get_meta("fig_002")
pct  <- label_percent(accuracy = 1)
df_002 <- read_levels(
  "fig_002",
  value_col    = "Share",
  levels_order = c("Bachelors", "Masters",
                   "All Doctorates (IPEDS)")
) %>%
  filter(!is.na(Share))

last_002  <- last_by_group(df_002, Year, Level)
first_002 <- df_002 %>% group_by(Level) %>% slice_min(Year, n = 1) %>% ungroup()

fig_002 <- ggplot(df_002, aes(Year, Share, color = Level, group = Level)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_label_repel_tpa(
    data         = last_002,
    mapping      = aes(label = paste0(pct(Share), " - ", Level)),
    nudge_x      = 0.5,
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  geom_label_repel_tpa(
    data         = first_002,
    mapping      = aes(label = pct(Share)),
    nudge_x      = -0.5,
    hjust        = 1,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_continuous(
    breaks = seq(min(df_002$Year), max(df_002$Year), 2),
    expand = expansion(mult = c(0.05, .5))
  ) +
  scale_y_continuous(labels = pct, expand = EXPAND_LINE_WIDE) +
  scale_color_levels() +
  guides(color = "none") +
  labs_meta(meta, y = "Share of completions") +
  theme_tpa()
save_chart(fig_002, "chapter1_fig_002")
```

### Chapter 1 - Figure 03 - Trends in Research Doctorate Recipients from U.S. Colleges and Universities, All Fields, by Year (1958-2023)
```{r figure 003}
meta <- get_meta("fig_003")

df_003 <- read_fig("fig_003") %>%
  select(Year, value = 2) %>%
  mutate(across(everything(), as.numeric))

labels_003 <- first_last_peak(df_003, Year, value)

fig_003 <- ggplot(df_003, aes(Year, value)) +
  geom_line(linewidth = LINE_WIDTH, color = tpa_colors[1]) +
  geom_label_repel_tpa(
    data         = labels_003,
    mapping      = aes(label = fmt_num_k(value)),
    color        = tpa_colors[1],
    nudge_y      = 3000,
    with_segment = TRUE
  ) +
  scale_x_years(df_003$Year) +
  scale_y_continuous(labels = label_comma(),
                     expand = EXPAND_LINE_TALL) +
  labs_meta(meta, y = "Doctorate recipients") +
  theme_tpa()

save_chart(fig_003, "chapter1_fig_003")
```

### Chapter 1 - Figure 05 - International Student Share of STEM Degree Completions
```{r figure 005}
meta <- get_meta("fig_005")
pct  <- label_percent(accuracy = 1)
df_005 <- read_levels(
  "fig_005",
  value_col    = "Share",
  levels_order = c("Bachelors", "Masters",
                   "All Doctorates (IPEDS)", "Research Doctorates (SED)")
) %>%
  filter(!is.na(Share))

last_005  <- last_by_group(df_005, Year, Level)
first_005 <- df_005 %>% group_by(Level) %>% slice_min(Year, n = 1) %>% ungroup()

fig_005 <- ggplot(df_005, aes(Year, Share, color = Level, group = Level)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_label_repel_tpa(
    data         = last_005,
    mapping      = aes(label = paste0(pct(Share), ": ", Level)),
    nudge_x      = 0.5,
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  geom_label_repel_tpa(
    data         = first_005,
    mapping      = aes(label = pct(Share)),
    nudge_x      = -0.5,
    hjust        = 1,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_continuous(
    breaks = seq(min(df_005$Year), max(df_005$Year), 2),
    expand = expansion(mult = c(0.05, .5))
  ) +
  scale_y_continuous(labels = pct, expand = EXPAND_LINE_WIDE) +
  scale_color_levels() +
  guides(color = "none") +
  labs_meta(meta, y = "Share of completions") +
  theme_tpa()
save_chart(fig_005, "chapter1_fig_005")
```


### Chapter 1 - Figure 07 - Projected Annual STEM PhD Production Per 100,000 of the 20- to 34-year-old population, by Country, 2011–2050
```{r figure 007}
meta <- get_meta("fig_007")

df_007 <- read_fig("fig_007") %>%
  pivot_observed_projected() %>%
  order_by_latest(Country, value)

labels_007 <- last_by_group(df_007, Year, Country)
proj_start <- proj_start_year(df_007)
y_max      <- max(df_007$value)

fig_007 <- ggplot(df_007, aes(Year, value, color = Country, group = Country)) +
  annotate_projected(proj_start, max(df_007$Year),
                     label_y = y_max * 0.98) +
  geom_op_lines(df_007) +
  geom_label_repel_tpa(
    data    = labels_007,
    mapping = aes(label = paste0(
      if_else(Country == "United Kingdom", "United\nKingdom", as.character(Country)),
      ": ", round(value)
    )),
    nudge_x    = 2,
    hjust      = 0,
    lineheight = 0.85,
    with_segment = TRUE
  ) +
  scale_x_years(df_007$Year, pad_right = PAD_RIGHT_NARROW) +
  scale_y_continuous(labels = label_comma(),
                     expand = EXPAND_LINE) +
  scale_color_tpa() +
  guides(color = "none") +
  labs_meta(meta, y = "STEM PhDs per year") +
  theme_op()

save_chart(fig_007, "chapter1_fig_007")
```

### Chapter 1 - Figure 08 - Projected Absolute Annual STEM PhD Production by Country, 2011–2050
```{r figure 008}
meta <- get_meta("fig_008")

df_008 <- read_fig("fig_008") %>%
  pivot_observed_projected() %>%
  order_by_latest(Country, value)

labels_008 <- last_by_group(df_008, Year, Country)
proj_start <- proj_start_year(df_008)
y_top      <- y_top_round(df_008$value, 50000)

milestones <- tribble(
  ~year, ~label,
  2019, "China Tops US in STEM PhDs",
  2033, "India Tops US in STEM PhDs",
  2042, "India Tops China in STEM PhDs, Both now produce 1.5x US",
  2050, "India Doubles US STEM Production"
)

fig_008 <- ggplot(df_008, aes(Year, value, color = Country, group = Country)) +
  annotate_projected(proj_start, max(df_008$Year) + 0.5,
                     ymax = y_top, label_y = y_top * 0.92) +
  geom_op_lines(df_008) +
  annotate_milestones(
    milestones, y_top,
    bg_for = \(year) if_else(year >= proj_start, "grey94", "white")
  ) +
  geom_label_repel_tpa(
    data         = labels_008,
    mapping      = aes(label = Country),
    nudge_x      = 2,
    hjust        = 0,
    with_segment = TRUE
  ) +
  scale_x_years(df_008$Year, pad_right = PAD_RIGHT_NARROW) +
  scale_y_continuous(labels = label_comma(),
                     expand = EXPAND_LINE) +
  scale_color_tpa() +
  guides(color = "none") +
  labs_meta(meta, y = "STEM PhDs per year") +
  theme_op()

save_chart(fig_008, "chapter1_fig_008")
```

### Chapter 1 - Figure 09 - International Students in US Higher Education - Total Enrollment and Share, 1949-2025
```{r figure 009}
meta <- get_meta("fig_009")

df_009 <- read_fig("fig_009") %>%
  rename(Year = 1, Total = 2, Share = 3) %>%
  mutate(across(everything(), as.numeric),
         Share = fill_gaps(Year, Share))

scale_factor <- dual_axis_scale(
  primary_max   = max(df_009$Total, na.rm = TRUE),
  secondary_max = 0.07
)

fig_009 <- ggplot(df_009, aes(Year)) +
  geom_col(aes(y = Total, fill = "Total"),
           width = 0.7, alpha = ALPHA_COL_LIGHT) +
  geom_line(aes(y = Share * scale_factor, color = "Share"),
            linewidth = LINE_WIDTH, na.rm = TRUE) +
  scale_x_years(df_009$Year) +
  scale_y_continuous(
    name   = "Count of international students",
    labels = label_comma(),
    expand = EXPAND_BAR,
    sec.axis = sec_axis_scaled(scale_factor,
                               name   = "Share of total US enrollment",
                               labels = label_percent(accuracy = 1))
  ) +
  scale_fill_manual(name = NULL, values = c("Total" = tpa_colors[2])) +
  scale_color_manual(name = NULL, values = c("Share" = tpa_colors[1])) +
  labs_meta(meta) +
  theme_dual_axis()

save_chart(fig_009, "chapter1_fig_009")
```

### Chapter 1 - Figure 98 - Trends in Doctoral Production by Degree Level
```{r figure 98}
meta <- get_meta("fig_098")

m1 <- function(x) sprintf("%.1fM", x / 1e6)

df_098 <- read_levels("fig_098", value_col = "Count",
                      levels_order = c("Bachelors", "Masters", "Doctorate"))

last_098  <- last_by_group(df_098, Year, Level)
first_098 <- df_098 %>% group_by(Level) %>% slice_min(Year, n = 1) %>% ungroup()

fig_098 <- ggplot(df_098, aes(Year, Count, color = Level, group = Level)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_label_repel_tpa(
    data         = last_098,
    mapping      = aes(label = paste0(Level, ": ", m1(Count))),
    nudge_x      = 0.5,
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  geom_label_repel_tpa(
    data         = first_098,
    mapping      = aes(label = m1(Count)),
    nudge_x      = -0.5,
    hjust        = 1,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_continuous(
    breaks = seq(min(df_098$Year), max(df_098$Year), 1),
    expand = expansion(mult = c(0.05, PAD_RIGHT_NARROW))
  ) +
  scale_y_continuous(labels = m1, expand = EXPAND_LINE) +
  scale_color_levels() +
  guides(color = "none") +
  labs_meta(meta, y = "Degrees awarded") +
  theme_tpa()

save_chart(fig_098, "chapter1_fig_098")
```

### Chapter 1 - Figure 99 - International Student Share of All Degree Completions
```{r figure 099}
meta <- get_meta("fig_099")
pct  <- label_percent(accuracy = 1)
df_099 <- read_levels(
  "fig_099",
  value_col    = "Share",
  levels_order = c("Bachelors", "Masters",
                   "All Doctorates (IPEDS)", "Research Doctorates (SED)")
) %>%
  filter(!is.na(Share))
last_099  <- last_by_group(df_099, Year, Level)
first_099 <- df_099 %>% group_by(Level) %>% slice_min(Year, n = 1) %>% ungroup()

fig_099 <- ggplot(df_099, aes(Year, Share, color = Level, group = Level)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_label_repel_tpa(
    data         = last_099,
    mapping      = aes(label = paste0(pct(Share), " - ", Level)),
    nudge_x      = 0.5,
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  geom_label_repel_tpa(
    data         = first_099,
    mapping      = aes(label = pct(Share)),
    nudge_x      = -0.5,
    hjust        = 1,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_continuous(
    breaks = seq(min(df_099$Year), max(df_099$Year), 2),
    expand = expansion(mult = c(0.05, .50))
  ) +
  scale_y_continuous(labels = pct, expand = EXPAND_LINE_WIDE) +
  scale_color_levels() +
  guides(color = "none") +
  labs_meta(meta, y = "Share of completions") +
  theme_tpa()
save_chart(fig_099, "chapter1_fig_099")
```

### Chapter 1 - Figure 100 - Trends in Research Doctorate Recipients from U.S. Colleges and Universities, All Fields, by Year (1958-2023)
```{r figure 100}
meta <- get_meta("fig_100")

df_100 <- read_fig("fig_100") %>%
  mutate(
    across(everything(), as.numeric),
    stem_share = 100 * `Just S&E fields` / `Doctorate recipients (number)`
  )

labels_100 <- first_last_peak(
  df_100,
  Year,
  `Doctorate recipients (number)`
)

stem_labels <- bind_rows(
  df_100 %>% slice_min(Year, n = 1),
  df_100 %>% slice_max(stem_share, n = 1)
) %>%
  distinct(Year, .keep_all = TRUE)

fig_100 <- ggplot(df_100, aes(x = Year)) +

  # STEM area
  geom_ribbon(
    aes(
      ymin = 0,
      ymax = `Just S&E fields`
    ),
    fill = tpa_colors[2],
    alpha = 0.30
  ) +

  # Non-STEM area stacked above STEM
  geom_ribbon(
    aes(
      ymin = `Just S&E fields`,
      ymax = `Doctorate recipients (number)`
    ),
    fill = "grey80",
    alpha = 0.45
  ) +

  # Original line
  geom_line(
    aes(y = `Doctorate recipients (number)`),
    linewidth = LINE_WIDTH,
    color = tpa_colors[1]
  ) +

  # Total doctorate labels
  geom_label_repel_tpa(
    data = labels_100,
    mapping = aes(
      y = `Doctorate recipients (number)`,
      label = fmt_num_k(`Doctorate recipients (number)`)
    ),
    color = tpa_colors[1],
    nudge_y = 5000,
    with_segment = TRUE
  ) +

  # STEM share labels (first year + peak share year)
  geom_label_repel_tpa(
    data = stem_labels,
    mapping = aes(
      y = `Just S&E fields`,
      label = paste0(round(stem_share), "%")
    ),
    color = tpa_colors[2],
    nudge_y = -5000,
    with_segment = TRUE
  ) +

  scale_x_years(df_100$Year) +
  scale_y_continuous(
    labels = label_comma(),
    expand = EXPAND_LINE_TALL
  ) +

  labs_meta(meta, y = "Doctorate recipients") +
  theme_tpa()

save_chart(fig_100, "chapter1_fig_100")
```

# Chapter 2

### Chapter 2 - Figure 10 - Economic Contributions of International Students to the U.S. Economy, 1999-2024
```{r figure 010}
meta <- get_meta("fig_010")

df_010 <- read_fig("fig_010") %>%
  rename(value = 2) %>%
  mutate(Year  = as.numeric(Year),
         value = as.numeric(value) / 1e9)

labels <- first_last_peak(df_010, Year, value)

fig_010 <- ggplot(df_010, aes(Year, value)) +
  geom_col(fill = tpa_colors[1], width = 0.7, alpha = ALPHA_COL) +
  geom_text_tpa(data    = labels,
                mapping = aes(label = sprintf("%.2f", value)),
                color   = tpa_colors[1]) +
  scale_x_years(df_010$Year, by = 1) +
  scale_y_continuous(labels = label_dollar(suffix = "B"),
                     expand = EXPAND_BAR_TALL) +
  labs_meta(meta, y = "USD (billions)") +
  theme_tpa()

save_chart(fig_010, "chapter2_fig_010")
```

### Chapter 2 - Figure 13 - Trends in International Student Enrollment in STEM Fields in U.S. Higher Education, by Country (2009–2024)
```{r figure 013}
meta <- get_meta("fig_013")

df_013 <- read_fig("fig_013") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Country", values_to = "Count") %>%
  filter(!is.na(Count))

top_countries <- df_013 %>%
  filter(Year == max(Year)) %>%
  slice_max(Count, n = 6) %>%
  pull(Country)

df_013 <- df_013 %>%
  filter(Country %in% top_countries) %>%
  order_by_latest(Country, Count)

labels_013 <- last_by_group(df_013, Year, Country)

fig_013 <- ggplot(df_013, aes(Year, Count, color = Country, group = Country)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_label_repel_tpa(
    data         = labels_013,
    mapping      = aes(label = paste0(Country, ": ", label_comma()(Count))),
    nudge_x      = 0.5,
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_013$Year, pad_right = .3) +
  scale_y_log10(labels = label_comma(),
                breaks = c(1000, 5000, 10000, 50000, 100000, 250000),
                expand = EXPAND_LINE) +
  scale_color_tpa() +
  guides(color = "none") +
  labs_meta(meta, y = "Enrolled students (log scale)") +
  theme_tpa()

save_chart(fig_013, "chapter2_fig_013")
```

### Chapter 2 - Figure 14 - Trends in International Student Enrollment in STEM Fields in U.S. Higher Education, by Broad STEM Field (2009–2024)
```{r figure 014}
meta <- get_meta("fig_014")

df_014 <- read_fig("fig_014") %>%
  select(-`Non STEM`,-`Social Sciences`) %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Field", values_to = "Count") %>%
  filter(!is.na(Count)) %>%
  order_by_latest(Field, Count)

labels_014 <- last_by_group(df_014, Year, Field)

fig_014 <- ggplot(df_014, aes(Year, Count, color = Field, group = Field)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_label_repel_tpa(
    data         = labels_014,
    mapping      = aes(label = paste0(Field, ": ", label_comma()(Count))),
    nudge_x      = 1,
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_014$Year, pad_right = .5) +
  scale_y_continuous(labels = label_comma(),
                     expand = EXPAND_LINE) +
  scale_color_tpa() +
  guides(color = "none") +
  labs_meta(meta, y = "Enrolled students") +
  theme_tpa()

save_chart(fig_014, "chapter2_fig_014")
```

### Chapter 2 - Figure 20 - International Postdoctoral Representation by Field, 1980–2023
```{r figure 020}
meta <- get_meta("fig_020")

df_020 <- read_fig("fig_020") %>%
  rename(Year = 1,
         Science     = 2,
         Engineering = 3,
         Health      = 4) %>%
  mutate(across(everything(), as.numeric)) %>%
  pivot_longer(-Year, names_to = "Field", values_to = "Share") %>%
  filter(!is.na(Share)) %>%
  order_by_latest(Field, Share)

labels_020 <- last_by_group(df_020, Year, Field)

fig_020 <- ggplot(df_020, aes(Year, Share, color = Field, group = Field)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_label_repel_tpa(
    data         = labels_020,
    mapping      = aes(label = paste0(Field, ": ", sprintf("%.1f%%", Share))),
    nudge_x      = 0.9,
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_020$Year, pad_right = .25) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(0, 80),
                     breaks = seq(0, 80, 10),
                     expand = EXPAND_BAR_TIGHT) +
  scale_color_tpa() +
  guides(color = "none") +
  labs_meta(meta, y = "International share of postdocs") +
  theme_tpa()

save_chart(fig_020, "chapter2_fig_020")
```

### Chapter 2 - Figure 21 - Trends in Postdocs at Federally Funded Research and Development Centers, by Citizenship (2012-2023)
```{r figure 021}
meta <- get_meta("fig_021")

df_021 <- read_fig("fig_021") %>%
  rename(Year = 1,
         `U.S. citizens and permanent residents` = 2,
         `Temporary visa holders`                = 3) %>%
  mutate(across(everything(), as.numeric)) %>%
  pivot_longer(-Year, names_to = "Status", values_to = "Count") %>%
  filter(!is.na(Year), !is.na(Count)) %>%
  order_by_latest(Status, Count)

labels_021 <- df_021 %>%
  nudge_by_series(Status,
                  above  = "Temporary visa holders",
                  amount = 120)

fig_021 <- ggplot(df_021, aes(Year, Count, color = Status, group = Status)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_point(size = 2) +
  geom_label_repel_tpa(
    data         = labels_021,
    mapping      = aes(label = label_comma()(Count)),
    nudge_y      = labels_021$nudge_y,
    direction    = "y",
    with_segment = FALSE
  ) +
  scale_x_years(df_021$Year, by = 1, pad_right = PAD_RIGHT_TIGHT) +
  scale_y_continuous(labels = label_comma(),
                     limits = c(0, 2500),
                     breaks = seq(0, 2500, 500),
                     expand = EXPAND_BAR_TIGHT) +
  scale_color_manual(values = citizenship_colors_alt) +
  labs_meta(meta, y = "Postdoctoral researchers") +
  theme_tpa()

save_chart(fig_021, "chapter2_fig_021")
```

### Chapter 2 - Figure 102 - International Students in U.S. Higher Education, Count by Academic Level
```{r fig_102}
meta <- get_meta("fig_102")

df_102 <- read_fig("fig_102") %>%
  rename(
    `Graduate (unspecified)` = `Graduate - Unspecified`,
    `Non-degree`             = `Non-Degree`,
    OPT                      = `Optional Practical Training (OPT)`
  ) %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Level", values_to = "Count") %>%
  order_by_latest(Level, Count)

labels_102 <- last_by_group(df_102, Year, Level)

fig_102 <- ggplot(df_102, aes(Year, Count, color = Level, group = Level)) +
  geom_line(linewidth = 1) +
  geom_label_repel_tpa(
    data = labels_102,
    mapping = aes(label = paste0(Level, ": ", label_comma()(Count))),
    nudge_x = 0.5,
    hjust = 0,
    direction = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_102$Year, pad_right = 0.50) +
  scale_y_continuous(labels = label_comma(),
                     expand = expansion(mult = c(0.02, 0.05))) +
  scale_color_manual(values = acad_level_colors) +
  guides(color = "none") +
  labs_meta(meta, y = "International students enrolled") +
  theme_tpa()

save_chart(fig_102, "chapter2_fig_102")
```

### Chapter 2 - Figure 103 - International Students in U.S. Higher Education, Share by Academic Level
```{r fig_103}
meta <- get_meta("fig_103")

df_103 <- read_fig("fig_103") %>%
  rename(
    `Graduate (unspecified)` = `Graduate - Unspecified`,
    `Non-degree`             = `Non-Degree`,
    OPT                      = `Optional Practical Training (OPT)`
  ) %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Level", values_to = "Count") %>%
  order_by_latest(Level, Count)

labels_103 <- last_by_group(df_103, Year, Level)

fig_103 <- ggplot(df_103, aes(Year, Count, color = Level, group = Level)) +
  geom_line(linewidth = 1) +
  geom_label_repel_tpa(
    data = labels_103,
    mapping = aes(label = paste0(Level, ": ", label_comma()(Count),"%")),
    nudge_x = 0.5,
    hjust = 0,
    direction = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_103$Year, pad_right = 0.50) +
  scale_y_continuous(labels = label_comma(),
                     expand = expansion(mult = c(0.02, 0.05))) +
  scale_color_manual(values = acad_level_colors) +
  guides(color = "none") +
  labs_meta(meta, y = "Share of International Enrollment (%)") +
  theme_tpa()

save_chart(fig_103, "chapter2_fig_103")
```

### Chapter 2 - Figure 104 - International Students in U.S. Higher Education, Share by Academic Level
```{r fig_104}
meta <- get_meta("fig_104")

df_104 <- read_fig("fig_104")

fig_104 <- ggplot() +
  block_arrow_layers(df_104, Category, Change,
                     fills = arrow_fill,
                     divider_after = 1) +
  labs_meta(meta) +
  theme_arrows()

save_chart(fig_104, "chapter2_fig_104", size = "standard")
```

### Chapter 2 - Figure 105 - Perspectives of Increases in International Enrollment
```{r fig_105}
meta <- get_meta("fig_105")

df_105 <- read_fig("fig_105") %>%
  rename(Reason = 1) %>%
  mutate(across(-Reason, as.numeric)) %>%
  pivot_longer(-Reason, names_to = "Cohort", values_to = "Share") %>%
  filter(!is.na(Share)) %>%
  mutate(
    Cohort = factor(Cohort, levels = c("Fall 2018", "Fall 2021", "Fall 2025")),
    Reason = str_wrap(Reason, width = 50),                 # <- lower width = narrower text column
    Reason = factor(Reason, levels = rev(unique(Reason)))
  )

cohort_colors <- c(
  "Fall 2018" = tpa_colors[1],
  "Fall 2021" = tpa_colors[2],
  "Fall 2025" = tpa_colors[3]
)

# Spread crowded value labels horizontally; gap is in percentage points.
gap <- 4
labels_105 <- df_105 %>%
  group_by(Reason) %>%
  arrange(Share, .by_group = TRUE) %>%
  mutate(
    spread  = accumulate(Share, ~ max(.y, .x + gap)),
    label_x = spread - (mean(spread) - mean(Share))
  ) %>%
  ungroup()

fig_105 <- ggplot(df_105, aes(Share, Reason, color = Cohort)) +
  geom_point(size = 3.5) +
  geom_text(
    data = labels_105,
    aes(x = label_x, label = paste0(Share, "%")),
    nudge_y = -0.34,
    size = (DATA_LABEL_SIZE - 2) / .pt,
    fontface = "bold",
    family = FONT_FAMILY,
    show.legend = FALSE
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.07, 0.07))) +
  scale_y_discrete(expand = expansion(add = c(0.8, 0.6))) +
  scale_color_manual(values = cohort_colors) +
  guides(color = guide_legend(override.aes = list(size = 3.5))) +
  coord_cartesian(clip = "off") +
  labs_meta(meta) +
  theme_dotplot()

save_chart(fig_105, "chapter2_fig_105", size="standard")
```

### Chapter 2 - Figure 106 - Perspectives of Decreases in International Enrollment
```{r fig_106}
meta <- get_meta("fig_106")

df_106 <- read_fig("fig_106") %>%
  rename(Reason = 1) %>%
  mutate(across(-Reason, as.numeric)) %>%
  pivot_longer(-Reason, names_to = "Cohort", values_to = "Share") %>%
  filter(!is.na(Share)) %>%
  mutate(
    Cohort = factor(Cohort, levels = c("Fall 2018", "Fall 2021", "Fall 2025")),
    Reason = str_wrap(Reason, width = 50),                 # <- lower width = narrower text column
    Reason = factor(Reason, levels = rev(unique(Reason)))
  )

cohort_colors <- c(
  "Fall 2018" = tpa_colors[1],
  "Fall 2021" = tpa_colors[2],
  "Fall 2025" = tpa_colors[3]
)

# Spread crowded value labels horizontally; gap is in percentage points.
gap <- 4
labels_106 <- df_106 %>%
  group_by(Reason) %>%
  arrange(Share, .by_group = TRUE) %>%
  mutate(
    spread  = accumulate(Share, ~ max(.y, .x + gap)),
    label_x = spread - (mean(spread) - mean(Share))
  ) %>%
  ungroup()

fig_106 <- ggplot(df_106, aes(Share, Reason, color = Cohort)) +
  geom_point(size = 3.5) +
  geom_text(
    data = labels_106,
    aes(x = label_x, label = paste0(Share, "%")),
    nudge_y = -0.34,
    size = (DATA_LABEL_SIZE - 2) / .pt,
    fontface = "bold",
    family = FONT_FAMILY,
    show.legend = FALSE
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.07, 0.07))) +
  scale_y_discrete(expand = expansion(add = c(0.8, 0.6))) +
  scale_color_manual(values = cohort_colors) +
  guides(color = guide_legend(override.aes = list(size = 3.5))) +
  coord_cartesian(clip = "off") +
  labs_meta(meta) +
  theme_dotplot()

save_chart(fig_106, "chapter2_fig_106", size="standard")
```

### Chapter 2 - Figure 107 - Institutional Undergraduate and Graduate Recruitment Priorities, Fall 2025
```{r fig_107}
meta <- get_meta("fig_107")

panel_colors <- c(
  "Undergraduate" = tpa_colors[1],   # gold
  "Graduate"      = tpa_colors[2]     # teal (not in tpa_colors; swap for tpa_colors[2] navy for house palette)
)

track_max <- 1.0   # grey track length; nudge up toward ~1.1 if fills look too long

df_107 <- read_fig("fig_107") %>%
  transmute(
    Rank                  = as.integer(str_remove(Rank, "Rank ")),
    Undergraduate_Country = `Undergraduate Country`,
    Undergraduate_Value   = as.numeric(`Undergraduate Percent`),
    Graduate_Country      = `Graduate Country`,
    Graduate_Value        = as.numeric(`Graduate Percent`)
  ) %>%
  pivot_longer(-Rank, names_to = c("Panel", ".value"), names_sep = "_") %>%
  mutate(
    Panel   = factor(Panel, levels = c("Undergraduate", "Graduate")),
    Country = str_wrap(Country, width = 18),
    Country = reorder_within(Country, -Rank, Panel)   # rank 1 on top, per panel
  )

fig_107 <- ggplot(df_107, aes(y = Country)) +
  geom_col(aes(x = track_max), fill = "grey90", width = 0.65) +
  geom_col(aes(x = Value, fill = Panel), width = 0.65) +
  geom_text(aes(x = track_max, label = label_percent(accuracy = 1)(Value)),
            hjust = -0.25, fontface = "bold", color = "grey20",
            size = DATA_LABEL_SIZE / .pt, family = FONT_FAMILY) +
  facet_wrap(~ Panel, scales = "free_y") +
  scale_x_continuous(expand = expansion(mult = c(0, 0.14))) +
  scale_y_reordered() +
  scale_fill_manual(values = panel_colors) +
  coord_cartesian(clip = "off") +
  labs_meta(meta) +
  theme_bullet()

save_chart(fig_107, "chapter2_fig_107")
```
### Chapter 2 - Figure 108 - Top 10 countries by international undergraduate enrollment (count)
```{r 108}
meta <- get_meta("fig_108")

df_108 <- read_fig("fig_108") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Country", values_to = "Count") %>%
  filter(!is.na(Count))

top_countries <- df_108 %>%
  filter(Year == max(Year)) %>%
  slice_max(Count, n = 8) %>%
  pull(Country)

df_108 <- df_108 %>%
  filter(Country %in% top_countries) %>%
  order_by_latest(Country, Count)

labels_108 <- last_by_group(df_108, Year, Country)

fig_108 <- ggplot(df_108, aes(Year, Count, color = Country, group = Country)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_label_repel_tpa(
    data         = labels_108,
    mapping      = aes(label = paste0(Country, ": ", label_comma()(Count))),
    nudge_x      = 0.5,
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_108$Year, pad_right = .3) +
  scale_y_continuous(labels = label_comma(),
                breaks = c(1000, 5000, 10000, 50000, 100000, 250000),
                expand = EXPAND_LINE) +
  scale_color_tpa() +
  guides(color = "none") +
  labs_meta(meta, y = "Enrolled International Students") +
  theme_tpa()

save_chart(fig_108, "chapter2_fig_108")
```

### Chapter 2 - Figure 109 - Top 10 countries by international undergraduate enrollment (share)
```{r 109}
meta <- get_meta("fig_109")

df_109 <- read_fig("fig_109") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Country", values_to = "Count") %>%
  filter(!is.na(Count))

top_countries <- df_109 %>%
  filter(Year == max(Year)) %>%
  slice_max(Count, n = 8) %>%
  pull(Country)

df_109 <- df_109 %>%
  filter(Country %in% top_countries) %>%
  order_by_latest(Country, Count)

labels_109 <- last_by_group(df_109, Year, Country)

fig_109 <- ggplot(df_109, aes(Year, Count, color = Country, group = Country)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_label_repel_tpa(
    data         = labels_109,
    mapping      = aes(label = paste0(Country, ": ", label_percent(accuracy = 0.1)(Count))),
    nudge_x      = 0.5,
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_109$Year, pad_right = .3) +
  scale_y_continuous(labels = label_percent(accuracy = 1),
                expand = EXPAND_LINE) +
  scale_color_tpa() +
  guides(color = "none") +
  labs_meta(meta, y = "Share of Enrolled International Students") +
  theme_tpa()

save_chart(fig_109, "chapter2_fig_109")
```

### Chapter 2 - Figure 110 - Top 10 countries by international graduate enrollment (count)
```{r 110}
meta <- get_meta("fig_110")

df_110 <- read_fig("fig_110") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Country", values_to = "Count") %>%
  filter(!is.na(Count))

top_countries <- df_110 %>%
  filter(Year == max(Year)) %>%
  slice_max(Count, n = 8) %>%
  pull(Country)

df_110 <- df_110 %>%
  filter(Country %in% top_countries) %>%
  order_by_latest(Country, Count)

labels_110 <- last_by_group(df_110, Year, Country)

fig_110 <- ggplot(df_110, aes(Year, Count, color = Country, group = Country)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_label_repel_tpa(
    data         = labels_110,
    mapping      = aes(label = paste0(Country, ": ", label_comma()(Count))),
    nudge_x      = 0.5,
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_110$Year, pad_right = .3) +
  scale_y_continuous(labels = label_comma(),
                expand = EXPAND_LINE) +
  scale_color_tpa() +
  guides(color = "none") +
  labs_meta(meta, y = "Enrolled International Students") +
  theme_tpa()

save_chart(fig_110, "chapter2_fig_110")
```

# Chapter 3
### Chapter 3 - Figure 30 - Top 15 Employers with NAICS Code 54 H-1B Visa Approvals
```{r figure 030}
# ---- shared fig_030 prep ----
# Full firm-level H-1B approvals (NAICS 54). Computes shares against
# the annual Grand Total row, excludes the partial 2026 column.
fig030_shares <- function() {
  raw <- read_fig("fig_030")
  names(raw) <- trimws(names(raw))

  total_row <- raw %>% filter(trimws(Company) == "Grand Total")
  totals <- total_row %>%
    select(-Company, -`Grand Total`) %>%
    pivot_longer(everything(), names_to = "Year", values_to = "Total") %>%
    mutate(Year = as.integer(Year), Total = as.numeric(Total))

  raw %>%
    filter(!trimws(Company) %in% c("Grand Total")) %>%
    mutate(Company = trimws(Company)) %>%
    select(-`Grand Total`) %>%
    pivot_longer(-Company, names_to = "Year", values_to = "Count") %>%
    mutate(Year = as.integer(Year),
           Count = as.numeric(replace_na(Count, 0))) %>%
    filter(Year != 2026) %>%                     # drop partial year
    left_join(totals, by = "Year") %>%
    mutate(Share = Count / Total * 100)
}

meta <- get_meta("fig_030")

short_names <- c(
  "COGNIZANT TECHNOLOGY SOLUTIONS US CORP"      = "Cognizant",
  "TATA CONSULTANCY SERVICES LIMITED"           = "Tata",
  "INFOSYS LIMITED"                             = "Infosys",
  "GOOGLE INC"                                  = "Google",
  "Deloitte and Subsidiaries"                   = "Deloitte",
  "WIPRO LIMITED"                               = "WIPRO",
  "ERNST & YOUNG LLP"                           = "Ernst & Young",
  "TECH MAHINDRA AMERICAS INC"                  = "Tech Mahindra",
  "HCL Technologies Limited and Subsidiaries"   = "HCL",
  "CAPGEMINI AMERICA INC"                       = "Capgemini",
  "LTIMINDTREE LIMITED"                         = "LTIMindtree",
  "IBM and subsidiaries"                        = "IBM",
  "LARSEN & TOUBRO INFOTECH LIMITED"            = "Larsen & Toubro"
)

shares <- fig030_shares()

# rank union firms by their max share across the two years, so the
# biggest bars (Cognizant's 2017, Google's 2025) sit on the left
df_bar <- shares %>%
  filter(Year %in% c(2017, 2025), Company %in% names(short_names)) %>%
  mutate(Company = recode(Company, !!!short_names),
         Year = factor(Year))

bar_order <- df_bar %>%
  group_by(Company) %>% summarise(m = max(Share)) %>%
  arrange(desc(m)) %>% pull(Company)

df_bar <- df_bar %>% mutate(Company = factor(Company, levels = bar_order))

fig_030a <- ggplot(df_bar, aes(Company, Share, fill = Year)) +
  geom_col(position = position_dodge(width = BAR_WIDTH),
           width = BAR_WIDTH, alpha = ALPHA_COL) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(0, 24), breaks = seq(0, 20, 5),
                     expand = EXPAND_BAR) +
  scale_fill_manual(values = c("2017" = tpa_colors[2], "2025" = tpa_colors[1])) +
  labs_meta(meta, y = "Share of H-1B approvals (NAICS 54)") +
  theme_tpa() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

save_chart(fig_030a, "chapter3_fig_030")
```

### Chapter 3 - Figure 31 - The Evolving Founder Demographics of the US-Based Forbes AI 50 Companies (2019–2025)
```{r figure 031}
meta <- get_meta("fig_031")

df_031 <- read_fig("fig_031") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Composition", values_to = "Share") %>%
  filter(!is.na(Share)) %>%
  mutate(
    Share       = Share * 100,
    Composition = case_when(
      Composition == "all us"       ~ "All U.S.",
      Composition == "all foreign"  ~ "All foreign",
      Composition == "some foreign" ~ "Some foreign"
    )
  ) %>%
  order_by_latest(Composition, Share)

palette_031 <- c(
  "All U.S."     = tpa_colors[1],
  "All foreign"  = tpa_colors[2],
  "Some foreign" = tpa_colors[4]
)

labels_031 <- last_by_group(df_031, Year, Composition)

fig_031 <- ggplot(df_031, aes(Year, Share, color = Composition, group = Composition)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_point(size = POINT_SIZE) +
  geom_label_repel_tpa(
    data         = labels_031,
    mapping      = aes(label = paste0(Composition, ": ", fmt_pct_1(Share))),
    nudge_x      = 0.5,
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_031$Year, by = 1, pad_right = PAD_RIGHT_STANDARD) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(20, 50),
                     breaks = seq(0, 50, 10),
                     expand = EXPAND_BAR_TIGHT) +
  scale_color_manual(values = palette_031) +
  guides(color = "none") +
  labs_meta(meta, y = "Share of Forbes AI 50") +
  theme_tpa()

save_chart(fig_031, "chapter3_fig_031")
```

### Chapter 3 - Figure 32 - Countries of Origin for Foreign Founders of U.S.-Based AI Companies Ranked in the Forbes AI5o (2019-2025)
```{r figure 032}
meta <- get_meta("fig_032")

df_032 <- read_fig("fig_032") %>%
  rename(Country = 1, Count = 2) %>%
  mutate(
    Count = as.numeric(Count),
    Country = case_when(
      Country == "the Netherlands"     ~ "Netherlands",
      Country == "United Arab Emirates" ~ "United Arab Emirates",
      Country == "Scotland"             ~ "United Kingdom",
      Country == "Hong Kong"            ~ "Hong Kong S.A.R.",
      TRUE                              ~ Country
    )
  ) %>%
  group_by(Country) %>%
  summarise(Count = sum(Count, na.rm = TRUE), .groups = "drop")

world_032 <- join_country_data(df_032)

fig_032 <- ggplot(world_032) +
  geom_sf(aes(fill = Count), color = "white", linewidth = 0.2) +
  geom_country_labels(world_032, value_col = "Count", min_value = 5) +
  scale_fill_gradient(
    low = TPA_RED_LIGHT,
    high = tpa_colors[1],
    na.value = "grey92",
    breaks = c(1, 20, 40, 60, 76),
    labels = label_comma(),
    name = "Founders"
  ) +
  coord_sf(crs = "+proj=robin", expand = FALSE) +
  labs_meta(meta) +
  theme_map()

save_chart(fig_032, "chapter3_fig_032")
```

### Chapter 3 - Figure 33 - Founder Composition and Valuation of Forbes AI50 Startups
```{r figure 033}
### Figure 33 - Founder Composition and Valuation of Forbes AI50 Startups
fig033_data <- function() {
  raw <- read_fig("fig_033")
  names(raw) <- make.unique(names(raw))   # de-dupe the repeated headers
  foreign <- raw %>%
    transmute(Company,
              Valuation = as.numeric(`Pitchbook Valuation`),
              Origin = "Foreign-founded") %>%
    filter(!is.na(Valuation))
  domestic <- raw %>%
    transmute(Company,
              Valuation = as.numeric(`Pitchbook Valuation.1`),
              Origin = "Domestic-founded") %>%
    filter(!is.na(Valuation))
  bind_rows(foreign, domestic) %>%
    mutate(ValB = Valuation / 1e9)
}

meta <- get_meta("fig_033")

df_top <- fig033_data() %>%
  slice_max(ValB, n = 25) %>%
  mutate(Company = fct_reorder(Company, ValB))

fig_033 <- ggplot(df_top, aes(ValB, Company, color = Origin)) +
  geom_point(size = 3.2) +
  scale_x_log10(labels = label_dollar(suffix = "B", accuracy = 1),
                breaks = c(1, 3, 10, 30, 100, 300),
                expand = expansion(mult = c(0.04, 0.08))) +
  scale_color_manual(values = c("Foreign-founded"  = "#C46C5D",
                                "Domestic-founded" = "#A6C5DE")) +
  labs_meta(meta, y = NULL) +
  labs(x = "Valuation (log scale)") +
  theme_dotplot() +
  theme(
    panel.grid.major.y = element_line(color = "grey92", linewidth = 0.3),
    # theme_dotplot blanks the x-axis (value is normally on each dot);
    # this chart has no dot labels, so restore the axis and title it
    axis.text.x  = element_text(family = FONT_FAMILY, color = "grey25",
                                size = AXIS_SIZE, angle = 0, hjust = 0.5,
                                margin = margin(t = 4)),
    axis.ticks.x = element_line(color = "grey50", linewidth = 0.35),
    axis.line.x  = element_line(color = "grey50", linewidth = 0.4),
    axis.title.x = element_text(family = FONT_FAMILY, color = "grey25",
                                size = AXIS_SIZE, margin = margin(t = 8))
  )

save_chart(fig_033, "chapter3_fig_033")
```

### Chapter 3 - Figure 34 - Geographic Affiliation Distribution of Research Paper Authors from Top AI Companies (2025)
```{r figure 034}
meta <- get_meta("fig_034")

company_order <- c("Alphabet","Anthropic", "OpenAI", "Alibaba", "DeepSeek", "Moonshot")
category_order <- c("Rest of World", "China", "United States")

df_034 <- read_fig("fig_034") %>%
  rename(Company = 1) %>%
  pivot_longer(-Company, names_to = "Affiliation", values_to = "Share") %>%
  mutate(
    Share = as.numeric(Share) * 100,
    Affiliation = str_remove(Affiliation, "^Current Affiliation - ?"),
    Affiliation = factor(Affiliation, levels = category_order),
    Company = factor(Company, levels = company_order)
  )

palette_034 <- c(
  "United States"      = tpa_colors[2],
  "China"              = tpa_colors[1],
  "Rest of World"      = tpa_colors[3]
)

fig_034 <- ggplot(df_034, aes(Share, Company, fill = Affiliation)) +
  geom_col(width = BAR_WIDTH_NARROW, alpha = ALPHA_COL) +
  geom_text(aes(label = fmt_pct_0(Share)),
            position = position_stack(vjust = 0.5),
            color = "white",
            fontface = "bold",
            family = FONT_FAMILY,
            size = DATA_LABEL_SIZE / .pt) +
  scale_x_continuous(labels = label_number(suffix = "%"),
                     breaks = seq(0, 100, 10),
                     limits = c(0, 101),
                     expand = expansion(mult = c(0, 0))) +
  scale_fill_manual(values = palette_034,
                    breaks = c("United States", "China", "Rest of World")) +
  labs_meta(meta, x = "Share of researchers") +
  theme_horizontal_bar()

save_chart(fig_034, "chapter3_fig_034")
```
### Chapter 3 - Figure 115 - Who Actually Works in STEM After Earning a STEM Degree?
```{r}
# shared prep -- combine US-born (fig_027) and foreign-born (fig_028)
# into one long frame with a Population column
prep_conversion <- function() {
  read_one <- function(name, pop) {
    read_fig(name) %>%
      mutate(across(-Year, as.numeric)) %>%
      pivot_longer(-Year, names_to = "Degree", values_to = "Rate") %>%
      mutate(Rate = Rate * 100, Population = pop)
  }

  bind_rows(
    read_one("fig_027", "U.S.-born"),
    read_one("fig_028", "Foreign-born")
  ) %>%
    mutate(
      Degree = recode(Degree,
        "bachelors"    = "Bachelors",
        "masters"      = "Masters",
        "doctorates"   = "Doctorates",
        "doctors"      = "Doctorates",
        "professional" = "Professional"),
      Degree = factor(Degree, levels = c("Doctorates", "Professional", "Masters", "Bachelors"))
    )
}

meta <- get_meta("fig_115")

first_yr <- 2003
last_yr  <- 2023

# long form: one row per Degree × Population × Year, with Year as a factor
# so it can drive the open/filled shape legend
df_db <- prep_conversion() %>%
  filter(Year %in% c(first_yr, last_yr)) %>%
  mutate(Year = factor(Year, levels = c(first_yr, last_yr)))

# wide, just for the connecting segments
df_seg <- df_db %>%
  select(Degree, Population, Year, Rate) %>%
  pivot_wider(names_from = Year, values_from = Rate, names_prefix = "Y")

fig_db <- ggplot(df_db, aes(y = Degree)) +
  geom_segment(data = df_seg,
               aes(x = Y2003, xend = Y2023, yend = Degree, color = Population),
               linewidth = 1.1, alpha = 0.55) +
  # both years drawn as shape 21; Year drives fill via the shape legend
  geom_point(aes(x = Rate, color = Population, shape = Year),
             fill = "white", stroke = 1.2, size = 3.8) +
  geom_point(data = filter(df_db, Year == last_yr),
             aes(x = Rate, color = Population, fill = Population),
             shape = 21, stroke = 0, size = 4) +
  scale_x_continuous(labels = label_number(suffix = "%"),
                     limits = c(20, 85), breaks = seq(20, 80, 10),
                     expand = expansion(mult = c(0.05, 0.10))) +
  scale_color_manual(values = c("U.S.-born"    = tpa_colors[2],
                                "Foreign-born" = tpa_colors[1])) +
  scale_fill_manual(values = c("U.S.-born"    = tpa_colors[2],
                               "Foreign-born" = tpa_colors[1]),
                    guide = "none") +
  scale_shape_manual(values = c("2003" = 21, "2023" = 19),
                     labels = c("2003" = "2003 (open)", "2023" = "2023 (filled)")) +
  guides(
    color = guide_legend(order = 1, override.aes = list(shape = 19, size = 4)),
    shape = guide_legend(order = 2,
                         override.aes = list(color = "grey40", fill = "white"))
  ) +
  labs_meta(meta, y = NULL) +
  theme_horizontal_bar()

save_chart(fig_db, "chapter3_fig_115")
```

### Chapter 3 - Figure 116 - Top-10 H1B Visa Granted Firms Share Over time
```{r fig_116}
meta <- get_meta("fig_116")

# true top-10-by-year share: rank all firms each year, sum the top 10
df_116 <- fig030_shares() %>%
  group_by(Year) %>%
  slice_max(Count, n = 10) %>%
  summarise(value = sum(Share), .groups = "drop") %>%
  mutate(Year = as.numeric(Year))

labels <- first_last_peak(df_116, Year, value) %>%
  mutate(hj = dplyr::case_when(
    Year == min(df_116$Year) ~ -0.4,   # first point: label to the RIGHT (inward)
    Year == max(df_116$Year) ~  1.3,   # last point:  label to the LEFT (inward)
    TRUE                       ~  0.5    # peak: stays centered
  ))

fig_116 <- ggplot(df_116, aes(Year, value)) +
  geom_line(color = tpa_colors[1], linewidth = LINE_WIDTH) +
  geom_point(data = labels, color = tpa_colors[1], size = POINT_SIZE) +
  # peak label: house wrapper, centered above the point
  geom_text_tpa(data    = filter(labels, hj == .5),
                mapping = aes(label = sprintf("%.1f%%", value)),
                color   = tpa_colors[1],
                vjust   = 1.5) +
  # endpoint labels: nudged inward so they don't clip the panel edges
  geom_text(data    = filter(labels, hj != 0.5),
            mapping = aes(label = sprintf("%.1f%%", value), hjust = hj),
            color   = tpa_colors[1], fontface = "bold",
            family  = FONT_FAMILY, size = DATA_LABEL_SIZE / .pt) +
  scale_x_years(df_116$Year, by = 1) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(50, 75),
                     breaks = seq(50, 75, 5),
                     expand = EXPAND_BAR_TALL) +
  labs_meta(meta, y = "Top 10 employers' share of approvals") +
  theme_tpa()

save_chart(fig_116, "chapter3_fig_116")
```

### Chapter 3 - Figure 117 - How Does Founder Origin Vary Across AI Startup Valuations?
```{r fig_117}
meta <- get_meta("fig_117")

tier_levels  <- c("<$1B", "$1–10B", "$10–100B", "≥$100B")
origin_colors <- c("Foreign-founded"  = tpa_colors[1],
                   "Domestic-founded" = tpa_colors[2])

df_tier <- read_fig("fig_117") %>%
  rename(Origin = Binary) %>%
  mutate(Origin = recode(Origin,
    "Has All U.S. Born Founders"            = "Domestic-founded",
    "Has At Least One Foreign-Born Founder" = "Foreign-founded")) %>%
  pivot_longer(-Origin, names_to = "Tier", values_to = "n") %>%
  mutate(
    n    = replace_na(as.numeric(n), 0),          # blank ≥$100B cell -> 0
    Tier = factor(Tier, levels = tier_levels),
    Origin = factor(Origin, levels = c("Domestic-founded", "Foreign-founded"))
  )

fig_117 <- ggplot(df_tier, aes(Tier, n, fill = Origin)) +
  geom_col(position = position_dodge(width = BAR_WIDTH),
           width = BAR_WIDTH, alpha = ALPHA_COL) +
  geom_text(aes(label = n, color = Origin),
            position = position_dodge(width = BAR_WIDTH),
            vjust = -0.4, size = DATA_LABEL_SIZE / .pt,
            fontface = "bold", show.legend = FALSE) +
  scale_y_continuous(expand = EXPAND_BAR_TALL) +
  scale_fill_manual(values = origin_colors) +
  scale_color_manual(values = origin_colors) +
  labs_meta(meta, y = "Number of companies") +
  theme_tpa() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5))

save_chart(fig_117, "chapter3_fig_117")
```

# Chapter 4

### Chapter 4 - Figure 35 - US PISA Scores Compared to China and OECD Average
```{r figure 035}
meta <- get_meta("fig_035")

subject_order <- c("Reading", "Science", "Math")   # Math last -> drawn on top

df_035 <- read_fig("fig_035") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Subject", values_to = "Score") %>%
  filter(!is.na(Score)) %>%
  mutate(Subject = factor(Subject, levels = subject_order))

palette_035 <- c(
  "Math"    = tpa_colors[1],   # red accent -- the chapter's argument
  "Science" = "grey62",
  "Reading" = "grey78"
)

df_math  <- filter(df_035, Subject == "Math")
df_muted <- filter(df_035, Subject != "Subject" & Subject != "Math")

# point-level labels: every Math point, muted series only at endpoints
math_labels  <- df_math
muted_labels <- df_muted %>% filter(Year %in% range(Year))

# right-edge series labels replacing the legend
end_labels <- last_by_group(df_035, Year, Subject)

fig_035 <- ggplot(df_035, aes(Year, Score, color = Subject, group = Subject)) +
  geom_hline(yintercept = 0, color = "grey50", linewidth = LINE_WIDTH_REF) +
  geom_line(data = df_muted, linewidth = LINE_WIDTH) +
  geom_point(data = df_muted, size = 1.8) +
  geom_line(data = df_math, linewidth = LINE_WIDTH + 0.4) +
  geom_point(data = df_math, size = 2.4) +
  geom_text(data = math_labels, aes(label = sprintf("%+d", Score)),
            vjust = 1.9, size = DATA_LABEL_SIZE / .pt,
            fontface = "bold", family = FONT_FAMILY, show.legend = FALSE) +
  geom_text(data = muted_labels, aes(label = sprintf("%+d", Score)),
            vjust = -1.1, size = DATA_LABEL_SIZE / .pt,
            fontface = "bold", family = FONT_FAMILY, show.legend = FALSE) +
  geom_label_repel_tpa(
    data    = end_labels,
    mapping = aes(label = Subject),
    nudge_x = 0.4, hjust = 0, direction = "y"
  ) +
  annotate("text", x = min(df_035$Year), y = 1.6, label = "OECD average",
           hjust = 0, vjust = 0, color = "grey50",
           size = (AXIS_SIZE - 2) / .pt, family = FONT_FAMILY) +
  scale_x_continuous(breaks = unique(df_035$Year),
                     expand = expansion(mult = c(0.03, 0.12))) +
  scale_y_continuous(limits = c(-20, 32), breaks = seq(-20, 30, 5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  scale_color_manual(values = palette_035) +
  guides(color = "none") +
  labs_meta(meta, y = "Points above or below OECD average") +
  theme_tpa() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5))

save_chart(fig_035, "chapter4_fig_035")
```

### Chapter 4 - Figure 36 - K-12 Education Expenditure - % of GDP and per pupil
```{r figure 036}
meta <- get_meta("fig_036")

subset_a <- c("United States", "OECD average", "Korea", "Japan", "Estonia",
              "Poland", "Finland", "Germany", "Norway", "Austria",
              "Netherlands", "Canada", "United Kingdom", "Iceland")

num <- function(x) parse_number(as.character(x))   # handles numeric or "15,799.00"

ratio_ord <- read_fig("fig_036") %>%
  mutate(K12 = num(`K-12 (in USD)`),
         Tert = num(`Tertiary (in USD)`),
         r = Tert / K12) %>%
  filter(Country %in% subset_a) %>%
  arrange(r) %>%
  pull(Country)

df_036 <- read_fig("fig_036") %>%
  mutate(
    K12  = num(`K-12 (in USD)`),
    Tert = num(`Tertiary (in USD)`)
  ) %>%
  filter(Country %in% subset_a) %>%
  select(Country, Group, K12, Tert) %>%
  pivot_longer(c(K12, Tert), names_to = "Level", values_to = "Per_pupil") %>%
  mutate(
    Level    = factor(if_else(Level == "K12", "K-12", "Tertiary"),
                      levels = c("K-12", "Tertiary")),
    Country  = factor(Country, levels = ratio_ord),
    fill_key = paste(if_else(Group == "United States", "US", "Other"), Level)
  )

# US bars in the red family so the US pops on BOTH bars;
# everyone else in two grays so the K-12/tertiary distinction still reads
fill_vals <- c(
  "Other K-12"     = "grey78",
  "Other Tertiary" = "grey55",
  "US K-12"        = TPA_RED_MID,
  "US Tertiary"    = tpa_colors[1]
)

fig_036 <- ggplot(df_036, aes(Per_pupil, Country, fill = fill_key)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.65, alpha = ALPHA_COL) +
  scale_x_continuous(labels = label_dollar(scale = 1e-3, suffix = "K"),
                     expand = EXPAND_BAR) +
  scale_fill_manual(values = fill_vals,
                    breaks = c("Other K-12", "Other Tertiary"),
                    labels = c("K-12", "Tertiary")) +
  labs_meta(meta, x = "Expenditure per student (USD, PPP)") +
  theme_horizontal_bar()

save_chart(fig_036, "chapter4_fig_036", size = "standard")
```

### Chapter 4 - Figure 37 - US-trained Students – Undergraduate – STEM Major Persistence Rate – Three and Five-Year
```{r figure 037}
meta <- get_meta("fig_037")

keep <- c("Never switched", "Switched once", "Switched 2+ times")

df_037 <- read_fig("fig_037") %>%
  rename(Survey = 1) %>%
  pivot_longer(-Survey, names_to = "Category", values_to = "Share") %>%
  filter(Category %in% keep) %>%
  mutate(
    Share    = as.numeric(Share) * 100,
    Category = factor(Category,
                      levels = keep,
                      labels = c("Never switched", "Switched once", "Switched 2+ times")),
    Survey   = factor(Survey, levels = c("2004/2009 Survey", "2012/2017 Survey"))
  )

palette_037 <- c(
  "2004/2009 Survey" = tpa_colors[2],
  "2012/2017 Survey" = tpa_colors[1]
)

fig_037 <- ggplot(df_037, aes(Category, Share, fill = Survey)) +
  geom_col_grouped() +
  geom_text(aes(label = fmt_pct_1(Share)),
            position = pos_dodge_grouped(),
            vjust = -0.6,
            size = DATA_LABEL_SIZE / .pt,
            fontface = "bold",
            family = FONT_FAMILY,
            color = "grey25") +
  scale_x_discrete(labels = function(x) str_wrap(x, 14)) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(0, 70),
                     breaks = seq(0, 70, 10),
                     expand = EXPAND_TIGHT_TOP) +
  scale_fill_manual(values = palette_037) +
  labs_meta(meta, y = "Share of students") +
  theme_tpa() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5))

save_chart(fig_037, "chapter4_fig_037")
```

### Chapter 4 - Figure 38 - STEM Graduation Outcomes by Initially Declared Major, All Cohorts
```{r figure 038}
meta <- get_meta("fig_038a")

outcome_levels <- c(
  "stem_bachelors"    = "Bachelor's in STEM",
  "nonstem_bachelors" = "Bachelor's in another field",
  "subbac_credential" = "Associate's or certificate",
  "still_enrolled"    = "Still enrolled, no degree",
  "left_no_degree"    = "Left without a degree"
)

df_038a <- read_fig("fig_038a") %>%
  mutate(across(-cohort, as.numeric)) %>%
  pivot_longer(-cohort, names_to = "outcome", values_to = "pct") %>%
  mutate(outcome = factor(recode(outcome, !!!outcome_levels),
                          levels = unname(outcome_levels)))

fig_038a <- ggplot(df_038a, aes(x = cohort, y = pct, fill = outcome)) +
  geom_col(width = 0.62) +
  geom_text(aes(label = fmt_pct_1(pct)),
            position = position_stack(vjust = 0.5),
            family = FONT_FAMILY, size = 3, color = "white") +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(0, 100), breaks = seq(0, 100, 25),
                     expand = EXPAND_BAR_TIGHT) +
  scale_fill_tpa() +   # if only scale_color_tpa exists: scale_fill_manual(values = tpa_colors)
  labs_meta(meta, y = "Share of STEM entrants") +
  theme_tpa() +
  theme(axis.title.y = element_text(family = FONT_FAMILY, color = "grey25",
                                    size = AXIS_SIZE, margin = margin(r = 8)))

save_chart(fig_038a, "chapter4_fig_038a")
```

### Chapter 4 - Figure 38b - STEM Graduation Outcomes by Initially Declared Major, All Cohorts
```{r figure 038b}
meta <- get_meta("fig_038b")

field_order <- c("Computer & information sciences", "Engineering",
                 "Bio/physical sciences, math & ag")
col_order   <- c(field_order, "Non-STEM field")

df_038b <- read_fig("fig_038b") %>%
  mutate(pct = as.numeric(pct),
         entry_field  = factor(entry_field, levels = rev(field_order)),  # top row = CIS
         degree_field = factor(degree_field, levels = col_order),
         cohort       = factor(cohort, levels = c("2003/04", "2011/12")),
         is_diag      = as.character(entry_field) == as.character(degree_field),
         lab = case_when(
           is.na(pct)                        ~ "n/a",
           reliability %in% c("unstable_30_50", "unstable_50plus") ~ paste0(fmt_pct_1(pct), "*"),
           TRUE                              ~ fmt_pct_1(pct)))

fig_038b <- ggplot(df_038b, aes(degree_field, entry_field, fill = pct)) +
  geom_tile(color = "white", linewidth = 1.5) +
  geom_tile(data = ~filter(.x, is_diag),
            color = "grey15", linewidth = 1, fill = NA) +
  geom_text(aes(label = lab, color = pct > 40),
            family = FONT_FAMILY, size = 3) +
  scale_fill_gradient(low = "#FCEBEB", high = tpa_colors[1],  # or "#791F1F"
                      na.value = "grey92", guide = "none") +
  scale_color_manual(values = c(`TRUE` = "#FCEBEB", `FALSE` = "#501313"),
                     na.value = "grey45", guide = "none") +
  facet_wrap(~cohort) +
  coord_fixed() +
  labs_meta(meta, x = "Graduated in", y = "Entered") +
  theme_tpa() +
  theme(axis.title.x = element_text(family = FONT_FAMILY, color = "grey25",
                                    size = AXIS_SIZE, margin = margin(t = 8)),
        axis.title.y = element_text(family = FONT_FAMILY, color = "grey25",
                                    size = AXIS_SIZE, margin = margin(r = 8)),
        axis.text.x  = element_text(angle = 30, hjust = 1),
        panel.grid   = element_blank())

save_chart(fig_038b, "chapter4_fig_038b")
```

### Chapter 4 - Figure 38c - STEM Graduation Outcomes by Initially Declared Major, All Cohorts
```{r figure 038c}
meta <- get_meta("fig_038c")

df_038c <- read_fig("fig_038c") %>%
  mutate(across(-field, as.numeric)) %>%
  pivot_longer(-field, names_to = "cohort", values_to = "pct") %>%
  mutate(cohort = recode(cohort,
                         entry_2003 = "2003/04 cohort",
                         entry_2011 = "2011/12 cohort"),
         field = factor(field, levels = c(
           "Computer & information sciences",
           "Engineering",
           "Bio/physical sciences, math & ag")))

fig_038c <- ggplot(df_038c, aes(x = field, y = pct, fill = cohort)) +
  geom_col(width = 0.6, position = position_dodge(width = 0.65)) +
  geom_text(aes(label = fmt_pct_1(pct)),
            position = position_dodge(width = 0.65), vjust = -0.4,
            family = FONT_FAMILY, size = 3.2) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(0, 80), breaks = seq(0, 80, 20),
                     expand = EXPAND_BAR_TIGHT) +
  scale_fill_tpa() +
  labs_meta(meta, y = "Graduated in entry field") +
  theme_tpa() +
  theme(axis.title.y = element_text(family = FONT_FAMILY, color = "grey25",
                                    size = AXIS_SIZE, margin = margin(r = 8)))

save_chart(fig_038c, "chapter4_fig_038c")
```

### Chapter 4 - Figure 39 - Trends in U.S. Students Studying Abroad (For Academic Credit), by Year (1989/90-2022/2023)
```{r figure 039}
meta <- get_meta("fig_039")

df_039 <- read_fig("fig_039") %>%
  mutate(across(-Year, as.numeric)) %>%
  mutate(Share = 100 * `U.S. Study Abroad` / `Total US Enrollment`)

lab_039 <- df_039 %>% slice_max(Year, n = 1) %>%
  mutate(label = paste0("Share of enrollment: ", fmt_pct_1(Share)))

fig_039 <- ggplot(df_039, aes(Year, Share)) +
  geom_line(linewidth = LINE_WIDTH_DENSE, color = tpa_colors[1]) +
  geom_label_repel_tpa(data = lab_039, aes(label = label),
                       nudge_x = 0.5, hjust = 0, direction = "y",
                       with_segment = TRUE, color = tpa_colors[1]) +
  annotate("text", x = 2021, y = 0.09, label = "COVID-19",
           family = FONT_FAMILY, size = 3, color = "grey40", vjust = 1.8) +
  scale_x_years(df_039$Year, by = 2, pad_right = .5) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(0, 2.5), breaks = seq(0, 2.5, 0.5),
                     expand = EXPAND_BAR_TIGHT) +
  labs_meta(meta, y = "Share of total U.S. enrollment") +
  theme_tpa()

save_chart(fig_039, "chapter4_fig_039")
```

### Chapter 4 - Figure 40 - Distribution of Field of Study in U.S. Study Abroad Students (2000/01-2023/24)
```{r figure 040}
meta <- get_meta("fig_040")
df   <- read_country_lines("fig_040", group_col = "Field", value_col = "Share")
fig_040  <- plot_country_lines(df, meta, "Share of U.S. students abroad",
                           group_col = "Field",
                           value_col = "Share",
                           label_fmt = fmt_pct_1,
                           x_by      = 2,
                           pad_right = .6,
                           y_args    = list(
                             labels = label_number(suffix = "%"),
                             limits = c(0, 35),
                             breaks = seq(0, 35, 5),
                             expand = EXPAND_BAR_TIGHT
                           ))
save_chart(fig_040, "chapter4_fig_040")
```

### Chapter 4 - Figure 41 - Distribution of Host Regions of U.S. Students Studying Abroad (For Academic Credit), by Year and By World Region (2004/05-2022/23)
```{r figure 041}
meta <- get_meta("fig_041")
df   <- read_country_lines("fig_041", group_col = "Region", value_col = "Share") %>%
  filter(Region != "Antarctica") %>%
  mutate(Region = recode(Region,
    "Latin America & Caribbean"  = "Latin America",
    "Middle East & North Africa" = "MENA",
    "Africa, Sub-Saharan"        = "Sub-Saharan Africa",
    "Multiple Destinations"      = "Multiple"))

fig <- plot_country_lines(df, meta, "Share of U.S. students abroad",
                          group_col = "Region",
                          value_col = "Share",
                          label_fmt = fmt_pct_1,
                          x_by      = 2,
                          pad_right = .4,
                          y_args    = list(
                            labels = label_number(suffix = "%"),
                            limits = c(0, 80),
                            breaks = seq(0, 80, 10),
                            expand = EXPAND_BAR_TIGHT
                          ))
save_chart(fig, "chapter4_fig_041")
```

### Chapter 4 - Figure 42 - Net flow of internationally mobile students (inbound - outbound) (UIS estimate)
```{r figure 042}
meta <- get_meta("fig_042")

df_042 <- read_fig("fig_042") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Country", values_to = "Flow") %>%
  mutate(Flow = Flow / 1e6) %>%              # keep NA rows so lines break at gaps
  order_by_latest(Country, Flow)

highlighted <- c("United States of America", "China")

df_042 <- df_042 %>% mutate(is_hi = Country %in% highlighted)

labels_042 <- df_042 %>%
  filter(!is.na(Flow), Country %in% highlighted) %>%
  group_by(Country) %>% slice_max(Year, n = 1) %>% ungroup() %>%
  mutate(label = paste0(if_else(Country == "United States of America",
                                "United States", Country),
                        ": ", sprintf("%.2fM", Flow)))

fig_042 <- ggplot() +
  geom_hline(yintercept = 0, color = "grey50", linewidth = LINE_WIDTH_REF) +
  geom_line(data = filter(df_042, !is_hi),
            aes(Year, Flow, group = Country),
            color = "grey80", linewidth = 0.35, na.rm = FALSE) +
  geom_line(data = filter(df_042, is_hi),
            aes(Year, Flow, color = Country, group = Country),
            linewidth = LINE_WIDTH_DENSE, na.rm = FALSE) +
  geom_label_repel_tpa(data = labels_042, aes(Year, Flow, label = label, color = Country),
                       nudge_x = 0.5, hjust = 0, direction = "y", with_segment = TRUE) +
  scale_x_years(df_042$Year, by = 2, pad_right = .3) +
  scale_y_continuous(labels = function(x) paste0(sprintf("%.1f", x), "M"),
                     expand = expansion(mult = c(0.05, 0.05))) +
  scale_color_manual(values = c("United States of America" = tpa_colors[2],
                                "China" = tpa_colors[1])) +
  guides(color = "none") +
  labs_meta(meta, y = "Net flow (millions of students)") +
  theme_tpa()

save_chart(fig_042, "chapter4_fig_042")
```

### Chapter 4 - Figure 83 - Year of Last Enrollment Among Students Who Did Not Complete a Degree, by Cohort
```{r figure 083}
meta <- get_meta("fig_083")

year_levels <- c(
  "No degree, left without return year 1" = "Year 1",
  "No degree, left without return year 2" = "Year 2",
  "No degree, left without return year 3" = "Year 3",
  "No degree, left without return year 4" = "Year 4",
  "No degree, left without return year 5" = "Year 5",
  "No degree, left without return year 6" = "Year 6"
)

df_083 <- read_fig("fig_083") %>%
  rename(Cohort = 1) %>%
  select(Cohort, all_of(names(year_levels))) %>%   # timing only; still-enrolled/attained live in fig_084 + text
  mutate(across(-Cohort, as.numeric)) %>%
  pivot_longer(-Cohort, names_to = "ExitYear", values_to = "Share") %>%
  mutate(ExitYear = recode(ExitYear, !!!year_levels),
         Cohort   = factor(Cohort, levels = c("1996/2001", "2004/2009", "2012/2017")))

fig_083 <- ggplot(df_083, aes(ExitYear, Share, fill = Cohort)) +
  geom_col(width = 0.7, position = position_dodge(width = 0.78)) +
  geom_text(aes(label = fmt_pct_1(Share)),
            position = position_dodge(width = 0.78), vjust = -0.4,
            family = FONT_FAMILY, size = 2.8) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(0, 9), breaks = seq(0, 8, 2),
                     expand = EXPAND_BAR_TIGHT) +
  scale_fill_tpa() +
  labs_meta(meta, y = "Share of STEM entrants") +
  theme_tpa() +
  theme(axis.title.y = element_text(family = FONT_FAMILY, color = "grey25",
                                    size = AXIS_SIZE, margin = margin(r = 8)))

save_chart(fig_083, "chapter4_fig_083")
```

### Chapter 4 - Figure 84 - Bachelor's Degree Attainment and Enrollment Status for Students Who Entered STEM, by Cohort
```{r figure 084}
meta <- get_meta("fig_084")

df_084 <- read_fig("fig_084") %>%
  rename(Cohort = 1) %>%
  mutate(across(-Cohort, as.numeric)) %>%
  pivot_longer(-Cohort, names_to = "Outcome", values_to = "Share") %>%
  mutate(
    Outcome = recode(Outcome,
      "Left without return"                                    = "Left without a degree",
      "Degree Attained, Or Still Enrolled by Last Survey Year" = "Degree attained or still enrolled"),
    Outcome = factor(Outcome, levels = c("Degree attained or still enrolled",
                                         "Left without a degree")),
    Cohort  = factor(Cohort, levels = rev(c("1990/1994", "1996/2001",
                                            "2004/2009", "2012/2017"))))

fig_084 <- ggplot(df_084, aes(Share, Cohort, fill = Outcome)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = fmt_pct_1(Share)),
            position = position_stack(vjust = 0.5),
            family = FONT_FAMILY, size = 3, color = "white") +
  scale_x_continuous(labels = label_number(suffix = "%"),
                     limits = c(0, 100), breaks = seq(0, 100, 20),
                     expand = expansion(mult = c(0, 0))) +
  scale_fill_manual(values = c("Degree attained or still enrolled" = tpa_colors[1],
                               "Left without a degree"             = tpa_colors[2])) +
  labs_meta(meta, x = "Share of STEM entrants") +
  theme_tpa() +
  theme(axis.title.x = element_text(family = FONT_FAMILY, color = "grey25",
                                    size = AXIS_SIZE, margin = margin(t = 8)))

save_chart(fig_084, "chapter4_fig_084")
```

### Chapter 4 - Figure 118 - How Lopsided Is U.S. Education Spending?
```{r fig_118}
meta <- get_meta("fig_118")

num <- function(x) parse_number(as.character(x))

df_118 <- read_fig("fig_036") %>%
  mutate(
    K12   = num(`K-12 (in USD)`),
    Tert  = num(`Tertiary (in USD)`),
    Ratio = Tert / K12
  ) %>%
  filter(!Country %in% c("EU25 average", "G20 average", "China")) %>%
  mutate(Country = fct_reorder(Country, Ratio))

pal_b <- highlight_colors(
  levels(df_118$Country),
  highlighted = c("United States", "OECD average"),
  palette     = c(tpa_colors[1], tpa_colors[3]),
  default     = tpa_colors[2]
)

fig_118 <- ggplot(df_118, aes(Ratio, Country, fill = Country)) +
  geom_vline(xintercept = 1, color = "grey50", linewidth = LINE_WIDTH_REF) +
  geom_col(width = 0.72, alpha = ALPHA_COL) +
  geom_text(aes(label = sprintf("%.1f\u00d7", Ratio),
                hjust = if_else(Ratio >= 1, -0.2, 1.2)),
            family = FONT_FAMILY, fontface = "bold",
            size = (DATA_LABEL_SIZE - 2) / .pt, color = "grey25") +
  annotate("text", x = 1.02, y = 25, label = "spends equally on both levels",
           angle = 90, hjust = 0, vjust = 0, color = "white",
           size = (DATA_LABEL_SIZE - 3) / .pt, family = FONT_FAMILY) +
  scale_x_continuous(breaks = seq(0, 3, 0.5),
                     labels = function(x) paste0(x, "\u00d7"),
                     expand = expansion(mult = c(0, 0.08))) +
  scale_fill_manual(values = pal_b) +
  guides(fill = "none") +
  coord_cartesian(clip = "off") +
  labs_meta(meta, x = "Tertiary-to-K-12 spending ratio (per student)") +
  theme_horizontal_bar() +
  theme(
    axis.text.y = element_text(size = AXIS_SIZE - 2),
    aspect.ratio = 0.68,
    axis.title.x = element_text(family = FONT_FAMILY, color = "grey25",
                             size = AXIS_SIZE, margin = margin(t = 8))
  )

save_chart(fig_118, "chapter4_fig_118", size = "standard")
```

### Chapter 4 - Figure 119 - Does Spending More on Schools Buy Better Results?
```{r fig_119}
meta <- get_meta("fig_119")

num <- function(x) parse_number(as.character(x))

df_119 <- read_fig("fig_036") %>%
  mutate(
    K12  = num(`K-12 (in USD)`),
    PISA = as.numeric(`OECD Math Average`)
  ) %>%
  filter(!is.na(K12), !is.na(PISA),
         !Country %in% c("EU25 average", "G20 average", "China")) %>%
  mutate(pt_group = case_when(
    Group == "United States" ~ "United States",
    Group == "OECD average"  ~ "OECD average",
    TRUE                     ~ "Other"
  ))

oecd_x <- df_119$K12[df_119$Country == "OECD average"]
oecd_y <- df_119$PISA[df_119$Country == "OECD average"]

label_c <- df_119 %>%
  filter(Country %in% c("United States", "OECD average", "Korea", "Japan",
                        "Peru", "Luxembourg", "Türkiye")) %>%
  mutate(
    nx = case_when(Country == "United States" ~  1200,
                   Country == "OECD average"  ~ -1600,
                   TRUE                       ~     0),
    ny = case_when(Country == "United States" ~  -6,
                   Country == "OECD average"  ~   9,
                   TRUE                       ~   0)
  )

pal_c <- c("United States" = tpa_colors[1],
           "OECD average"  = tpa_colors[3],
           "Other"         = "grey80")

fig_119 <- ggplot(df_119, aes(K12, PISA)) +
  geom_hline(yintercept = oecd_y, color = "grey70",
             linetype = "dashed", linewidth = LINE_WIDTH_REF) +
  geom_vline(xintercept = oecd_x, color = "grey70",
             linetype = "dashed", linewidth = LINE_WIDTH_REF) +
  geom_point(aes(color = pt_group, size = pt_group), alpha = ALPHA_BUBBLE) +
  geom_label_repel_tpa(
    data         = label_c,
    mapping      = aes(label = Country, color = pt_group),
    nudge_x      = label_c$nx,
    nudge_y      = label_c$ny,
    with_segment = TRUE,
    direction    = "both"
  ) +
  annotate("text", x = max(df_119$K12) * 0.98, y = oecd_y + 3,
           label = "OECD average score", hjust = 1, vjust = 0,
           color = "grey55", size = (DATA_LABEL_SIZE - 3) / .pt,
           family = FONT_FAMILY) +
  annotate("text", x = oecd_x + 300, y = min(df_119$PISA) + 4,
           label = "OECD average spending", angle = 90,
           hjust = 0, vjust = 1, color = "grey55",
           size = (DATA_LABEL_SIZE - 3) / .pt, family = FONT_FAMILY) +
  scale_x_continuous(labels = label_dollar(scale = 1e-3, suffix = "K"),
                     expand = expansion(mult = c(0.04, 0.08))) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.08))) +
  scale_color_manual(values = pal_c) +
  scale_size_manual(values = c("United States" = 4, "OECD average" = 3.5, "Other" = 2.4)) +
  guides(color = "none", size = "none") +
  coord_cartesian(clip = "off") +
  labs_meta(meta,
            x = "K-12 spending per student (USD, PPP)",
            y = "PISA math score, 2022") +
  theme_tpa() +
  theme(
    axis.text.y = element_text(size = AXIS_SIZE - 2),
    aspect.ratio = 0.68,
    axis.title.x = element_text(family = FONT_FAMILY, color = "grey25",
                             size = AXIS_SIZE, margin = margin(t = 8))
  )

save_chart(fig_119, "chapter4_fig_119", size = "standard")
```

### Chapter 4 - Figure 120 - PISA Scores in Mathematics
```{r fig_120}
meta <- get_meta("fig_120")
num  <- function(x) parse_number(as.character(x))

OECD_MATH_MEAN <- 472  # OECD 2022 mathematics average

oecd_members <- c(
  "Australia","Austria","Belgium","Canada","Chile","Colombia","Costa Rica",
  "Czech Republic","Denmark","Estonia","Finland","France","Germany","Greece",
  "Hungary","Iceland","Ireland","Israel","Italy","Japan","Korea","Lat\via",
  "Lithuania","Mexico","Netherlands","New Zealand","Norway","Poland","Portugal",
  "Slovak Republic","Slovenia","Spain","Sweden","Switzerland","Türkiye",
  "United Kingdom","United States"
)

df_120 <- read_fig("fig_120") %>%
  filter(Country %in% oecd_members) %>%
  mutate(Gap = num(PISA_Math_Mean) - OECD_MATH_MEAN,
         Country = fct_reorder(Country, Gap))

palette_120 <- highlight_colors(
  all_values  = levels(df_120$Country),
  highlighted = "United States",
  palette     = tpa_colors[1],   # red
  default     = tpa_colors[2]    # navy
)

fig_120 <- ggplot(df_120, aes(Gap, Country, fill = Country)) +
  geom_vline(xintercept = 0, color = "grey50", linewidth = LINE_WIDTH_REF) +
  geom_col(width = 0.75, alpha = ALPHA_COL) +
  geom_text(aes(label = if_else(Gap == 0, "0", sprintf("%+.0f", Gap)),
                hjust = if_else(Gap >= 0, -0.2, 1.2)),
            family = FONT_FAMILY,
            fontface = "bold",
            size = (DATA_LABEL_SIZE - 2) / .pt,
            color = "grey25") +
  scale_x_continuous(limits = c(-100, 75),
                     breaks = seq(-100, 75, 25),
                     expand = EXPAND_NONE) +
  scale_fill_manual(values = palette_120) +
  guides(fill = "none") +
  labs_meta(meta, x = "Points above or below the OECD average") +
  theme_horizontal_bar() +
  theme(axis.text.y  = element_text(size = AXIS_SIZE - 2),
        axis.title.x = element_text(family = FONT_FAMILY,
                                    size = AXIS_SIZE - 2,
                                    color = "grey25",
                                    margin = margin(t = 6)))
save_chart(fig_120, "chapter4_fig_120", size = "long")
```

### Chapter 4 - Figure 121 - PISA Scores in Science
```{r fig_121}
meta <- get_meta("fig_121")
num  <- function(x) parse_number(as.character(x))

OECD_SCIENCE_MEAN <- 485  # OECD 2022 science average

oecd_members <- c(
  "Australia","Austria","Belgium","Canada","Chile","Colombia","Costa Rica",
  "Czech Republic","Denmark","Estonia","Finland","France","Germany","Greece",
  "Hungary","Iceland","Ireland","Israel","Italy","Japan","Korea","Latvia",
  "Lithuania","Mexico","Netherlands","New Zealand","Norway","Poland","Portugal",
  "Slovak Republic","Slovenia","Spain","Sweden","Switzerland","Türkiye",
  "United Kingdom","United States"
)

df_121 <- read_fig("fig_120") %>%
  filter(Country %in% oecd_members) %>%
  mutate(Gap = num(PISA_Science_Mean) - OECD_SCIENCE_MEAN,
         Country = fct_reorder(Country, Gap))

palette_121 <- highlight_colors(
  all_values  = levels(df_121$Country),
  highlighted = "United States",
  palette     = tpa_colors[1],   # red
  default     = tpa_colors[2]    # navy
)

fig_121 <- ggplot(df_121, aes(Gap, Country, fill = Country)) +
  geom_vline(xintercept = 0, color = "grey50", linewidth = LINE_WIDTH_REF) +
  geom_col(width = 0.75, alpha = ALPHA_COL) +
  geom_text(aes(label = if_else(Gap == 0, "0", sprintf("%+.0f", Gap)),
                hjust = if_else(Gap >= 0, -0.2, 1.2)),
            family = FONT_FAMILY,
            fontface = "bold",
            size = (DATA_LABEL_SIZE - 2) / .pt,
            color = "grey25") +
  scale_x_continuous(limits = c(-100, 75),
                     breaks = seq(-100, 75, 25),
                     expand = EXPAND_NONE) +
  scale_fill_manual(values = palette_121) +
  guides(fill = "none") +
  labs_meta(meta, x = "Points above or below the OECD average") +
  theme_horizontal_bar() +
  theme(axis.text.y  = element_text(size = AXIS_SIZE - 2),
        axis.title.x = element_text(family = FONT_FAMILY,
                                    size = AXIS_SIZE - 2,
                                    color = "grey25",
                                    margin = margin(t = 6)))
save_chart(fig_121, "chapter4_fig_121", size = "long")
```

### Chapter 4 - Figure 122 - How Do OECD Countries Rank in Math Proficiency?
```{r}
meta <- get_meta("fig_122")
num  <- function(x) parse_number(as.character(x))

keep_rows <- c(
  "Australia","Austria","Belgium","Canada","Chile","Colombia","Costa Rica",
  "Czech Republic","Denmark","Estonia","Finland","France","Germany","Greece",
  "Hungary","Iceland","Ireland","Israel","Italy","Japan","Korea","Latvia",
  "Lithuania","Mexico","Netherlands","New Zealand","Norway","Poland","Portugal",
  "Slovak Republic","Slovenia","Spain","Sweden","Switzerland","Türkiye",
  "United Kingdom","United States","OECD average"
)

df_122 <- read_fig("fig_122") %>%
  rename(Country = 1) %>%
  mutate(Country = str_remove(Country, "\\*+$"),      # drop sampling asterisks
         `Level 5` = num(`Level 5`),
         `Level 6` = num(`Level 6`),
         Top = `Level 5` + `Level 6`) %>%
  filter(Country %in% keep_rows) %>%
  mutate(Country = fct_reorder(Country, Top))

palette_122 <- highlight_colors(
  all_values  = levels(df_122$Country),
  highlighted = c("United States", "OECD average"),
  palette     = c(tpa_colors[1], tpa_colors[3]),   # red, gold
  default     = tpa_colors[2]                       # navy
)

fig_122 <- ggplot(df_122, aes(Top, Country, fill = Country)) +
  geom_col(width = 0.75, alpha = ALPHA_COL) +
  geom_text(aes(label = sprintf("%.1f", Top), hjust = -0.2),
            family = FONT_FAMILY,
            fontface = "bold",
            size = (DATA_LABEL_SIZE - 2) / .pt,
            color = "grey25") +
  scale_x_continuous(limits = c(0, 25),
                     breaks = seq(0, 25, 5),
                     expand = EXPAND_NONE) +
  scale_fill_manual(values = palette_122) +
  guides(fill = "none") +
  labs_meta(meta, x = "Share of students scoring at Level 5 or 6 (%)") +
  theme_horizontal_bar() +
  theme(axis.text.y  = element_text(size = AXIS_SIZE - 2),
        axis.title.x = element_text(family = FONT_FAMILY,
                                    size = AXIS_SIZE - 2,
                                    color = "grey25",
                                    margin = margin(t = 6)))
save_chart(fig_122, "chapter4_fig_122", size = "long")
```

### Chapter 4 - Figure 123 - How Do OECD Countries Rank in Science Proficiency?
```{r}
meta <- get_meta("fig_123")
num  <- function(x) parse_number(as.character(x))

keep_rows <- c(
  "Australia","Austria","Belgium","Canada","Chile","Colombia","Costa Rica",
  "Czech Republic","Denmark","Estonia","Finland","France","Germany","Greece",
  "Hungary","Iceland","Ireland","Israel","Italy","Japan","Korea","Latvia",
  "Lithuania","Mexico","Netherlands","New Zealand","Norway","Poland","Portugal",
  "Slovak Republic","Slovenia","Spain","Sweden","Switzerland","Türkiye",
  "United Kingdom","United States","OECD average"
)

df_123 <- read_fig("fig_123") %>%
  rename(Country = 1) %>%
  mutate(Country = str_remove(Country, "\\*+$"),
         `Level 5` = num(`Level 5`),
         `Level 6` = num(`Level 6`),
         Top = `Level 5` + `Level 6`) %>%
  filter(Country %in% keep_rows) %>%
  mutate(Country = fct_reorder(Country, Top))

palette_123 <- highlight_colors(
  all_values  = levels(df_123$Country),
  highlighted = c("United States", "OECD average"),
  palette     = c(tpa_colors[1], tpa_colors[3]),   # red, gold
  default     = tpa_colors[2]                       # navy
)

fig_123 <- ggplot(df_123, aes(Top, Country, fill = Country)) +
  geom_col(width = 0.75, alpha = ALPHA_COL) +
  geom_text(aes(label = sprintf("%.1f", Top), hjust = -0.2),
            family = FONT_FAMILY,
            fontface = "bold",
            size = (DATA_LABEL_SIZE - 2) / .pt,
            color = "grey25") +
  scale_x_continuous(limits = c(0, 25),
                     breaks = seq(0, 25, 5),
                     expand = EXPAND_NONE) +
  scale_fill_manual(values = palette_123) +
  guides(fill = "none") +
  labs_meta(meta, x = "Share of students scoring at Level 5 or 6 (%)") +
  theme_horizontal_bar() +
  theme(axis.text.y  = element_text(size = AXIS_SIZE - 2),
        axis.title.x = element_text(family = FONT_FAMILY,
                                    size = AXIS_SIZE - 2,
                                    color = "grey25",
                                    margin = margin(t = 6)))
save_chart(fig_123, "chapter4_fig_123", size = "long")
```
### Chapter 4 - Figure 124 - Share of Countries' Students Who Study Abroad
```{r fig_124}
meta <- get_meta("fig_124")
num  <- function(x) parse_number(as.character(x))   # parse_number chokes on numeric cols

# Wide CSV: one column per year (1997–2023); first column is labelled
# "Year" but holds country names. Pivot long, coerce, keep latest year.
mobility <- read_fig("fig_124") %>%
  rename(country = 1) %>%
  select(-any_of("Grand Total")) %>%
  pivot_longer(-country, names_to = "year", values_to = "ratio") %>%
  mutate(year = as.integer(year), ratio = num(ratio)) %>%
  filter(!is.na(ratio))

# OECD members present in the data (Korea = "South Korea") plus the two
# rivals. Luxembourg excluded: ratio > 100% (no full domestic system).
oecd_plus <- c(
  "Australia", "Austria", "Belgium", "Canada", "Chile", "Colombia",
  "Costa Rica", "Czechia", "Denmark", "Estonia", "Finland", "France",
  "Germany", "Greece", "Hungary", "Iceland", "Ireland", "Israel", "Italy",
  "Japan", "South Korea", "Latvia", "Lithuania", "Mexico", "Netherlands",
  "New Zealand", "Norway", "Poland", "Portugal", "Slovakia", "Slovenia",
  "Spain", "Sweden", "Switzerland", "Turkey", "United Kingdom",
  "United States", "China", "India"
)

df_124 <- mobility %>%
  filter(country %in% oecd_plus) %>%
  group_by(country) %>%
  slice_max(year, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(
    Country = recode(country, "South Korea" = "Korea"),
    grp = case_when(
      country == "United States"      ~ "United States",
      country %in% c("China", "India") ~ "Rival",
      TRUE                             ~ "OECD"
    )
  ) %>%
  arrange(ratio) %>%
  mutate(Country = fct_inorder(Country))   # US (lowest) at the floor

# House highlight pattern, mirroring fig_120/122: US in red, the rivals
# picked out in gold, all other OECD in navy.
palette_124 <- highlight_colors(
  all_values  = df_124$grp,
  highlighted = c("United States", "Rival"),
  palette     = c(tpa_colors[1], tpa_colors[3]),
  default     = tpa_colors[2]
)

fig_124 <- ggplot(df_124, aes(ratio, Country, color = grp)) +
  geom_segment(aes(x = 0, xend = ratio, yend = Country),
               linewidth = LINE_WIDTH_REF) +
  geom_point(size = POINT_SIZE) +
  geom_text(aes(label = sprintf("%.1f", ratio)), hjust = -0.35,
            family = FONT_FAMILY, fontface = "bold",
            size = (DATA_LABEL_SIZE - 2) / .pt, color = "grey25") +
  scale_x_continuous(labels = label_number(suffix = "%"),
                     expand = expansion(mult = c(0, 0.08))) +
  scale_color_manual(values = palette_124) +
  guides(color = "none") +
  coord_cartesian(clip = "off") +
  labs_meta(meta, x = "Share of tertiary students studying abroad") +
  theme_dotplot() +
  theme(axis.text.y  = element_text(size = AXIS_SIZE - 2),
        axis.title.x = element_text(family = FONT_FAMILY, color = "grey25",
                                    size = AXIS_SIZE - 2, margin = margin(t = 6)))

save_chart(fig_124, "chapter4_fig_124")
```

# Chapter 5
### Chapter 5 - Figure 44 - Fractional Shares of Top-Cited Emerging-Technology Publications: China–US Trajectories, 2015–2024
```{r figure 044}
meta <- get_meta("fig_044")

df_044 <- read_fig("fig_044") %>%
  filter(!is.na(Year)) %>%
  mutate(across(everything(), as.numeric)) %>%
  pivot_longer(-Year, names_to = "Country", values_to = "Share") %>%
  mutate(Share   = Share * 100,
         Country = case_when(Country == "china" ~ "China",
                             Country == "us"    ~ "United States")) %>%
  group_by(Year) %>%
  mutate(label_y = Share + ifelse(Share == max(Share), 1.3, -1.3)) %>%
  ungroup()

fig_044 <- ggplot(df_044, aes(Year, Share, color = Country, group = Country)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_point(size = 2) +
  geom_text(aes(y = label_y, label = fmt_pct_0(Share)),
            size = DATA_LABEL_SIZE / .pt, fontface = "bold",
            family = FONT_FAMILY, show.legend = FALSE) +
  scale_x_years(df_044$Year, by = 1) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(0, 25), breaks = seq(0, 25, 5),
                     expand = EXPAND_LINE) +
  scale_color_china_us() +
  labs_meta(meta, y = "Fractional share of top-cited publications") +
  coord_cartesian(clip = "off") +
  theme_tpa() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
        plot.margin  = margin(t = 10, r = 24, b = 10, l = 24))

save_chart(fig_044, "chapter5_fig_044")
```

### Chapter 5 - Figure 45 - U.S. and Chinese authorship shares in top emerging technology publications, 2015–2025
```{r figure 045}
meta <- get_meta("fig_045")

df_045 <- read_fig("fig_045") %>%
  filter(!is.na(Year)) %>%
  mutate(across(everything(), as.numeric)) %>%
  pivot_longer(-Year, names_to = "Country", values_to = "Share") %>%
  mutate(Share   = Share * 100,
         Country = case_when(Country == "pct_papers_with_us" ~ "United States",
                             Country == "pct_papers_with_cn" ~ "China")) %>%
  group_by(Year) %>%
  mutate(label_y = Share + ifelse(Share == max(Share), 0.9, -0.9)) %>%
  ungroup()

fig_045 <- ggplot(df_045, aes(Year, Share, color = Country, group = Country)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_point(size = 2) +
  geom_text(aes(y = label_y, label = fmt_pct_1(Share)),
            size = DATA_LABEL_SIZE / .pt, fontface = "bold",
            family = FONT_FAMILY, show.legend = FALSE) +
  scale_x_years(df_045$Year, by = 1) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(15, 35), breaks = seq(15, 35, 2),
                     expand = EXPAND_LINE) +
  scale_color_china_us() +
  labs_meta(meta, y = "Share of top papers with at least one author") +
  coord_cartesian(clip = "off") +
  theme_tpa() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
        plot.margin  = margin(t = 10, r = 30, b = 10, l = 30))

save_chart(fig_045, "chapter5_fig_045")
```

### Chapter 5 - Figure 46 - Share of the World's Most-Cited Researchers, US vs. China
```{r figure 046}
meta <- get_meta("fig_046")

df_046 <- read_fig("fig_046") %>%
  filter(!is.na(Year)) %>%
  mutate(across(everything(), as.numeric)) %>%
  transmute(Year,
            China         = CN_Share * 100,
            `United States` = US_Share * 100) %>%
  pivot_longer(-Year, names_to = "Country", values_to = "Share")

labels_046 <- last_by_group(df_046, Year, Country)

fig_046 <- ggplot(df_046, aes(Year, Share, color = Country, group = Country)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_point(size = POINT_SIZE) +
  geom_label_repel_tpa(
    data         = labels_046,
    mapping      = aes(label = paste0(Country, ": ", fmt_pct_0(Share))),
    nudge_x      = 0.5,
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_046$Year, by = 1, pad_right = PAD_RIGHT_WIDE) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(0, 60),
                     breaks = seq(0, 60, 15),
                     expand = EXPAND_LINE) +
  scale_color_china_us() +
  guides(color = "none") +
  labs_meta(meta, y = "Share of global top 1% most-cited researchers") +
  theme_tpa() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5))

save_chart(fig_046, "chapter5_fig_046")
```



### Chapter 5 - Figure 49 - Rolling Average of US Recipients of Top Awards in STEM Fields (2000-2024)
```{r figure 049}
meta <- get_meta("fig_049")

short_names <- c(
  "10-Year Rolling Average of US Affiliated Scientists Who Won A Nobel Prize in STEM" = "Nobel Prize - US Affiliated (10yr)",
  "10-Year Rolling Average of US Born Scientists Who Won A Nobel Prize in STEM"       = "Nobel Prize - US Born (10yr)",
  "10-Year Rolling Average of US Born Scientists Who Won A Turing Award"              = "Turing Award - US Born (10yr)",
  "10-Year Rolling Average of US Affiliated Scientists Who Won A Turing Award"        = "Turing Award - US Affiliated (10yr)"
)

df_049 <- read_fig("fig_049") %>%
  filter(!is.na(Year)) %>%
  mutate(across(everything(), as.numeric)) %>%
  pivot_longer(-Year, names_to = "Award", values_to = "Count") %>%
  filter(!is.na(Count), !str_detect(Award, "Fields Medal")) %>%
  mutate(Award = recode(Award, !!!short_names)) %>%
  order_by_latest(Award, Count)

labels_049 <- last_by_group(df_049, Year, Award)

fig_049 <- ggplot(df_049, aes(Year, Count, color = Award, group = Award)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_label_repel_tpa(
    data         = labels_049,
    mapping      = aes(label = paste0(Award, ": ", Count)),
    nudge_x      = 0.5,
    hjust        = 0,
    direction    = "y",
    with_segment = TRUE
  ) +
  scale_x_years(df_049$Year, by = 2, pad_right = .6) +
  scale_y_continuous(limits = c(0, 60),
                     breaks = seq(0, 60, 10),
                     expand = EXPAND_LINE) +
  scale_color_tpa() +
  guides(color = "none") +
  labs_meta(meta, y = "Rolling average of recipients") +
  theme_tpa()

save_chart(fig_049, "chapter5_fig_049")
```

### Chapter 5 - Figure 50 - Distribution of Citizenship Status of U.S. Affiliated Nobel Laureates (1900-2025)
```{r figure 050}
meta <- get_meta("fig_050")

df_050 <- read_fig("fig_050") %>%
  rename(Status = 1) %>%
  pivot_longer(c(chemistry, medicine, physics),
               names_to = "Prize", values_to = "Count") %>%
  group_by(Status) %>%
  summarise(Count = sum(Count, na.rm = TRUE), .groups = "drop") %>%
  mutate(Share = Count / sum(Count) * 100)

palette_050 <- c("Born Citizen" = tpa_colors[2], "Immigrant" = tpa_colors[1])

fig_050 <- plot_composition_pie(
  df_050,
  category_col   = "Status",
  value_col      = "Share",
  palette        = palette_050,
  category_order = c("Born Citizen", "Immigrant"),
  label_style    = "full",     # "Born Citizen: 66.8%"
  label_fmt      = fmt_pct_1,
  legend         = FALSE
) +
  labs_meta(meta)

save_chart(fig_050, "chapter5_fig_050")
```

### Chapter 5 - Figure 51 - Distribution of Country of Origin of U.S. Affiliated Immigrant Nobel Laureates (1900-2025)
```{r figure 051}
meta <- get_meta("fig_051")

df_051 <- read_fig("fig_051") %>%
  rename(Country = 1, Count = 2) %>%
  mutate(Count   = as.numeric(Count),
         Country = case_when(Country == "the Netherlands" ~ "Netherlands",
                             Country == "South Korea"     ~ "Republic of Korea",
                             TRUE                         ~ Country)) %>%
  group_by(Country) %>%
  summarise(Count = sum(Count, na.rm = TRUE), .groups = "drop")

world_051 <- join_country_data(df_051)

fig_051 <- ggplot(world_051) +
  geom_sf(aes(fill = Count), color = "white", linewidth = 0.2) +
  # on-map counts removed: clustered European centroids overlapped illegibly.
  # Magnitude now reads off the color legend.
  scale_fill_gradient(low = TPA_RED_LIGHT, high = tpa_colors[1], na.value = "grey92",
                      trans = "sqrt", breaks = c(1, 3, 6, 10, 15),
                      labels = label_comma(), name = "Immigrant laureates") +
  coord_sf(crs = "+proj=robin", expand = FALSE) +
  labs_meta(meta) +
  theme_map()

save_chart(fig_051, "chapter5_fig_051")
```

### Chapter 5 - Figure 54 - Cited By Count
```{r fig_054}
meta <- get_meta("fig_054")
dat  <- read_bibliometric("fig_054")
ord  <- dat %>% arrange(median_50th) %>% pull(Company) %>% as.character()

fig_054 <- plot_bibliometric_dotrange(
  dat,
  x_label       = "Citations",
  log_scale     = FALSE,
  company_order = ord
) +
  labs_meta(meta)

save_chart(fig_054, "chapter5_fig_054")
```

### Chapter 5 - Figure 55 - H Index
```{r fig_055}
meta <- get_meta("fig_055")
dat  <- read_bibliometric("fig_055")
# same order as fig_054 so the two panels line up company-by-company
ord  <- read_bibliometric("fig_054") %>% arrange(median_50th) %>% pull(Company) %>% as.character()

fig_055 <- plot_bibliometric_dotrange(
  dat,
  x_label       = "h-index",
  log_scale     = FALSE,
  company_order = ord
) +
  labs_meta(meta)

save_chart(fig_055, "chapter5_fig_055")
```

### Chapter 5 - Figure 63 - Gross domestic expenditures on R&D, by selected region, country, or economy:
```{r figure 063}
meta <- get_meta("fig_063")
df   <- read_country_lines("fig_063", group_col = "Region", value_col = "Value") %>%
  mutate(Value = Value / 1000)

fig <- plot_country_lines(df, meta, "R&D expenditure (USD billions, PPP)",
                          group_col   = "Region",
                          value_col   = "Value",
                          label_fmt   = function(x) paste0("$", sprintf("%.0f", x), "B"),
                          pad_right   = PAD_RIGHT_WIDE,
                          highlight   = c("United States", "China"),
                          color_scale = scale_color_china_us(),
                          y_args      = list(
                            labels = function(x) paste0("$", x, "B"),
                            limits = c(0, 1000),
                            breaks = seq(0, 1000, 100),
                            expand = EXPAND_BAR_TIGHT
                          ))

save_chart(fig, "chapter5_fig_063")
```

### Chapter 5 - Figure 64 - R&D Intensity in Critical Technology Sectors by Country
```{r figure 064}
meta <- get_meta("fig_064")
df   <- read_country_lines("fig_064", group_col = "Country", value_col = "Share") %>%
  filter(Country != "Russia") %>%
  mutate(Country = if_else(Country == "US", "United States", Country),
         Share   = Share * 100)

fig <- plot_country_lines(df, meta, "R&D intensity (% of GDP)",
                          value_col   = "Share",
                          label_fmt   = fmt_pct_1,
                          pad_right   = PAD_RIGHT_WIDE,
                          highlight   = c("United States", "China"),
                          color_scale = scale_color_china_us(),
                          y_args      = list(
                            labels = label_number(suffix = "%"),
                            limits = c(0, 12),
                            breaks = seq(0, 12, 2),
                            expand = EXPAND_BAR_TIGHT
                          ))

save_chart(fig, "chapter5_fig_064")
```


### Chapter 5 - Figure 125 - R&D Intensity by Type
```{r figure 125}
meta <- get_meta("fig_125")

rd_type_colors <- c(
  "Basic research"           = tpa_colors[1],  
  "Applied research"         = tpa_colors[2],  
  "Experimental development" = tpa_colors[3]  
)

df <- read_fig("fig_125") %>%
  rename(Year = 1, Country = 2, Type = 3, value = 4) %>%
  mutate(
    Year    = as.integer(round(as.numeric(as.character(Year)))),
    value   = as.numeric(value),
    Country = factor(Country, levels = c("United States", "China")),
    Type    = recode(as.character(Type),
                     "Basic Research"           = "Basic research",
                     "Applied Research"          = "Applied research",
                     "Experimental Development"  = "Experimental development"),
    Type    = factor(Type, levels = c("Basic research", "Applied research",
                                      "Experimental development"))
  )

fig <- ggplot(df, aes(Year, value, fill = Type)) +
  geom_area(position = position_stack(reverse = TRUE), alpha = ALPHA_AREA) +
  facet_wrap(~ Country) +
  scale_fill_manual(values = rd_type_colors, breaks = names(rd_type_colors)) +
  scale_x_continuous(breaks = c(1991, 2000, 2010, 2024), expand = EXPAND_LINE) +
  scale_y_continuous(labels = scales::label_dollar(suffix = "B"), expand = EXPAND_BAR) +
  labs_meta(meta) +
  theme_facet()

save_chart(fig, "chapter5_fig_125")
```

### Chapter 5 - Figure 126 - R&D Spending by Sector, U.S. and China
```{r fig_126}
meta <- get_meta("fig_126")

sector_colors <- c(
  "Business enterprise" = tpa_colors[4],
  "Government"          = tpa_colors[5],
  "Higher education"    = tpa_colors[6],
  "Private non-profit"  = tpa_colors[7]
)

df <- read_fig("tab_022") %>%
  rename(Year = 1, Country = 2, Sector = 3, value = 4) %>%
  mutate(
    Year   = as.integer(round(as.numeric(as.character(Year)))),
    value  = as.numeric(value),
    Sector = recode(as.character(Sector),
                    "Business enterprise sector" = "Business enterprise",
                    "Government sector"          = "Government",
                    "Higher education sector"    = "Higher education",
                    "Private non-profit sector"  = "Private non-profit")
  ) %>%
  filter(Sector %in% names(sector_colors)) %>%
  complete(Country, Sector, Year, fill = list(value = 0)) %>%   # China private non-profit -> $0 all years
  mutate(
    Country = factor(Country, levels = c("United States", "China")),
    Sector  = factor(Sector, levels = names(sector_colors))
  )

fig_126 <- ggplot(df, aes(Year, value, fill = Sector)) +
  geom_area(position = position_stack(reverse = TRUE), alpha = ALPHA_AREA) +
  facet_wrap(~ Country) +
  scale_fill_manual(values = sector_colors, breaks = names(sector_colors)) +
  scale_x_continuous(breaks = c(1991, 2000, 2010, 2024), expand = EXPAND_LINE) +
  scale_y_continuous(labels = scales::label_dollar(suffix = "B"), expand = EXPAND_BAR) +
  labs_meta(meta) +
  theme_facet()
save_chart(fig_126, "chapter5_fig_126")
```


# Chapter 6

### Chapter 6 - Figure 65 - International PhD Recipients – Intention to Stay
```{r figure 065}
meta <- get_meta("fig_065")

df_065 <- read_fig("fig_065") %>%
  rename(Year = 1, Share = 2) %>%
  filter(!is.na(Year)) %>%
  mutate(across(everything(), as.numeric),
         Share = Share * 100)

fig_065 <- ggplot(df_065, aes(Year, Share)) +
  geom_col(fill = tpa_colors[2], width = 0.7, alpha = ALPHA_COL) +
  geom_text(aes(label = fmt_pct_0(Share)),
            vjust = -0.5,
            color = tpa_colors[2],
            size = DATA_LABEL_SIZE / .pt,
            fontface = "bold",
            family = FONT_FAMILY) +
  scale_x_years(df_065$Year, by = 1) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(60, 80),
                     breaks = seq(60, 80, 2),
                     expand = EXPAND_TIGHT_TOP,
                     oob = scales::oob_keep) +
  coord_cartesian(ylim = c(60, 80)) +
  labs_meta(meta, y = "Intention to stay in the United States") +
  theme_tpa()

save_chart(fig_065, "chapter6_fig_065")
```

### Chapter 6 - Figure 67 - International STEM PhD Recipients – Stay Rates
### Chapter 6 - Figure 67 - International STEM PhD Recipients – Stay Rates
```{r figure 067}
meta <- get_meta("fig_067")
df_067 <- read_fig("fig_067") %>%
  filter(!is.na(Year)) %>%
  mutate(across(everything(), as.numeric)) %>%
  pivot_longer(-Year, names_to = "Rate", values_to = "Share") %>%
  filter(!is.na(Share)) %>%
  mutate(
    Share = Share * 100,
    Rate  = factor(Rate, levels = c("5-year stay rate", "10-year stay rate"))
  )

# first point, last point, and peak of each line; first-point labels
# left-align so they extend right off the marker instead of over the axis
labels_067 <- flp_by_group(df_067, Year, Share, Rate) %>%
  mutate(hjust = if_else(Year == min(Year), 1, 0.5))

palette_067 <- c(
  "5-year stay rate"  = tpa_colors[2],   # navy
  "10-year stay rate" = tpa_colors[1]    # red
)

fig_067 <- ggplot(df_067, aes(Year, Share, color = Rate, group = Rate)) +
  geom_line(linewidth = LINE_WIDTH) +
  geom_point(size = 2) +
  geom_text_tpa(
    data        = labels_067,
    mapping     = aes(label = fmt_pct_1(Share), hjust = hjust),
    show.legend = FALSE
  ) +
  scale_x_years(df_067$Year, by = 2, pad_right = .09, pad_left = 0.09) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(50, 75),
                     breaks = seq(50, 75, 5),
                     expand = expansion(mult = c(0, 0.05))) +
  scale_color_manual(values = palette_067) +
  coord_cartesian(clip = "off") +
  labs_meta(meta, y = "Stay rate") +
  theme_tpa() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5))

save_chart(fig_067, "chapter6_fig_067")
```

### Chapter 6 - Figure 69 - Trends in Distribution of Employment Sector of Doctorate Recipients with Definite Postgraduation Commitments for Non-postdoc Employment in the United States (1993-2023) - TVH Recipients
```{r figure 069}
meta <- get_meta("fig_069")
df   <- read_sector_employment("fig_069")
fig  <- plot_sector_employment(df, meta, "Share of TVH doctorate recipients",
                               y_limits = c(0, 80),
                               y_breaks = seq(0, 80, 10))
save_chart(fig, "chapter6_fig_069")
```

### Chapter 6 - Figure 70 - Trends in Distribution of Employment Sector of Doctorate Recipients with Definite Postgraduation Commitments for Non-postdoc Employment in the United States (1993-2023) - US Recipients
```{r figure 070}
meta <- get_meta("fig_070")
df   <- read_sector_employment("fig_070")
fig  <- plot_sector_employment(df, meta, "Share of US doctorate recipients",
                               y_limits = c(0, 60),
                               y_breaks = seq(0, 60, 10))
save_chart(fig, "chapter6_fig_070")
```

### Chapter 6 - Figure 71 - STEM OPT Approvals, Applications Received, Approval Rates
```{r figure 071}
meta <- get_meta("fig_071")

opt_years <- c(2008, 2013, 2018, 2023, 2025)
# opt_years <- c(2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024, 2025)


df_071 <- read_fig("fig_071") %>%
  filter(!is.na(Year)) %>%
  mutate(across(everything(), as.numeric)) %>%
  filter(Year %in% opt_years) %>%
  select(Year, Received, Approved) %>%
  pivot_longer(-Year, names_to = "Status", values_to = "Count") %>%
  mutate(Status = factor(Status, levels = c("Received", "Approved")))

fig_071 <- ggplot(df_071, aes(factor(Year), Count, fill = Status)) +
  geom_col(position = position_dodge(width = 0.8),
           width = 0.7, alpha = ALPHA_COL) +
  geom_text(aes(label = label_comma()(Count)),
            position = position_dodge(width = 0.8),
            vjust = -0.5,
            size = (DATA_LABEL_SIZE - 2) / .pt,
            fontface = "bold",
            family = FONT_FAMILY) +
  scale_y_continuous(labels = label_comma(),
                     expand = EXPAND_TIGHT_TOP) +
  scale_fill_manual(name = NULL,
                    values = c("Received" = tpa_colors[2],
                               "Approved" = tpa_colors[1])) +
  labs_meta(meta, y = "STEM OPT applications") +
  coord_cartesian(clip = "off") +
  theme_tpa()

save_chart(fig_071, "chapter6_fig_071")
```

### Chapter 6 - Figure 76 - Distribution of PERM Applications by Status (2008-2024)
```{r figure 076}
meta <- get_meta("fig_076")

df_076 <- read_fig("fig_076") %>%
  filter(!is.na(Year)) %>%
  select(Year,
         `Used`    = Certified,
         `Expired` = `Certified-expired`,
         `Denied`  = Denied) %>%
  mutate(across(everything(), as.numeric)) %>%
  pivot_longer(-Year, names_to = "Status", values_to = "Count") %>%
  filter(!is.na(Count)) %>%
  mutate(Status = factor(Status, levels = c("Denied", "Expired", "Used")))

palette_076 <- c(
  "Used"    = tpa_colors[3],   # gold — certifications actually used
  "Expired" = tpa_colors[1],   # red  — certified but lapsed unused
  "Denied"  = tpa_colors[6]
)

fig_076 <- ggplot(df_076, aes(Year, Count, fill = Status)) +
  geom_area(alpha = ALPHA_AREA) +
  scale_x_years(df_076$Year, by = 2) +
  scale_y_continuous(labels = label_comma(), expand = EXPAND_BAR) +
  scale_fill_manual(values = palette_076,
                    breaks = c("Used", "Expired", "Denied")) +
  labs_meta(meta, y = "PERM certifications") +
  theme_tpa()

save_chart(fig_076, "chapter6_fig_076")
```


### Chapter 6 - Figure 78 - Trends in J-1 Visa Participation by Country (Top 10 overall from the last decade)
```{r figure 078}
meta <- get_meta("fig_078")

df_078 <- read_fig("fig_078") %>%
  filter(!is.na(Year)) %>%
  mutate(across(everything(), as.numeric)) %>%
  pivot_longer(-Year, names_to = "Country", values_to = "Share") %>%
  filter(!is.na(Share)) %>%
  mutate(Share    = Share * 100,
         is_china = Country == "China")

china_078 <- filter(df_078, is_china)
other_078 <- filter(df_078, !is_china)
china_last <- china_078 %>% filter(Year == max(Year))

fig_078 <- ggplot() +
  geom_line(data = other_078,
            aes(Year, Share, group = Country),
            color = "grey75", linewidth = LINE_WIDTH) +
  geom_line(data = china_078,
            aes(Year, Share),
            color = tpa_colors[1], linewidth = LINE_WIDTH_DENSE) +
  geom_point(data = china_last,
             aes(Year, Share), color = tpa_colors[1], size = 2.5) +
  geom_label_repel_tpa(
    data    = china_last,
    mapping = aes(Year, Share, label = paste0("China: ", fmt_pct_1(Share))),
    nudge_x = 0.4, hjust = 0, direction = "y", with_segment = TRUE
  ) +
  annotate("text", x = 2015, y = 11.6, label = "China", hjust = 0,
           color = tpa_colors[1], fontface = "bold",
           family = FONT_FAMILY, size = DATA_LABEL_SIZE / .pt) +
  annotate("text", x = 2020.2, y = 7.0, label = "Other top-five countries",
           hjust = 0, color = "grey55",
           family = FONT_FAMILY, size = (DATA_LABEL_SIZE - 2) / .pt) +
  scale_x_years(df_078$Year, by = 1, pad_right = PAD_RIGHT_STANDARD) +
  scale_y_continuous(labels = label_number(suffix = "%"),
                     limits = c(0, 12), breaks = seq(0, 12, 2),
                     expand = EXPAND_BAR_TIGHT) +
  labs_meta(meta, y = "Share of J-1 participants") +
  theme_tpa()

save_chart(fig_078, "chapter6_fig_078")
```

### Chapter 6 - Figure 79 - Immigrant vs Native-Born Workforce Participation Rate Gap, by Country (2024)
```{r figure 079}
meta <- get_meta("fig_079")

df_079 <- read_fig("fig_079") %>%
  rename(Gap = `Gap with the native-born`) %>%
  mutate(Gap = as.numeric(Gap),
         Country = fct_reorder(Country, Gap))

palette_079 <- highlight_colors(
  all_values  = levels(df_079$Country),
  highlighted = c("United States", "OECD average"),
  palette     = c(tpa_colors[1], tpa_colors[3]),   # red, gold
  default     = tpa_colors[2]                       # navy
)

fig_079 <- ggplot(df_079, aes(Gap, Country, fill = Country)) +
  geom_vline(xintercept = 0, color = "grey50", linewidth = LINE_WIDTH_REF) +
  geom_col(width = 0.75, alpha = ALPHA_COL) +
  geom_text(aes(label = sprintf("%.1f", Gap),
                hjust = if_else(Gap >= 0, -0.2, 1.2)),
            family = FONT_FAMILY,
            fontface = "bold",
            size = (DATA_LABEL_SIZE - 2) / .pt,
            color = "grey25") +
  scale_x_continuous(limits = c(-15, 20),
                     breaks = seq(-15, 20, 5),
                     expand = EXPAND_NONE) +
  scale_fill_manual(values = palette_079) +
  guides(fill = "none") +
  labs_meta(meta, x = "Percentage point gap with the native-born") +
  theme_horizontal_bar() +
  theme(axis.text.y = element_text(size = AXIS_SIZE - 2))

save_chart(fig_079, "chapter6_fig_079", size = "long")
```
### Chapter 6 - Figure 87 - Trends in H-1B Denials, by Application Type (2009-2025)
```{r figure 087}
meta <- get_meta("fig_087")

df_087 <- read_hb_denials("fig_087") %>% filter(Type == "New Denials") %>%
  select(Year, All = Count) %>%
  left_join(
    read_hb_denials("fig_088") %>% filter(Type == "New Denials") %>%
      select(Year, Tech = Count),
    by = "Year"
  )

fig_087 <- ggplot(df_087, aes(Year)) +
  geom_area(aes(y = All),  fill = tpa_colors[2], alpha = ALPHA_AREA) +
  geom_area(aes(y = Tech), fill = tpa_colors[1], alpha = ALPHA_AREA) +
  geom_line(aes(y = All),  color = tpa_colors[2], linewidth = LINE_WIDTH) +
  geom_line(aes(y = Tech), color = tpa_colors[1], linewidth = LINE_WIDTH) +
  annotate("text", x = 2017, y = 32000, label = "All industries",
           color = tpa_colors[2], fontface = "bold", hjust = 0.5,
           family = FONT_FAMILY, size = DATA_LABEL_SIZE / .pt) +
  annotate("text", x = 2018.75, y = 14000, label = "Tech occupations\n(NAICS 54)",
           color = "white", fontface = "bold", hjust = 0.5, lineheight = 0.9,
           family = FONT_FAMILY, size = (DATA_LABEL_SIZE - 1) / .pt) +
  scale_x_years(df_087$Year, by = 2) +
  scale_y_continuous(labels = label_comma(), expand = EXPAND_BAR) +
  labs_meta(meta, y = "New H-1B denials") +
  theme_tpa()

save_chart(fig_087, "chapter6_fig_087")
```
### Chapter 6 - Figure 72 - Trends in H-1B Approvals, by Application Type (2009-2025)
```{r figure 087}
meta <- get_meta("fig_072")

total_approvals <- function(name) {
  read_fig(name) %>%
    filter(!is.na(Year)) %>%
    select(Year, `New Approved`, `Renewal Approved`) %>%
    mutate(across(everything(), as.numeric)) %>%
    transmute(Year, Total = `New Approved` + `Renewal Approved`)
}

df_072 <- total_approvals("fig_072") %>% rename(All = Total) %>%
  left_join(total_approvals("fig_074") %>% rename(Tech = Total), by = "Year")

fig_072 <- ggplot(df_072, aes(Year)) +
  geom_area(aes(y = All),  fill = tpa_colors[2], alpha = ALPHA_AREA) +
  geom_area(aes(y = Tech), fill = tpa_colors[1], alpha = ALPHA_AREA) +
  geom_line(aes(y = All),  color = tpa_colors[2], linewidth = LINE_WIDTH) +
  geom_line(aes(y = Tech), color = tpa_colors[1], linewidth = LINE_WIDTH) +
  annotate("text", x = 2017, y = 150000, label = "All industries",
           color = "white", fontface = "bold", hjust = 0.5,
           family = FONT_FAMILY, size = DATA_LABEL_SIZE / .pt) +
  annotate("text", x = 2017, y = 50000, label = "Tech occupations\n(NAICS 54)",
           color = "white", fontface = "bold", hjust = 0.5, lineheight = 0.9,
           family = FONT_FAMILY, size = (DATA_LABEL_SIZE - 1) / .pt) +
  scale_x_years(df_072$Year, by = 2) +
  scale_y_continuous(labels = label_comma(), expand = EXPAND_BAR) +
  labs_meta(meta, y = "H-1B approvals") +
  theme_tpa()

save_chart(fig_072, "chapter6_fig_072")
```

### Chapter 6 - Figure 127 - Lose Pipeline
```{r fig_127}
# fig_127 (EVJ213) — chapter closer, by field.
# One cohort (2012-14 doctorate recipients), one horizon (~10-year U.S. stay
# rate), split into four STEM fields. Each field's departures valued at that
# field's own time to degree x the NSF GRFP support rate. Benchmark, not a
# per-person accounting (see note).

cohort <- read_fig("fig_127") %>%
  mutate(across(c(recipients, stay_rate_10yr, ttd_years), as.numeric))

COST_PER_YEAR <- 53000

fld <- cohort %>%
  mutate(
    stayed     = recipients * stay_rate_10yr,
    left       = recipients - stayed,
    left_share = left / recipients,
    cost_phd   = COST_PER_YEAR * ttd_years,
    value_left = left * cost_phd
  ) %>%
  arrange(left_share) %>%                       # lowest leak -> bottom, highest -> top
  mutate(field = factor(field, levels = field))

value_total <- sum(fld$value_left)              # ~3.94e9

# long form for stacking; Left leftmost so its right edge = departure count
seg <- fld %>%
  select(field, Stayed = stayed, Left = left) %>%
  pivot_longer(c(Stayed, Left), names_to = "outcome", values_to = "count") %>%
  mutate(outcome = factor(outcome, levels = c("Stayed", "Left")))

# leak label anchored just past the red segment's right edge (x = left)
leak_lab <- fld %>%
  transmute(field, left,
            llab = paste0(label_comma(accuracy = 100)(left),
                          "  (", round(left_share * 100), "%)"))

outcome_fill <- c("Stayed" = tpa_colors[9],   # muted
                  "Left"   = tpa_colors[1])   # emphasis red

# footnote appended under the data source (caption is ggtext; <br> = line break)
footnote <- paste0(
  "Note: Bars show one cohort (2012-14 doctorate recipients) at a single horizon ",
  "\u2014 the share still in the U.S. about ten years after graduation. The dollar ",
  "figures are a benchmark, not a per-person accounting: each field's departures are ",
  "valued at the NSF Graduate Research Fellowship support rate ($53,000 per ",
  "student-year, a $37,000 stipend plus a $16,000 education allowance) times that ",
  "field's median graduate-school time to degree (NSF Survey of Earned Doctorates, ",
  "2024, table 3-5), about $320,000-$370,000 per PhD. Composite field medians are ",
  "enrollment-weighted across constituent SED fields. The rate covers stipend and ",
  "tuition only and excludes laboratory, equipment, and institutional overhead, so it ",
  "understates full training cost. Most departing recipients held temporary visas, ",
  "funded through research-grant assistantships rather than the fellowship itself, ",
  "which costs a comparable amount per year."
)

meta <- get_meta("fig_127")
meta$caption <- paste0(meta$caption, "<br><br>", footnote)

x_max <- max(fld$recipients)

p_127 <- ggplot(seg, aes(x = count, y = field, fill = outcome)) +
  geom_col(width = 0.68, alpha = ALPHA_COL) +
  geom_text(
    data = leak_lab, inherit.aes = FALSE,
    aes(x = left, y = field, label = llab),
    hjust = -0.08, family = FONT_FAMILY, fontface = "bold",
    size = DATA_LABEL_SIZE / .pt, color = tpa_colors[1]
  ) +
  annotate(
    "text", x = 0, y = 0.32, hjust = 0,
    label = sprintf("\u2248 $%.1f billion in STEM training investment left with them",
                    value_total / 1e9),
    family = FONT_FAMILY, fontface = "bold",
    size = TITLE_SIZE / .pt, color = tpa_colors[1]
  ) +
  scale_x_continuous(labels = label_comma(), limits = c(0, x_max),
                     expand = expansion(mult = c(0, 0.05))) +
  scale_y_discrete(expand = expansion(add = c(0.95, 0.6))) +
  scale_fill_manual(values = outcome_fill, breaks = c("Stayed", "Left")) +
  labs_meta(meta, x = NULL, y = NULL) +
  theme_horizontal_bar()

save_chart(p_127, "fig_127", size = "standard")
```