title: "GTCI Tables"
author: "Emerson Johnston"
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = FALSE,
  warning   = FALSE,
  message   = FALSE,
  fig.width  = 8.0,
  fig.height = 5.5,
  dpi       = 300,
  fig.align = "center",
  dev       = "ragg_png"
)

library(tidyverse)
library(scales)
library(showtext)
library(ggtext)
library(ggrepel)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(gt)
library(gtExtras)
library(webshot2)
library(magick)
library(pdftools)

source("R/theme.R")
source("R/helpers.R")
source("R/tables.R")
source("R/chart_layers.R")
source("R/maps.R")
```

# Chapter 1 

### Chapter 1 - Table 6 - Change (1995-2024) in Int. Share of STEM Degrees Awarded by Field - Bachelors
```{r table 006}
# meta_a <- get_meta("tab_006a")
# meta_b <- get_meta("tab_006b")
# meta_c <- get_meta("tab_006c")
meta_d  <- get_meta("tab_006d")

# tab_006a <- build_change_table("tab_006a", meta_a)
# tab_006b <- build_change_table("tab_006b", meta_b)
# tab_006c <- build_change_table("tab_006c", meta_c)
tab_006d <- build_change_table_wide("tab_006d", meta_d)

# save_table(tab_006a, "chapter1_tab_006a")
# save_table(tab_006b, "chapter1_tab_006b")
# save_table(tab_006c, "chapter1_tab_006c")
save_table(tab_006d, "chapter1_tab_006d")
```

# Chapter 2 

### Chapter 2 - Table 7 - Change (1995-2024) in Int. Share of STEM Degrees Awarded by Field - Bachelors
```{r table 007}
meta_007 <- get_meta("tab_007")

tab_007 <- build_region_table(
  name = "tab_007",
  meta = meta_007
)

save_table(tab_007, "chapter2_tab_007")
```

### Chapter 2 - Table 8 - What Share of Each Country's International Students Are Enrolled in Key STEM Fields?
```{r table 008}
meta <- get_meta("tab_008")

field_files <- c(
  "Engineering"              = "fig_112",
  "Math & Computer Science"  = "fig_113",
  "Physical & Life Sciences" = "fig_114"
)

field_labels <- names(field_files)

read_field_long <- function(name) {
  read_fig(name) %>%
    mutate(across(-Year, ~ suppressWarnings(as.numeric(na_if(as.character(.x), "-"))))) %>%
    pivot_longer(-Year, names_to = "Country", values_to = "Share") %>%
    filter(!is.na(Share))
}

long_all <- imap(field_files, ~ read_field_long(.x) %>% mutate(Field = .y)) %>%
  bind_rows()

summ <- long_all %>%
  filter(Year %in% c(2009, 2024)) %>%
  mutate(point = if_else(Year == 2009, "First", "Last")) %>%
  select(Field, Country, point, Share) %>%
  pivot_wider(names_from = point, values_from = Share) %>%
  filter(!is.na(First), !is.na(Last)) %>%
  mutate(First = First * 100, Last = Last * 100,
         Change = (Last - First) / First * 100)

# keep only countries with both 2009 and 2024 values in all three fields
keep_countries <- summ %>% count(Country) %>% filter(n == 3) %>% pull(Country)
summ <- summ %>% filter(Country %in% keep_countries)

ord_cols <- as.vector(t(outer(field_labels, c("First", "Last", "Change"),
                              paste, sep = "_")))

country_order <- c(
  "China",
  "India",
  "South Korea",
  "Canada",
  "Japan",
  "Taiwan",
  "Saudi Arabia",
  "Mexico",
  "Vietnam",
  "Brazil",
  "Turkey/Türkiye",
  "Nigeria",
  "United Kingdom",
  "Germany",
  "Hong Kong",
  "Nepal",
  "Indonesia",
  "France",
  "Colombia",
  "Pakistan"
)

df_008 <- summ %>%
  pivot_wider(
    names_from  = Field,
    values_from = c(First, Last, Change),
    names_glue  = "{Field}_{.value}"
  ) %>%
  select(Country, all_of(ord_cols)) %>%
  arrange(match(Country, country_order)) %>%
  # ordered by mean field-share change; swap to e.g.
  # arrange(desc(`Engineering_Last`)) to order by 2024 engineering share
  mutate(.avg = rowMeans(across(ends_with("_Change")), na.rm = TRUE)) %>%
  select(-.avg)

tab_008 <- build_field_share_table(df_008, "tab_008", meta, field_labels)
save_table(tab_008, "chapter2_tab_008")
```

### Chapter 2 - Table 09 - International Share of U.S. Graduate Enrollment, by Field and Citizenship Status (1980–2024)
```{r tab_009}
meta <- get_meta("tab_009")

cite_files <- c(
  "Science"     = "fig_015",
  "Engineering" = "fig_016",
  "Health"      = "fig_017"
)

read_cite_long <- function(name) {
  read_fig(name) %>%
    mutate(across(-Year, ~ suppressWarnings(as.numeric(na_if(as.character(.x), "-"))))) %>%
    pivot_longer(-Year, names_to = "Group", values_to = "Count") %>%
    filter(!is.na(Count))
}

long_all <- imap(cite_files, ~ read_cite_long(.x) %>% mutate(Field = .y)) %>%
  bind_rows() %>%
  mutate(Group = recode(Group,
    "U.S. citizens and permanent residents" = "USC",
    "Temporary visa holders"                = "TVH"))

first_year <- min(long_all$Year, na.rm = TRUE)
last_year  <- max(long_all$Year, na.rm = TRUE)

# counts -> within-field citizenship share, per Field × Year
shares <- long_all %>%
  filter(Year %in% c(first_year, last_year)) %>%
  group_by(Field, Year) %>%
  mutate(Share = Count / sum(Count) * 100) %>%
  ungroup()

summ <- shares %>%
  mutate(point = if_else(Year == first_year, "First", "Last")) %>%
  select(Field, Group, point, Share) %>%
  pivot_wider(names_from = point, values_from = Share) %>%
  mutate(Change = (Last - First) / First * 100)   # relative change in share

ord_cols <- as.vector(t(outer(c("USC", "TVH"),
                              c("First", "Last", "Change"), paste, sep = "_")))

field_order <- c("Science", "Engineering", "Health")

df_009 <- summ %>%
  pivot_wider(
    names_from  = Group,
    values_from = c(First, Last, Change),
    names_glue  = "{Group}_{.value}"
  ) %>%
  select(Field, all_of(ord_cols)) %>%
  arrange(match(Field, field_order))

tab_009 <- build_citizenship_table(df_009, "tab_009", meta,
                        first_year = first_year, last_year = last_year,
                        value_is_share = TRUE,
                        footnote = paste(
                          "Within-field citizenship share of U.S. graduate enrollment in",
                          "science, engineering, and health. Change is the relative change",
                          "in share between", first_year, "and", last_year, "."))

save_table(tab_009, "chapter2_tab_009")
```

### Chapter 2 - Table 10 - Research Doctorate Recipients, by Field and Citizenship Status (2014–2024)
```{r table_010}
meta <- get_meta("tab_010")

phd_files <- c(
  "USC" = "fig_018",
  "TVH" = "fig_019"
)

read_phd_long <- function(name) {
  read_fig(name) %>%
    mutate(across(-Year, ~ suppressWarnings(as.numeric(na_if(as.character(.x), "-"))))) %>%
    pivot_longer(-Year, names_to = "Field", values_to = "Count") %>%
    filter(!is.na(Count)) %>%
    mutate(Field = recode(Field, "Social Science" = "Social Sciences"))
}

long_all <- imap(phd_files, ~ read_phd_long(.x) %>% mutate(Group = .y)) %>%
  bind_rows()

first_year <- min(long_all$Year, na.rm = TRUE)
last_year  <- max(long_all$Year, na.rm = TRUE)

summ <- long_all %>%
  filter(Year %in% c(first_year, last_year)) %>%
  mutate(point = if_else(Year == first_year, "First", "Last")) %>%
  select(Field, Group, point, Count) %>%
  pivot_wider(names_from = point, values_from = Count) %>%
  mutate(Change = (Last - First) / First * 100)

# wide counts: USC_First/Last/Change, TVH_First/Last/Change
counts_wide <- summ %>%
  pivot_wider(
    names_from  = Group,
    values_from = c(First, Last, Change),
    names_glue  = "{Group}_{.value}"
  )

# international share = TVH / (TVH + USC), per year, + relative change
df_010 <- counts_wide %>%
  mutate(
    SHR_First  = TVH_First / (TVH_First + USC_First) * 100,
    SHR_Last   = TVH_Last  / (TVH_Last  + USC_Last)  * 100,
    SHR_Change = (SHR_Last - SHR_First) / SHR_First * 100
  ) %>%
  select(Field,
         USC_First, USC_Last, USC_Change,
         TVH_First, TVH_Last, TVH_Change,
         SHR_First, SHR_Last, SHR_Change) %>%
  arrange(Field == "Non-sciences", desc(TVH_Last))

tab_010 <- build_citizenship_table(df_010, "tab_010", meta,
                        first_year = first_year, last_year = last_year,
                        value_is_share = FALSE,
                        share_group = TRUE,
                        footnote = paste(
                          "Research doctorate recipients by field and citizenship.",
                          "Change is the percent change between", first_year, "and", last_year, ".",
                          "International share is temporary visa holders as a percentage of",
                          "all recipients in the field; its change is relative.",
                          "Fields ordered by 2024 international count; non-sciences shown for comparison."))

save_table(tab_010, "chapter2_tab_010")
```

# Chapter 3 

### Chapter 3 - Table 11 - Temporary visa holder share of employment by occupation, 2017 / 2021 / 2023 
```{r}
meta <- get_meta("tab_011")

short_names <- c(
  "Biological, agricultural, and other life scientists" = "Biological scientists",
  "Computer and mathematical scientists"                = "Computer/math scientists",
  "Physical and related scientists"                     = "Physical scientists",
  "Social and related scientists"                        = "Social scientists",
  "Engineers"                                            = "Engineers",
  "S&E-related occupations"                               = "S&E-related occupations",
  "Non-S&E occupations"                                   = "Non-S&E occupations"
)

row_order <- c("Computer/math scientists", "Engineers", "Biological scientists",
               "Physical scientists", "Social scientists",
               "S&E-related occupations", "Non-S&E occupations")

df_011 <- read_fig("tab_011") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Occupation", values_to = "Rate") %>%
  filter(!is.na(Rate), Year %in% c(2017, 2019, 2021, 2023)) %>%
  mutate(Rate = Rate * 100,
         Occupation = recode(Occupation, !!!short_names)) %>%
  pivot_wider(names_from = Year, values_from = Rate, names_prefix = "Y") %>%
  mutate(
    Chg2019 = (Y2019 - Y2017) / Y2017 * 100,
    Chg2021 = (Y2021 - Y2019) / Y2019 * 100,
    Chg2023 = (Y2023 - Y2021) / Y2021 * 100,
    Change  = (Y2023 - Y2017) / Y2017 * 100,
    Occupation = factor(Occupation, levels = row_order)
  ) %>%
  arrange(Occupation) %>%
  select(Occupation, Y2017, Chg2019, Chg2021, Chg2023, Y2023, Change)

tab_011 <- build_occupation_share_table(
  df_011, "tab_011", meta,
  level_cols = c("Y2017", "Y2023"),
  chg_cols   = c("Chg2019", "Chg2021", "Chg2023")
)

save_table(tab_011, "chapter3_tab_011")
tab_011
```


### Chapter 3 - Table 12 - Top H1B Visa Employers in 2017 and 2025
```{r tab_012}
meta <- get_meta("tab_012")

shares <- fig030_shares()

# union of firms that were top-10 in EITHER year
union_firms <- shares %>%
  filter(Year %in% c(2017, 2025)) %>%
  group_by(Year) %>%
  slice_max(Count, n = 10) %>%
  ungroup() %>%
  distinct(Company) %>%
  pull(Company)

wide <- shares %>%
  filter(Year %in% c(2017, 2025), Company %in% union_firms) %>%
  select(Company, Year, Share) %>%
  pivot_wider(names_from = Year, values_from = Share, names_prefix = "Y")

df_012 <- wide %>%
  mutate(
    Company = recode(Company, !!!short_names),
    Y1 = Y2017, Y2 = Y2025,
    Change = if_else(Y2017 > 0, (Y2025 - Y2017) / Y2017 * 100, NA_real_)
  ) %>%
  arrange(desc(Change)) %>%          # order by 2025 share (today's leaders first)
  select(Company, Y1, Y2, Change)

tab_012 <- build_h1b_top10_table(df_012, "tab_012", meta,
                                 first_year = 2017, last_year = 2025)
save_table(tab_012, "chapter3_tab_012")
tab_012
```

### Chapter 3 - Table 13 - Labor Force Participation Rate of International Students as a Share of All Employed
```{r tab_013}
meta <- get_meta("tab_013")

# short row labels -- the raw CSV categories are very long
short_cats <- c(
  "Employed in the U.S. Labor Force in a STEM Occupation Directed Related to Science and Engineering"   = "STEM, directly S&E-related",
  "Employed in the U.S. Labor Force in a STEM Occupation Indirectly Related to Science and Engineering" = "STEM, indirectly S&E-related",
  "Employed in the U.S. Labor Force in a STEM Occupation with a STEM middle skill"                      = "STEM, middle-skill",
  "Employed in the U.S. Labor Force in a non-STEM Occupation"                                            = "Non-STEM occupation",
  "Unemployed in the U.S. Labor Force"                                                                   = "Unemployed",
  "Not in the U.S. labor force"                                                                          = "Not in labor force",
  "Total"                                                                                                = "Total"
)

row_order <- c("STEM, directly S&E-related", "STEM, indirectly S&E-related",
               "STEM, middle-skill", "Non-STEM occupation",
               "Unemployed", "Not in labor force", "Total")

df_013 <- read_fig("fig_025") %>%
  mutate(
    total_n = US_citizen + TPV,
    Citizen = US_citizen / total_n * 100,
    TVH     = TPV / total_n * 100,
    Category = recode(Category, !!!short_cats),
    Category = factor(Category, levels = row_order)
  ) %>%
  arrange(Category) %>%
  select(Category, Citizen, TVH)

tab_013 <- build_labor_citizenship_table(df_013, "tab_013", meta)

save_table(tab_013, "chapter3_tab_013")
tab_013
```

### Chapter 3 - Table 14 - STEM Degree to STEM Job Converstion Rate
```{r tab_14}
meta <- get_meta("tab_014")   # or reuse fig_029's meta

df_029t <- read_fig("fig_029") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Level", values_to = "Share") %>%
  filter(!is.na(Share), Year %in% c(1993, 2003, 2013, 2023)) %>%
  mutate(
    Share = Share * 100,
    Level = case_when(
      str_detect(Level, regex("bach", ignore_case = TRUE)) ~ "Bachelors",
      str_detect(Level, regex("mast", ignore_case = TRUE)) ~ "Masters",
      str_detect(Level, regex("doct", ignore_case = TRUE)) ~ "Doctorate"
    )
  ) %>%
  pivot_wider(names_from = Year, values_from = Share, names_prefix = "Y") %>%
  mutate(
    Chg2003 = (Y2003 - Y1993) / Y1993 * 100,
    Chg2013 = (Y2013 - Y2003) / Y2003 * 100,
    Chg2023 = (Y2023 - Y2013) / Y2013 * 100,
    Change  = (Y2023 - Y1993) / Y1993 * 100,
    Occupation = factor(Level, levels = c("Bachelors", "Masters", "Doctorate"))
  ) %>%
  arrange(Occupation) %>%
  select(Occupation, Y1993, Chg2003, Chg2013, Chg2023, Y2023, Change)

tab_014 <- build_occupation_share_table(
  df_029t, "tab_014", meta,
  level_cols    = c("Y1993", "Y2023"),
  chg_cols      = c("Chg2003", "Chg2013", "Chg2023"),
  spanner_label = "Share employed in STEM",
  footnote = paste(
    "1993 is the share of graduates at each degree level employed in a STEM",
    "occupation. 2003, 2013, and 2023 show the relative change from the prior",
    "period shown rather than the share itself. Net Change is the relative",
    "change in share from 1993 to 2023."
  )
)

save_table(tab_014, "chapter3_tab_014")
tab_014
```

# Chapter 4 

### Chapter 4 - Table 15 - How Many Times Do STEM Majors Change Majors
```{r tab_015}
meta <- get_meta("tab_015")
tab_015 <- build_bookend_table(
  "tab_015", meta,
  first = "2004/2009 Survey", last = "2012/2017 Survey",
  id_col = "Survey",
  first_label = "2004/2009", last_label = "2012/2017",
  spanner_label = "Share of STEM entrants",
  change_type = "relative",
  recode_cats = c("Never"          = "Never switched",
                  "Once"           = "Switched once",
                  "More than Once" = "Switched 2+ times"),
  cat_order = c("Never switched", "Switched once", "Switched 2+ times"),
  footnote = paste(
    "Share of STEM entrants by number of major switches. Net change is the",
    "relative change from the 2004/2009 to the 2012/2017 cohort."
  )
)
save_table(tab_015, "chapter4_tab_015")
tab_015
```

### Chapter 4 - Table 16 = What Do Americans Study Abroad?
```{r tab_016}
meta <- get_meta("tab_016")
tab_016 <- build_bookend_table(
  "fig_040", meta,
  first = 2005, last = 2024,
  spanner_label = "Share of students",
  change_type = "relative",
  sort_by = "change",
  recode_cats = c("STEM Fields"           = "STEM",
                  "Business & Management" = "Business & management",
                  "Social Sciences"       = "Social sciences",
                  "Fine and Applied Arts" = "Fine & applied arts"),
  flagged_rows = c("Social sciences", "Humanities", "Other"),
  flag_note = paste(
    "Open Doors revised its field taxonomy between 2013 and 2014; shares for",
    "these categories are not directly comparable across that break, so the",
    "reported change reflects reclassification as well as any real shift."
  ),
  footnote = paste(
    "Share of U.S. study-abroad students by field of study.",
    "Net change is the relative change from 2005 to 2024."
  )
)
save_table(tab_016, "chapter4_tab_016")
tab_016
```

### Chapter 4 - Table 17 - Where Do American's Study Abroad
```{r tab_017}
meta <- get_meta("tab_017")
tab_017 <- build_bookend_table(
  "fig_041", meta,
  first = 2005, last = 2024,
  spanner_label = "Share of students",
  change_type = "relative",
  sort_by = "last",
  drop_cols = "Antarctica",
  recode_cats = c("Africa, Sub-Saharan"   = "Sub-Saharan Africa",
                  "Multiple Destinations" = "Multiple destinations"),
  footnote = paste(
    "Share of U.S. study-abroad students by host region. Net change is the",
    "relative change from 2005 to 2024, which is volatile for small-share",
    "regions (e.g., the Middle East & North Africa moved 1.0% to 1.6%)."
  )
)
save_table(tab_017, "chapter4_tab_017")
tab_017
```
### Chapter 4 - Table 18 - Where Do American's Study Abroad
```{r tab_018}
meta <- get_meta("tab_018")   # or a dedicated tab_038a entry

tab_038a <- build_bookend_table(
  name          = "fig_038a",
  meta          = meta,
  first         = "2003/04",
  last          = "2011/12",
  id_col        = "cohort",
  spanner_label = "Share of STEM entrants",
  change_type   = "relative",
  recode_cats = c(
    stem_bachelors    = "Completed Bachelor's Degree in STEM Field",
    nonstem_bachelors = "Completed Bachelor's Degree in Another, Non-STEM Field",
    subbac_credential = "Completed A Subbac Credential",
    still_enrolled    = "Still Enrolled",
    left_no_degree    = "Left with No Degree"
  ),
  cat_order = c(
    "Completed Bachelor's Degree in STEM Field",
    "Completed Bachelor's Degree in Another, Non-STEM Field",
    "Completed A Subbac Credential",
    "Still Enrolled",
    "Left with No Degree"
  ),
  footnote = paste(
    "Values are the share of each entering STEM cohort ending in each outcome;",
    "each year column sums to 100%. Net change is the relative change from the",
    "2003/04 cohort to the 2011/12 cohort."
  )
)
save_table(tab_038a, "chapter4_tab_018")
```

# Chapter 5 

### Chapter 5 - Table 1 - US Universities in the Global Top 100 Over Time
```{r tab_001}
meta <- get_meta("tab_001")
num  <- function(x) parse_number(as.character(x))

field_order <- c("Physical Sciences", "Life Sciences", "Engineering", "Computer Science")
years_shown <- c(2014, 2018, 2022, 2026)

raw <- read_fig("tab_001") %>%
  rename(Year = 1) %>%
  mutate(Year = as.integer(num(Year)))

cs_2017 <- raw %>% filter(Year == 2017) %>% pull(`Computer Science`) %>% num()

df_001 <- raw %>%
  select(Year, all_of(field_order)) %>%                 # drop precomputed Average
  filter(Year %in% years_shown) %>%
  pivot_longer(-Year, names_to = "Field", values_to = "value") %>%
  mutate(value = num(value)) %>%
  pivot_wider(names_from = Year, values_from = value, names_prefix = "Y") %>%
  mutate(
    # computer science was first ranked in 2017; drop its 2017 count into the
    # 2014 slot so its change and net columns chain consistently
    Y2014   = if_else(Field == "Computer Science" & is.na(Y2014), cs_2017, Y2014),
    Chg2018 = Y2018 - Y2014,
    Chg2022 = Y2022 - Y2018,
    Chg2026 = Y2026 - Y2022,
    Change  = Y2026 - Y2014,
    Field   = factor(Field, levels = field_order)
  ) %>%
  arrange(Field) %>%
  select(Field, Y2014, Chg2018, Chg2022, Chg2026, Y2026, Change)

tab_001 <- build_ranking_delta_table(
  df_001, "tab_001", meta,
  level_cols  = c("Y2014", "Y2026"),
  chg_cols    = c("Chg2018", "Chg2022", "Chg2026"),
  change_type = "absolute",
  footnote = paste(
    "Counts are U.S. universities in the global top 100 in each field, in 2014 and",
    "2026. Interior columns show the change in that count from the prior shown year;",
    "Net Change is the change from 2014 to 2026."
  ),
  cell_footnote       = paste(
    "Computer science was first ranked in 2017; its 2017 count is shown here in the",
    "2014 column so its change and net columns can be computed."
  ),
  cell_footnote_field = "Computer Science",
  cell_footnote_col   = "Y2014"
)
save_table(tab_001, "chapter5_tab_001")
tab_001
```


### Chapter 5 - Table 19 - By Conference Publication Changes
```{r table 019}
conf_names <- c(
  nips    = "NeurIPS (ML / AI)",      icml   = "ICML (ML)",
  iclr    = "ICLR (ML)",              aistats = "AISTATS (Stats / ML)",
  acml    = "ACML (ML)",              colt   = "COLT (Theory)",
  uai     = "UAI (Probabilistic ML)", cvpr   = "CVPR (Vision)",
  iccv    = "ICCV (Vision)",          wacv   = "WACV (Vision)",
  acl     = "ACL (NLP)",              emnlp  = "EMNLP (NLP)",
  naacl   = "NAACL (NLP)",            coling = "COLING (NLP)",
  ijcai   = "IJCAI (AI)",             icra   = "ICRA (Robotics)",
  iros    = "IROS (Robotics)",        corl   = "CoRL (Robotics / ML)",
  rss     = "RSS (Robotics)",         siggraphasia = "SIGGRAPH Asia (Graphics)"
)

build_conference_summary_table(
  name       = "tab_019",     # output id; adjust to the next free table number
  meta       = get_meta("tab_019"),
  conf_names = conf_names,
  data_name  = "fig_048"      # reads fig_048.csv (the long author-share file)
) %>%
  save_table("chapter5_tab_019")
```

### Chapter 5 - Table 20 - Patents By Field, Country
```{r table 020}
build_patents_table(
  name    = "tab_020",     # output id; adjust to next free table number
  meta    = get_meta("tab_020"),
  sources = c(
    "Biotechnology"       = "fig_059",
    "Semiconductors"      = "fig_060",
    "Computer technology" = "fig_061"
  )
) %>%
  save_table("chapter5_tab_020")
```

### Chapter 5 - Table 21 - R&D Investment By Type
```{r table 020}
meta <- get_meta("tab_021")

type_labels <- c(
  "Basic Research"           = "Basic research (for tomorrow)",
  "Applied Research"          = "Applied research (for today)",
  "Experimental Development"  = "Experimental development (for today)"
)
type_order <- c("Basic research (for tomorrow)",
                "Applied research (for today)",
                "Experimental development (for today)")

df_125 <- read_fig("fig_125") %>%
  rename(Year = 1, Country = 2, Type = 3, value = 4) %>%
  mutate(Year  = as.numeric(as.character(Year)),
         value = as.numeric(value)) %>%
  filter(Year %in% c(1994, 2004, 2014, 2024)) %>%
  mutate(Type = recode(Type, !!!type_labels)) %>%
  pivot_wider(names_from = Year, values_from = value, names_prefix = "Y") %>%
  mutate(
    Chg2004 = (Y2004 - Y1994) / Y1994 * 100,
    Chg2014 = (Y2014 - Y2004) / Y2004 * 100,
    Chg2024 = (Y2024 - Y2014) / Y2014 * 100,
    Change  = (Y2024 - Y1994) / Y1994 * 100,
    Country = factor(Country, levels = c("United States", "China")),
    Type    = factor(Type, levels = type_order)
  ) %>%
  arrange(Country, Type) %>%
  select(Country, Type, Y1994, Chg2004, Chg2014, Chg2024, Y2024, Change)

tab_021 <- build_rd_type_delta_table(
  df_125, "tab_022", meta,
  level_cols = c("Y1994", "Y2024"),
  chg_cols   = c("Chg2004", "Chg2014", "Chg2024")
)
save_table(tab_021, "chapter5_tab_021")

```

### Chapter 5 - Table 22 - R&D Spending by Sector, U.S. and China
```{r tab_022}
meta <- get_meta("tab_022")

sector_labels <- c(
  "Business enterprise sector" = "Business enterprise",
  "Government sector"          = "Government",
  "Higher education sector"    = "Higher education",
  "Private non-profit sector"  = "Private non-profit"
)
sector_order <- c("Business enterprise", "Government",
                  "Higher education", "Private non-profit")

df_022 <- read_fig("tab_022") %>%
  rename(Year = 1, Country = 2, Type = 3, value = 4) %>%
  mutate(Year  = as.numeric(as.character(Year)),
         value = as.numeric(value)) %>%
  filter(Type %in% names(sector_labels),
         Year %in% c(1994, 2004, 2014, 2024)) %>%
  mutate(Type = recode(Type, !!!sector_labels)) %>%
  complete(Country, Type, Year, fill = list(value = 0)) %>%   # China private non-profit -> $0
  pivot_wider(names_from = Year, values_from = value, names_prefix = "Y") %>%
  mutate(
    Chg2004 = (Y2004 - Y1994) / Y1994 * 100,
    Chg2014 = (Y2014 - Y2004) / Y2004 * 100,
    Chg2024 = (Y2024 - Y2014) / Y2014 * 100,
    Change  = (Y2024 - Y1994) / Y1994 * 100,
    across(c(Chg2004, Chg2014, Chg2024, Change), ~ ifelse(is.finite(.), ., 0)),  # 0/0 -> 0%
    Country = factor(Country, levels = c("United States", "China")),
    Type    = factor(Type, levels = sector_order)
  ) %>%
  arrange(Country, Type) %>%
  select(Country, Type, Y1994, Chg2004, Chg2014, Chg2024, Y2024, Change)

tab_022 <- build_rd_type_delta_table(
  df_022, "tab_022", meta,
  level_cols = c("Y1994", "Y2024"),
  chg_cols   = c("Chg2004", "Chg2014", "Chg2024"),
  footnote = paste(
    "The first and last columns are spending levels in billions of US dollars (PPP",
    "converted). The interior columns show the relative change from the prior shown",
    "year rather than the level. Net Change is the relative change across the full",
    "window. Sectors denote where R&D is performed. China reports no private non-profit",
    "R&D; it is shown as $0, and its relative-change cells (change from a zero base is",
    "undefined) appear as 0%."
  )
)
save_table(tab_022, "chapter5_tab_022")
tab_022
```

### Chapter 5 - Table 23 - Patents by Company
```{r tab_023}
meta <- get_meta("tab_023")
num  <- function(x) parse_number(as.character(x))

flag_slug <- c(
  "United States" = "united-states",
  "Japan"         = "japan",
  "Taiwan"        = "taiwan",
  "South Korea"   = "south-korea",
  "Germany"       = "germany",
  "China"         = "china"
)

df_023 <- read_fig("tab_023") %>%
  rename(Company = 1, Country = 2, Y2005 = 3, Y2015 = 4, Y2025 = 5) %>%
  mutate(across(c(Y2005, Y2015, Y2025), num)) %>%
  mutate(
    flag    = paste0("flags/", recode(Country, !!!flag_slug), ".png"),
    Chg2015 = Y2015 - Y2005,
    Chg2025 = Y2025 - Y2015,
    Change  = Y2025 - Y2005
  ) %>%
  arrange(desc(Change)) %>%
  select(Company, flag, Y2005, Chg2015, Y2015, Chg2025, Y2025, Change)

tab_023 <- build_patent_company_delta_table(
  df_023, "tab_023", meta,
  level_cols  = c("Y2005", "Y2015", "Y2025"),
  chg_cols    = c("Chg2015", "Chg2025"),
  change_type = "absolute"
)
save_table(tab_023, "chapter5_tab_023")
```

### Chapter 5 - Table 24 - Count and Share of Top Researchers
```{r tab_024}
meta <- get_meta("tab_024")   # add a tab_046 row to titles_and_sources

tab_024 <- build_count_share_bookend_table(
  name  = "fig_046",          # reads the same CSV as the figure
  meta  = meta,
  first = 2014,
  last  = 2025,
  entities      = list("United States" = "US", "China" = "CN"),
  count_spanner = "Researchers in global top 1%",
  share_spanner = "Share of global top 1%",
  footnote = paste(
    "Count is the number of top-1% most-cited researchers with a primary",
    "affiliation in each country; share is that country's percentage of the",
    "global top 1%. Change columns are the absolute change in count and the",
    "percentage-point change in share between 2014 and 2025."
  )
)

save_table(tab_024, "chapter5_tab_024")
```

# Chapter 6

### Chapter 6 - Table 25 - Net Change in Intentions 
```{r table 025}
meta <- get_meta("tab_025")

tab_025 <- build_bookend_table(
  "fig_068", meta,
  first = 2021, last = 2024,
  spanner_label = "Share with U.S. plans",
  sort_by = "change",
  scale = 100
)

save_table(tab_025, "chapter6_tab_025")
```

### Chapter 6 - Table 26 - Employment-Based Visa Processing Wait Times
```{r table 026}
meta <- get_meta("tab_026")

tab_026 <- build_eb_combined_table(meta)

save_table(tab_003, "chapter6_tab_026")
tab_026
```

### Chapter 6 - Table 27 - Employment-Based Visa Processing Wait Times
```{r table 027}
meta <- get_meta("tab_027")

tab_027 <- build_sector_bookend_table(meta)

save_table(tab_027, "chapter6_tab_027")
tab_027
```

### Chapter 6 - Table 28 - Share Intending to Stay in the U.S., by World Region
```{r table 028}
meta <- get_meta("tab_028")

tab_028 <- build_bookend_table(
  "fig_066", meta,
  first = 2002, last = 2024,
  spanner_label = "Intend to stay",
  scale = 100,
  drop_cols = "Region and country or economy unknown",
  change_type = "ppt"
)

save_table(tab_066, "chapter6_tab_028")
tab_028
```

### Chapter 6 - Table 29 - How Many PERM Certifications Go Unused
```{r table 029}
meta <- get_meta("tab_029")

tab_029 <- build_perm_expiry_table(meta)

save_table(tab_029, "chapter6_tab_029")
tab_029
```